### 2023-11-11 lit.qmd
Initial commit

---
title: "Lit Review"
author: "Susan Vanderplas"
format: html
editor: visual
bibliography: references.bib
---

## Screwdrivers/Chisels

-   @garcia2017

    -   Overall Rating: 🌈

    -   Parameter definitions:\
        ![](fig/Garcia_2017_fig1.png)

        -   Angle of attack $\alpha$ vs. axial rotation $\gamma$

    -   Experimental Design:

        -   5 Sencys CR-V 10x200mm electrician's chisels (tip = 9.35x1.85mm)

        -   $\alpha = 0$

        -   $\gamma = \{0, 15, 30, 45, 60, 75\}^\circ$

        -   2 repeat observations

        -   marks were cast, scanned with 20x objective magnification, 800nm vertical resolution, $2\mu m$ lateral resolution, \~10x0.7mm scanning areas, 32 bit precision.

    -   "Tool marks have to be created in a consistent manner that reduces factors that might induce variation in or between marks. Most of these factors are caused by the manual creation of marks" (pg 206)

        -   does this mean that manual creation of marks is less reliable?

        -   Can toolmarks hold up to practical situations where both the examiner and the accused would have made manual marks?

    -   Results:

        -   Based on a preliminary investigation it was observed that the tool mark width (y-axis) compressed with increasing axial rotation. Based on the widths of the experimental marks the average deviation from the theoretical compression factors (1.000, 0.966, 0.866, 0.707, 0.500 and, 0.258) was calculated to be 0.792% $\pm$ 2.068

        -   Without resizing, can only compare $0^\circ$ to $15^\circ$ within guidelines; with resampling and resizing, can compare across the $\gamma$ values chosen

        -   Used wavelength filters to compare structural similarities:

            -   Wavelength ranges included were 1000--5, 1000--200, 200--100, 100--50, 50--25, 25--10 and 10--5 mm. This resulted in a similarity scores which depended solely on the comparison of structural details belonging to a certain wavelength range.

        -   Without resizing:![](fig/Garcia_2017_tab1.png)

        -   With resizing:\
            ![](fig/Garcia_2017_tab2-3.png)

        -   ![](fig/Garcia_2017_fig12.png)

        -   Results somewhat dependent on noise removal threshold

-   @bachrach2010

-   @yang2019

-   @petraco2012

## Wire Cutters

-   @heikkinen2014

    -   Overall rating: 💩💩

    -   Introduction

        -   When diagonal cutters are used to cut wires, they leave marks that were created by a part of the blade on the cut surface (4,5). Small (less than mm2) partial marks may not allow matching the tool to a mark using standard comparison microscopy (e.g., Leica FSC used in this study \[8\]) because there may be too few recognizable features to declare a match or nonmatch with sufficient certainty [@nichols2]. 3D imaging provides a potential way to address this issue: It allows seeing small features and potentially recognizing a match from a small partial toolmark. With precise 3D imaging, fewer striations are needed to declare that two marks were made using the same tool (Table 1). [@nichols2]

            -   Quote from @nichols2:

                > In three dimensional toolmarks when at least two different groups of at least three consecutive matching striae appear in the same relative position, or one group of six consecutive matching striae are in agreement in an evidence toolmark compared to a test toolmark. In two dimensional toolmarks when at least two groups of at least five consecutive matching striae appear in the same relative position, or one group of eight consecutive matching striae are in agreement in an evidence toolmark. For these criteria to apply, however, the possibility of sub-class characteristics must be ruled out.

    -   Design:

        -   2mm wires cut with 2 different diagonal cutters from the same production batch and what seems to be a 3rd cutter that wasn't from the batch? (description NOT clear)

        -   blades marked to ensure cuts were with same part of the blade (5 cutting positions, but only 3 )

        -   3D optical imaging

    -   Results

        -   74/80 profiles identified using CMS (even using images taken w/ different 3D scanners)

        -   Wear is visible after 500 cuts but can still be assessed using CMS (and from Fig 7, probably using crosscorr as well)

        -   Different blades have different characteristics

        -   different parts of the same blade have different characteristics

        -   3D imaging provides more details relative to 2D imaging

## Bolt Cutters

-   @butcher1975

-   @yang2019

## Pliers

-   @bachrach2010

-   @kassamakov2010

-   @yang2019

### 2023-11-11 shop_tools.qmd
Initial commit

---
title: "A Standard Engineer's Shop"
format: html
editor: visual
bibliography: references.bib
---

## Wire Cutting Surfaces

Basic Parameters

-   We will only count things someone might reasonably use to cut wires "on the fly" - scissors, x-acto knives, wire cutters, pliers, and so on. We excluded saw blades, most knives on pocket knives and multitools, chisels, screwdrivers, wrenches, and so on.

-   We're only counting tools which are easily available (not ransacking the whole shop to do this analysis). I did include the scissors available throughout the house, but this isn't a huge factor in the results of the analysis.

::: {#fig-shop-layout layout-ncol=2}

![Shop Bench Tool rack](fig/PXL_20231110_190517827.jpg)

![Shop Side Tool rack](fig/PXL_20231110_190522765.jpg)

We consider only tools which might be easily used for wire cutting that are easily available.

:::

```{r}
#| label: tbl-shop-surfaces
#| tbl-cap: "Shop cutting surfaces. All lengths in cm."
#| echo: false
#| message: false
library(tibble)
library(dplyr)

cutting_surfaces <- tribble(
  ~type, ~length, ~surfaces, ~tools,
  "Pliers", 1, 2, 10,
  "Needlenose pliers", 1, 4, 3,
  "Side cutters", 1, 4, 1,
  "Side cutters", 2, 4, 2,
  "Scissors", 8.89, 2, 14,
  "Tin snips", 3.81, 2, 4,
  "Lineman pliers", 2, 4, 2,
  "Lineman pliers", 1, 4, 1,
  "Wire strippers", 1, 2, 2,
  "Kneep tongs", 2, 4, 1,
  "Kneep tongs", 3, 4, 1,
  "Utility knife blades", 7.62, 2, 50
) %>%
  mutate(total_cutting_length = length*surfaces*tools)

total_cutting_surface = sum(cutting_surfaces$total_cutting_length)

knitr::kable(cutting_surfaces)
```

In our shop we have approximately `r total_cutting_surface` cm of wire cutting surface, as shown in @tbl-shop-surfaces.

## Stranded Wire

Using the standard stranded wire that we happen to have in our electrical toolbox, we can measure the approximate pixels of each wire strand and compare to the approximate pixel measurement of the 1cm rule shown on the right side of the picture below. This is not precise to forensic standards, but is sufficient for an order-of-magnitude calculation.

![A sample of stranded wire of unknown gauge.](fig/wire_stranded.png)

```{r}
#| echo: false
#| message: false
wire_meas <- c(
  5.8, 5.4, 5.4, 5.4, 6.1, 7.3, 6.1, 5.0, 6.0, 7.0, 6.0, 6.1, 5.0
)

cm_meas <- 346.4
wire_thickness_cm <- mean(wire_meas)/cm_meas

comparisons <- ceiling(total_cutting_surface/wire_thickness_cm)
```

Using the average of the `r length(wire_meas)` strands, we come up with a mean wire diameter of `r wire_thickness_cm` cm.

## Fundamental Problem

If a forensic analyst were to compare the end of a single fragment of stranded wire recovered from a crime scene to all of the tools easily available in our shop, he or she would have to do at least `r comparisons` comparisons, and likely multiple times this number to account for e.g. replication of marks, variability, partial offsets, etc.

We know that factors like cutting angle also make a difference in terms of what features are recorded, but for the moment we'll ignore this factor.

Even if we accept that the error rate for toolmark comparisons is extremely low, e.g. 0.003-0.005 as in @murdock2017 we can work out the chance of a random match in all of the comparisons as follows:

\begin{align}
1 \text{ comp} & P(\text{Random match}) = 0.003 \\
& P(\text{No random match}) = 0.997\\
N \text{ comp} &  P(\text{no random matches in} N \text{ comp}) = 0.997^N\\
& P(\text{at least one random match in} N \text{ comp}) = 1 - 0.997^N = `r sprintf("%.06f", 1-0.997^comparisons)`
\end{align}

In fact, if we want to control for the overall random match probability amid $N$ comparisons, we can calculate what our estimated RMP for a single comparison might be:

Let $R_n:=\{\text{at least one random match in }n\text{ comparisons}\}$

\begin{align}
P(R_1) = r
P(\overline{R_1}) = 1-r
P(R_n) = \left[1-r\right]^n
P(\overline{R_n}) = 1 - \left[P(\overline{R_1})\right]^n = 1 - \left[1 - r\right]^n
\end{align}

Thus, if we wish to control $P(R_n)$, we can calculate out $r$ accordingly: $$r = 1 - \left[1-P(R_n)\right]^{1/n}$$

```{r}
#| echo: false
#| message: false
library(ggplot2)
library(tidyr)
single_rmp = exp(seq(-1000, 0, .1))
N = c(10, 100, 1000, 10000, 100000)

res = crossing(single_rmp, N) %>%
  mutate(overall = 1 - (1 - single_rmp)^N)

ggplot(res, aes(x = single_rmp, y = overall, group = factor(N), color = factor(N))) + 
  geom_line() +
  coord_cartesian(xlim = c(0, 0.005), ylim = c(0.05, 1)) + 
  ylab("Probability of at least one random match\nin N comparisons") + 
  xlab("Probability of random match for 1 comparison") + 
  ggtitle("Wire Comparison Overall Random Match Probability") + 
  scale_color_discrete("# comparisons") + 
  theme_bw()

```

Fundamentally, once you start doing comparisons on wire cutting surfaces, you're in deep trouble - the total length of surfaces you might test in an investigation is large relative to the surface diameter of the evidence you may have. As a result, the probability of a random match is hard to control even with the assumption that a random match is extremely unlikely.

### 2023-11-11 Preparation/shop_tools.qmd
Reorganizing

--- 2023-11-11 shop_tools.qmd
@@
 editor: visual
-bibliography: references.bib
+bibliography: ../references.bib
 ---
@@
 
-::: {#fig-shop-layout layout-ncol=2}
+::: {#fig-shop-layout layout-ncol="2"}
+![Shop Bench Tool rack](../fig/PXL_20231110_190517827.jpg)
 
-![Shop Bench Tool rack](fig/PXL_20231110_190517827.jpg)
-
-![Shop Side Tool rack](fig/PXL_20231110_190522765.jpg)
+![Shop Side Tool rack](../fig/PXL_20231110_190522765.jpg)
 
 We consider only tools which might be easily used for wire cutting that are easily available.
-
 :::
@@
 
-![A sample of stranded wire of unknown gauge.](fig/wire_stranded.png)
+![A sample of stranded wire of unknown gauge.](../fig/wire_stranded.png)
 
@@
 
+```{=tex}
 \begin{align}
@@
 \end{align}
-
+```
 In fact, if we want to control for the overall random match probability amid $N$ comparisons, we can calculate what our estimated RMP for a single comparison might be:
@@
 \end{align}
-
+```
 Thus, if we wish to control $P(R_n)$, we can calculate out $r$ accordingly: $$r = 1 - \left[1-P(R_n)\right]^{1/n}$$

### 2023-11-11 PNAS/index.Rnw
PNAS template

\documentclass[9pt,twocolumn,twoside]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasbriefreport} % Choose template
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\title{Template for preparing your Brief Report submission to PNAS using Overleaf}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1,2]{Susan Vanderplas}
\author[b,1]{Heike Hofmann}
\author[b]{Alicia Carriquiry}

\affil[a]{Statistics Department, University of Nebraska Lincoln. 350 Hardin Hall, 3310 Holdrege North Wing, Lincoln, NE 68503}
\affil[b]{Affiliation Two}
\affil[c]{Affiliation Three}

% Please give the surname of the lead author for the running footer
\leadauthor{Lead author last name}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{Please provide details of author contributions here.}
\authordeclaration{Please declare any competing interests here.}
\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: susan.vanderplas@unl.edu}

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol
\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}

\begin{abstract}
Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\firstpage{12}
% Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting. In this example, there are a total of 11 paragraphs on the first page, counting the first level heading as a paragraph. The value {12} represents the number of the paragraph starting the second page. If a paragraph runs over onto the second page, include a bracket with the paragraph line number starting the second page, followed by the paragraph number in curly brackets, e.g. "\firstpage[4]{11}".

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.
\dropcap{T}his PNAS journal template is provided to help you write your work in the correct journal format. Instructions for use are provided below.

Note: please start your introduction without including the word ``Introduction'' as a section heading (except for math articles in the Physical Sciences section); this heading is implied in the first paragraph.

\section*{Guide to using this template on Overleaf}

Please note that whilst this template provides a preview of the typeset manuscript for submission, to help in this preparation, it will not necessarily be the final publication layout. For more detailed information please see the \href{ https://www.pnas.org/page/authors/format}{PNAS Information for Authors}.

If you have a question while using this template on Overleaf, please use the help menu (``?'') on the top bar to search for \href{https://www.overleaf.com/help}{help and tutorials}. You can also \href{https://www.overleaf.com/contact}{contact the Overleaf support team} at any time with specific questions about your manuscript or feedback on the template.

\subsection*{Author Affiliations}

Include department, institution, and complete address, with the ZIP/postal code, for each author. Use lower case letters to match authors with institutions, as shown in the example. PNAS strongly encourages authors to supply an \href{https://orcid.org/}{ORCID identifier} for each author. Individual authors must link their ORCID account to their PNAS account at \href{http://www.pnascentral.org/}{www.pnascentral.org}. For proper authentication, authors must provide their ORCID at submission and are not permitted to add ORCIDs on proofs.

\subsection*{Submitting Manuscripts}

All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}. If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.

\subsection*{Format}

Many authors find it useful to organize their manuscripts with the following order of sections: title, author line and affiliations, keywords, abstract, introduction, results, discussion, materials and methods, acknowledgments, and references. Other orders and headings are permitted.

\subsection*{Manuscript Length}

Brief Reports are limited to 3 pages, which is approximately 1,600 words (including the manuscript text, title page, abstract, and figure legends), and 15 references. Supporting information (SI) is limited to extended methods, essential supporting datasets, and videos (no additional tables or figures).

\subsection*{References}

References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g. \cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric,phdthesis,masterthesis}. All references cited in the main text should be included in the main manuscript file.

\subsection*{Data Archival}

PNAS must be able to archive the data essential to a published article. Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the \href{https://www.pnas.org/author-center/editorial-and-journal-policies#materials-and-data-availability}{Information for Authors}, is acceptable.

\subsection*{Language-Editing Services}
Prior to submission, authors who believe their manuscripts would benefit from professional editing are encouraged to use a language-editing service (see list at https://www.pnas.org/author-center/language-editing). PNAS does not take responsibility for or endorse these services, and their use has no bearing on acceptance of a manuscript for publication.

\begin{figure}[t!]
\centering
\includegraphics[width=.8\linewidth]{frog}
\caption{Placeholder image of a frog with a long example legend to show justification setting.}
\label{fig:frog}
\end{figure}

\begin{SCfigure*}[\sidecaptionrelwidth][t]
\centering
\includegraphics[width=11.4cm,height=11.4cm]{frog}
\caption{This legend would be placed at the side of the figure, rather than below it.}\label{fig:side}
\end{SCfigure*}

\begin{table}[t!]
\centering
\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
\begin{tabular}{lrrr}
Species & CBS & CV & G3 \\
\midrule
1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
\bottomrule
\end{tabular}

\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
\end{table}

\subsection*{Digital Figures}

EPS, high-resolution PDF, and PowerPoint are preferred formats for figures that will be used in the main manuscript. Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format. Color images must be in RGB (red, green, blue) mode. Include the font files for any text.

Images must be provided at final size, preferably 1 column width (8.7cm). Figures wider than 1 column should be sized to 11.4cm or 17.8cm wide. Numbers, letters, and symbols should be no smaller than 6 points (2mm) and no larger than 12 points (6mm) after reduction and must be consistent.

Figures and tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.

Figure \ref{fig:frog} shows an example of how to insert a column-wide figure. To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment. Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide. Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side legends.

\subsection*{Tables}
Tables should be included in the main manuscript file and should not be uploaded separately.

\subsection*{Single column equations}

Authors may use 1- or 2-column equations in their article, according to their preference.

To allow an equation to span both columns, use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures.

Note that the use of the \verb|widetext| environment for equations is not recommended, and should not be used.

\begin{figure*}[bt!]
\begin{align*}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3.
\end{align*}
\end{figure*}

\subsection*{Supporting Information (SI)}

Authors are limited to the following types of SI: extended methods, datasets, videos, and 3D figures. Extended discussion is not permitted.

\subsection*{SI Appendix}

Supply extended materials and methods information in a separate SI Appendix PDF file. Include SI movie legends. SI will be published as provided by the authors; it will not be edited or composed. Supplementary figures and tables are not allowed.

\subsubsection*{SI Datasets}

Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be published in raw format and will not be edited or composed.

\subsubsection*{SI Movies}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and and include a brief legend for each movie in the main manuscript file. All movies should be submitted at the desired reproduction size and length. Movies should be no more than 10 MB in size.

\matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required.

\subsection*{Subsection for Method}
Example text for subsection.
}

\showmatmethods{} % Display the Materials and Methods section

\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}

\showacknow{} % Display the acknowledgments section

\bibsplit[3]
%Use \bibsplit to split the references from the body of the text. Value "[3]" represents the number of reference in the left column (Note: Please avoid single column figures & tables on this page.)

% Bibliography
\bibliography{pnas-sample}

\end{document}

### 2023-11-12 LPR/index.qmd
LPR skeleton draft

---
title: "Multiple Comparisons in Toolmark Evidence"
author: "Susan Vanderplas, Heike Hofmann"
format: pdf
---
```{r setup, include = F}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
```

## Introduction

-   Problems with database searches and multiple comparisons in forensics
-   Fingerprint - madrid bombing
-   DNA databases
-   Firearms and toolmarks?? NIBIN/IBIS issues, but more broadly there are issues

## Thought Experiment

```{r shop-data}
library(tibble)
library(dplyr)

cutting_surfaces <- tribble(
  ~type, ~length, ~surfaces, ~tools,
  "Pliers", 1.5, 2, 10,
  "Needlenose pliers", 1, 4, 3,
  "Side cutters", 1, 4, 1,
  "Side cutters", 2, 4, 2,
  "Scissors", 8.89, 2, 14,
  "Tin snips", 3.81, 2, 4,
  "Lineman pliers", 2, 4, 2,
  "Lineman pliers", 1, 4, 1,
  "Wire strippers", 1, 2, 2,
  "Long handled cutting pliers", 2, 4, 1,
  "Long handled cutting pliers", 3, 4, 1,
  "Utility knife blades", 7.62, 2, 50
) %>%
  mutate(total_cutting_length = length*surfaces*tools)

total_cutting_surface = sum(cutting_surfaces$total_cutting_length)
```

Consider, for example a situation where someone builds a bomb, and a wire fragment is recovered from un-detonated explosive. We will examine how the forensic examination might proceed in three different cases:

1.  Examiners identify a suspect and locate a single tool which is believed to have been used to construct the bomb. Examiners would like to determine whether this tool is a forensic match to wires within the undetonated bomb.

2.  Examiners identify a suspect and locate a garage full of tools, one of which may have been used to construct the bomb. Examiners would like to identify which tool(s) were used to cut the wires during bomb construction.

3.  Examiners identify several suspect, each of whom has a set of tools, and wish to determine which suspect(s) may have constructed the bomb based off of the toolmark comparisons.

We will consider wires ranging in diameter from 2mm (solid 12 gauge wire) to 0.0116 mm (stranded 12 gauge wire may be composed of 49 strands of 0.0116 mm diameter wire) for the purposes of this thought experiment; these values represent wire easily obtainable at a local hardware store.

<!-- https://www.calmont.com/wp-content/uploads/calmont-eng-wire-gauge.pdf -->

<!-- https://en.wikipedia.org/wiki/American_wire_gauge -->

### Single Tool Comparisons

Let us first consider the situation where a single tool is recovered from the suspect's house that is viewed as likely to have made the wire cut. Then, a forensic examiner might take the tool and use it to cut a piece of metal, known as a blade, in order to record the markings of the entire cutting surface simultaneously. This process might be repeated at multiple angles in order to record the effect of different contact surfaces on the resulting striations. For simplicity, let us assume that angles are not a factor. Let us further define the blade to be of length $\ell$ and the wire to have diameter $d$ which is fully covered with striations that are untouched by the blast.

At minimum, the examiner will have to do $\ell/d$ comparisons in order to assess whether the tool matches the wire; when a comparison algorithm is used, and both the blade and the wire are scanned at the same resolution of $r$ mm, there may be as many as $\ell/r- d/r + 1$ comparisons performed automatically in order to find the optimal alignment between the two samples[^1].

[^1]: This assumes we restrict the wire to be completely contained within the bounds of the blade surface. If we allow for partial matches and restrict these matches to require at least $k$ points of overlap, then this becomes $\ell/r - d/r + k + 1$.

To make this more concrete, consider the pair of pliers shown in @fig-wire-cutter, which has a 1.5 cm cutting surface machined on both sides to produce a peaked cross-section. When used to cut a wire, the wire rests against a rectangular surface used to hold it in place; the blade is pushed into the wire, producing striae resulting from imperfections in side A on one half of the wire and striae from imperfections and wear in side B on the other.

![Wire cutter with 1.5 cm cutting surface beveled on both sides](../fig/wire_cutter.png){#fig-wire-cutter}

A scan of a cut made by the blade at $0.625\mu m$ resolution would produce approximately $15/0.0000625 = 240000$ observations; a scan of a 2mm wire would produce $2/0.0000625 = 32000$ observations, and computationally aligning these two data sequences using maximum cross-correlation would require $240000 - 32000 + 1 = 208001$ comparisons. Under an optical microscope, an examiner might move the wire at 0.2mm increments along the blade cut surface for a total of $15/.2 - 2/.2 + 1 = 66$ distinct and discrete comparisons. Then, we must consider that the examiner or algorithm must compare the wire to both side A and side B; we then double the number of total comparisons made to $416002$ and $132$ for the algorithm and the examiner, respectively.

Ultimately, in order to compare tool marks in this scenario, the examiner or algorithm completes the equivalent to a database search resulting from a single piece of evidence. The dangers of such searches are well known and have been cited in forensic missteps such as the Madrid bombing case, as discussed in @li2023; @newman2007; @mayfield2006report [Pg 137].

While 132 visual comparisons might not seem so terrible, this is the simplest of the 3 scenarios we consider; each scenario presents a progressively more complex problem.

### A Garage of Comparisons

In our next scenario, investigators identify a suspect and collect all of the tools in the suspect's residence which might have been used to construct the bomb.

Let us consider the garage of one of the authors of this paper; @fig-shop-layout shows the garage work bench and tool storage in typical condition. While there is no reason to think the author's garage is representative of all garages, the general setup and number of tools in the garage does not seem unusual for a suburban household where residents may dabble in automotive repair, home improvement, or woodworking.

::: {#fig-shop-layout layout-ncol="2"}
![Shop Bench Tool rack](../fig/PXL_20231110_190517827.jpg)

![Shop Side Tool rack](../fig/PXL_20231110_190522765.jpg)

Layout of tool storage in the author's garage. For this study, we considered only tools which might reasonably be used to cut wires.
:::

```{r}
#| label: tbl-shop-surfaces
#| tbl-cap: "Shop cutting surfaces. All lengths in cm."
#| echo: false
#| message: false
knitr::kable(cutting_surfaces)
```
```{r}
#| label: tbl-shop-comparisons
#| tbl-cap: "Comparions performed by tool type, assuming digital scans at a resolution of $0.625\\mu m$ and visual comparisons performed at $0.2 mm$ intervals along the test blade impression."
#| echo: false
#| message: false
comparisons <- cutting_surfaces %>%
  mutate(digital = (length*10/0.0000625 - 2/0.0000625+1)*surfaces*tools,
         analog = (length*10/.2 - 2/.2 + 1)) %>%
  group_by(type) %>%
  summarize(digital = sum(digital), analog = sum(analog)) %>%
  rename(`Tool type` = type, Algorithm = digital, Examiner = analog)

knitr::kable(comparisons)
```

We assembled the easily available tools from the house and garage which might reasonably be used to cut wires[^2], including pliers, wire cutters and strippers, scissors, tongs, and utility knives. Estimated cutting surface length and quantity of each type of tool is provided in @tbl-shop-surfaces. The total length of all cutting surfaces of relevant tools in the author's garage is `r total_cutting_surface` cm; primary contributions to the total length are the 50 pack of utility knife blades and the 14 pairs of full-size scissors which could be located throughout the house. Of course, a thorough search of the garage and house would come up with far more cutting surfaces to compare against during an actual crime investigation.

[^2]: Reasonably is defined as we have used a tool like this to cut wires during electrical work in the past.

As in Scenario 1, the wire fragment recovered from the un-detonated explosive would be compared to blade surface test impressions generated from each cutting surface in the garage. However, unlike in Scenario 1, we have `r total_cutting_surface` cm of cutting surfaces to work with. We can apply our comparison formula on a by-surface basis to yield the digital and analog comparison estimates shown in @tbl-shop-comparisons. Overall, our algorithm would make `r sum(comparisons$Algorithm)` comparisons and an examiner doing a visual inspection would make `r sum(comparisons$Examiner)` comparisons under the assumptions we made in Scenario 1. XXX fill this in with proper formatting XXX

### Multiple Suspects

## The Problems with Database Searches

### Random Match Probability

### Score-based Distributions

## Potential Solutions

We need to have an explicit relationship between random match probability and amount of signal (assessed by length, number of features, overall striae depth, etc.). Until we have this kind of relationship, it will be hard to use error rates (even if they were known for toolmarks) from algorithms or  black-box studies in practical situations.

We need much more study of real data, examiner evaluations, and performance of algorithms in situations which mimic real casework. We also need to quantify the strength, distinctiveness, and total number of striations to ensure that comparisons are being made only in situations where random matches are unlikely to occur.

### 2023-11-13 LPR/index.qmd
Add refs and write part of intro

@@
 ---
+
 ```{r setup, include = F}
@@
 -   Firearms and toolmarks?? NIBIN/IBIS issues, but more broadly there are issues
+
+One of the common suggestions in the 2016 PCAST report across pattern disciplines in forensic science was the importance of creating huge databases of e.g.
+known prints and firearms to support research into matches and closest non-matches\[@pcast, fingerprints pg 11, 88, firearms pg 12\].
+As researchers, we concur with this suggestion, as it is critical for developing objective algorithms and creating useful black-box studies.
+We have spent considerable time assembling large data sets of tool marks, bullets, and cartridge cases to support our research as well as create public resources to accelerate future research \[@csafeDataPortal\].
+However, we are wary about the implications of database searches across forensic disciplines, for reasons also mentioned in the PCAST report: the necessity of understanding the random match probability for searches against these databases, in particular when using degraded evidence often found in crime scenes\[@pcast, pg 52\].
+
+This issue has been raised in the past with regard to DNA [@thompson2003] and fingerprints [@fine2006]; one of the many issues identified in the aftermath of the 2004 Madrid bombing case was that the IAFIS database used to locate similar prints is extremely large and thus it is possible for the database to locate unusually similar non-matches.
+This issue will almost certainly become an issue in other pattern disciplines as databases of firearms evidence such as NIBIN @nibin2021 become more commonly utilized by law enforcement.
+In this paper, however, we examine a more subtle issue present in tool mark examinations: searches within evidence collected as part of a single case.
+We begin with a hypothetical set of three scenarios involving the collection of tool mark evidence and then assess the statistical issues involved with each scenario and explore possible resolutions to those issues.
 
@@
 
-We will consider wires ranging in diameter from 2mm (solid 12 gauge wire) to 0.0116 mm (stranded 12 gauge wire may be composed of 49 strands of 0.0116 mm diameter wire) for the purposes of this thought experiment; these values represent wire easily obtainable at a local hardware store.
+We will consider wires ranging in diameter from 2mm (solid 12 gauge wire) to 0.0116 mm (stranded 12 gauge wire may be composed of 49 strands of 0.0116 mm diameter wire) for the purposes of this thought experiment; these values represent types of wire easily obtainable at a local hardware store.
 
@@
 ```
+
 ```{r}

### 2023-11-13 Preparation/effective_obs.qmd
Update renv and start thinking about effective observations

---
title: "Time Series"
format: html
editor: visual
---

## Effective Number of Observations in a Single Series

In order to calculate the number of comparisons performed by an algorithm, we must calculate the effective number of observations. After all, if we use two different microscopes with different resolutions, we will see more detail but still have effectively the same evidence.

Laying this out mathematically, let us assume that we have a series of observations of a physical object of length $\ell$, $x_i = {x_1, …, x_n}$ taken at resolution $\ell/n$.

```{r}
# Using Hamby 44 barrel 6, Bullet 1, Land 1 as an example
library(x3ptools)
library(bulletxtrctr)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)

bullet_x3p <- read_bullet("../data/Barrel 6/Bullet 1/")

df <- bullet_x3p %>%
  mutate(x3p = x3p %>% map(x3p_m_to_mum),
         scale = map_dbl(x3p, x3p_get_scale))

```

Useful refs:

-   https://imedea.uib-csic.es/master/cambioglobal/Modulo_V_cod101615/Theory/TSA_theory_part1.pdf

-   https://andrewcharlesjones.github.io/journal/21-effective-sample-size.html

-

### 2023-11-13 Preparation/effective_obs.qmd
Updates

@@
 ---
+
+```{r data-setup, include = F}
+library(readr)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+
+wire_sigs <- read_csv("../data/data-manual/wire-signatures.csv")
+
+```
 
@@
 
+```{r wire-cut-example-data}
+wce <- wire_sigs %>% 
+  filter(id == "T1AW-LI-R1", edge == "A") 
+```
+
+For instance, consider the cut edge of 16 gauge wire (1.5 mm diameter), scanned at a resolution of 0.645 $\mu m$.
+We have `r nrow(wce)` observations spaced at $0.645 \mu m$ for a total sample length of $2144*0.645/1000 = `r sprintf("%.02f", nrow(wce)*0.645/1000)` mm$, which represents the cut edge of the 1.5mm diameter wire, as shown in @fig-wire-scan-process.
+
+![The tip of the cut wire is imaged and then processed to produce the wire signature](../data/JPG-Images/Image-18.jpg){#fig-wire-scan-process fig-align="center" style="width:430px; height:300px;object-position='50% 150px';object-fit:cover;"}
+
+```{r, echo = F, warning = F}
+#| fig-width: 8
+#| fig-height: 2
+#| dpi: 300
+
+wce %>%
+  ggplot(aes(x = x, y = sig)) + 
+  geom_line() + 
+  ggtitle("T1AW-LI-R1 A signature") + 
+  theme_bw() + 
+  ylab("Signature height") + xlab("Position")
+```
+No matter how small our resolution, the real-world surface is never more than 1.383 mm in length and the underlying striations are essentially fixed - what changes is what level of detail we can observe.
+At some point, adding additional resolution doesn't get us any more information about the striations; thus, we cannot depend solely on $n$, the number of observations, to tell us how much information is contained in our sequence.
+In order to correct for this, time series analysts use an effective sample size calculation of $N^\prime = N \cdot \frac{1-r_1}{1+r_1}$ when there is first-order autocorrelation.
+
+We can test for autocorrelation by looking at the auto correlation function plot, which tests the correlation between $x_t$ and $x_{t+i}$, where $i$ is the order of the autocorrelation function (e.g.
+the number of positions offset from the observation).
+
 ```{r}
-# Using Hamby 44 barrel 6, Bullet 1, Land 1 as an example
-library(x3ptools)
-library(bulletxtrctr)
-library(tibble)
-library(dplyr)
-library(tidyr)
-library(purrr)
+acf(na.omit(wce$sig), type = "correlation")
+myacf <- acf(na.omit(wce$sig), plot = F)
+```
 
-bullet_x3p <- read_bullet("../data/Barrel 6/Bullet 1/")
+It is obvious that there is a significant amount of autocorrelation in our striae.
 
-df <- bullet_x3p %>%
-  mutate(x3p = x3p %>% map(x3p_m_to_mum),
-         scale = map_dbl(x3p, x3p_get_scale))
+```{r}
+library(lmtest)
 
+wce_ts <- wce %>%
+  mutate(sig1 = c(NA, diff(sig, 1)),
+         sig2 = c(NA, diff(sig1, 1)))
+model <- lm(sig ~ x, data = wce_ts)
+lmtest::dwtest(model)
+```
+The null hypothesis of the Durbin-Watson test is that there is no autocorrelation; we can conclude that there is evidence of autocorrelation in our data.
+
+Our next task is to examine different orders of autocorrelation. We can use the `auto.arima` function to determine the best class of time series model structure for our data.
+
+```{r}
+library(forecast)
+
+df <- dplyr::filter(wce, !is.na(sig))
+aamodel <- auto.arima(y = df$sig, stepwise = FALSE, parallel = T)
+summary(aamodel)
+
+ggtsdisplay(df$sig, theme = ggplot2::theme_bw(), main = "Signature")
+
+ggtsdisplay(resid(aamodel), theme = ggplot2::theme_bw(), main = "Residuals from Auto.Arima model")
+```
+
+There are multiple ways to calculate "effective sample size" which seem to differ in what level of ARIMA modeling they account for and the underlying method to compute the autocorrelations.
+
+The `coda` package uses $ESS = N \frac{\lamda^2}{\sigma^2}$ where $\lamda^2$ is the variance and $\sigma^2$ is the spectral density at frequency 0.
+
+```{r}
+coda::effectiveSize(df$sig)
+# Sampling every other point
+coda::effectiveSize(df$sig[seq(1, length(df$sig), 2)])
+# Sampling every 4th point
+coda::effectiveSize(df$sig[seq(1, length(df$sig), 4)])
+```
+
+Another calculation for first-order autocorrelation (which doesn't apply here...) is
+
+```{r}
+ar1_ess <- sum(!is.na(df$sig)) * (1-myacf$acf[2])/(1 + myacf$acf[2])
+```
+
+$ESS = N \frac{1 - \rho(1)}{1 + \rho(1)}$ where $\rho(1)$ is the sample autocorrelation at lag 1.
+As we have an ARMA(3,2) model, we can't really use this with a straight face, however, if we did, the result would be `r ar1_ess`...
+e.g.
+not reasonable on its face.
+
+I'm going to quit now, but this turns out to be a real pain in the ass.
+I don't find 25 or 26 observations to be particularly realistic, but neither do I think .5 observations is realistic.
+To my mind, something like 50-100 observations would be reasonable, but I don't know that I have enough of a gut feeling for this yet.
+
+```{r, echo = F, warning = F}
+#| fig-width: 8
+#| fig-height: 2
+#| dpi: 300
+
+wce %>%
+  ggplot(aes(x = x, y = sig)) + 
+  geom_line() + 
+  ggtitle("T1AW-LI-R1 A signature") + 
+  theme_bw() + 
+  ylab("Signature height") + xlab("Position")
 ```
@@
 Useful refs:
+
+- https://stats.stackexchange.com/questions/429470/what-is-the-correct-effective-sample-size-ess-calculation
 

### 2023-11-13 LPR/index.qmd
Update to intro

@@
 One of the common suggestions in the 2016 PCAST report across pattern disciplines in forensic science was the importance of creating huge databases of e.g.
-known prints and firearms to support research into matches and closest non-matches\[@pcast, fingerprints pg 11, 88, firearms pg 12\].
+known prints and firearms to support research into matches and closest non-matches[@pcast, fingerprints pg 11, 88, firearms pg 12].
 As researchers, we concur with this suggestion, as it is critical for developing objective algorithms and creating useful black-box studies.
-We have spent considerable time assembling large data sets of tool marks, bullets, and cartridge cases to support our research as well as create public resources to accelerate future research \[@csafeDataPortal\].
-However, we are wary about the implications of database searches across forensic disciplines, for reasons also mentioned in the PCAST report: the necessity of understanding the random match probability for searches against these databases, in particular when using degraded evidence often found in crime scenes\[@pcast, pg 52\].
+We have spent considerable time assembling large data sets of tool marks, bullets, and cartridge cases to support our research as well as create public resources to accelerate future research [@csafeDataPortal].
+However, we are wary about the implications of database searches across forensic disciplines, for reasons also mentioned in the PCAST report: the necessity of understanding the random match probability for searches against these databases, in particular when using degraded evidence often found in crime scenes[@pcast, pg 52].
 

### 2023-11-13 Preparation/effective_obs.qmd
add in some Heike-suggested diagnostics

@@
 
--
+## Leaving out the last X%
+
+```{r}
+library(forecast)
+
+df <- dplyr::filter(wce, !is.na(sig))
+
+df[1:1500,]
+aamodel <- auto.arima(y = df$sig[1:1500], xreg = df$x[1:1500], stepwise = FALSE, parallel = T)
+summary(aamodel)
+
+pred <- predict(aamodel, n.ahead = 500, newxreg = df$x[1501:2000]) %>%
+  as.data.frame() %>%
+  mutate(x = df$x[1501:2000], ypred = pred, ylow = ypred - 2*se, yhigh = ypred + 2*se)
+
+ggplot(df, aes(x = x, y = sig)) + geom_line() + 
+  geom_line(data = pred, aes(x = x, y = ypred), color = "red") + 
+  geom_ribbon(data = pred, aes(x = x, y = ypred, ymin = ylow, ymax = yhigh), fill = "red", alpha = .2)
+```

### 2023-11-17 LPR/index.qmd
Updates

@@
 library(dplyr)
+library(tidyr)
+library(purrr)
 
@@
   "Pliers", 1.5, 2, 10,
-  "Needlenose pliers", 1, 4, 3,
+  "Pliers", 1, 4, 3,
   "Side cutters", 1, 4, 1,
   "Side cutters", 2, 4, 2,
-  "Scissors", 8.89, 2, 14,
-  "Tin snips", 3.81, 2, 4,
+  "Scissors", 8, 2, 14,
+  "Tin snips", 4, 2, 4,
   "Lineman pliers", 2, 4, 2,
@@
   "Long handled cutting pliers", 3, 4, 1,
-  "Utility knife blades", 7.62, 2, 50
+  "Utility knife blades", 6.2, 2, 50
 ) %>%
@@
 total_cutting_surface = sum(cutting_surfaces$total_cutting_length)
+
+dad_surfaces <- tribble(
+  ~type, ~length, ~surfaces, ~tools,
+  "Pliers", 1.8, 2, 1,
+  "Pliers", 1.6, 2, 1,
+  "Pliers", 1.1, 2, 1,
+  "Wire cutters", 2, 2, 1,
+  "Wire cutters", 1.5, 2, 1,
+  "Wire cutters", 1, 2, 2,
+  "Other cutters", 8, 2, 1,
+  "Other cutters", 4, 4, 1,
+  "Scissors", 7.5, 2, 5,
+  "Utility knife blades", 6.2, 2, 160,
+  "Razor blades", 4, 2, 14,
+  "Knife", 4, 2, 1,
+  "Knife", 2, 2, 1,
+  "Clippers", 4, 2, 1
+) %>%
+  mutate(total_cutting_length = length*surfaces*tools)
+
+dad_total_cutting_surface = sum(dad_surfaces$total_cutting_length)
 ```
@@
 
-A scan of a cut made by the blade at $0.625\mu m$ resolution would produce approximately $15/0.0000625 = 240000$ observations; a scan of a 2mm wire would produce $2/0.0000625 = 32000$ observations, and computationally aligning these two data sequences using maximum cross-correlation would require $240000 - 32000 + 1 = 208001$ comparisons.
-Under an optical microscope, an examiner might move the wire at 0.2mm increments along the blade cut surface for a total of $15/.2 - 2/.2 + 1 = 66$ distinct and discrete comparisons.
-Then, we must consider that the examiner or algorithm must compare the wire to both side A and side B; we then double the number of total comparisons made to $416002$ and $132$ for the algorithm and the examiner, respectively.
+```{r}
+resolution <- 0.645 # res in mum
+length_blade <- 15 # length in mm
+wire_diam <- 2 # diameter in mm
+mm_to_mum <- 1000 # conversion
+
+obs_blade <- round(length_blade * mm_to_mum/resolution)
+obs_wire <- round(wire_diam * mm_to_mum/resolution)
+alg_compare <- obs_blade - obs_wire + 1
+
+examiner_increment <- 0.2 # increment in mm
+examiner_compare <- (length_blade - wire_diam)/examiner_increment + 1
+```
+
+A scan of a cut made by the blade at $`r resolution`\mu m$ resolution will produce approximately $`r length_blade` \times `r mm_to_mum`/`r resolution` = `r format(obs_blade, big.mark = ',')`$ observations; a scan of a 2mm wire would produce $`r wire_diam` \times `r mm_to_mum`/`r resolution` = `r format(obs_wire, big.mark = ',')`$ observations, and computationally aligning these two data sequences using maximum cross-correlation would require $`r obs_blade` - `r obs_wire` + 1 = `r format(alg_compare, big.mark = ',')`$ (obviously dependent) comparisons.
+Under an optical microscope, an examiner might move the wire at 0.2mm increments along the blade cut surface for a total of $`r length_blade`/`r examiner_increment` - `r wire_diam`/`r examiner_increment` + 1 = `r format(examiner_compare, big.mark = ',')`$ distinct comparisons.
+Then, we must consider that the examiner or algorithm must compare the wire to both side A and side B; we then double the number of total comparisons made to $`r format(alg_compare*2, big.mark = ',')`$ and $`r examiner_compare*2`$ for the algorithm and the examiner, respectively.
 
@@
 
-While 132 visual comparisons might not seem so terrible, this is the simplest of the 3 scenarios we consider; each scenario presents a progressively more complex problem.
+While $`r examiner_compare*2`$ visual comparisons might not seem so terrible, this is the simplest of the 3 scenarios we consider; each scenario presents a progressively more complex problem.
 
@@
 Let us consider the garage of one of the authors of this paper; @fig-shop-layout shows the garage work bench and tool storage in typical condition.
-While there is no reason to think the author's garage is representative of all garages, the general setup and number of tools in the garage does not seem unusual for a suburban household where residents may dabble in automotive repair, home improvement, or woodworking.
+While there is no reason to think the author's garage is representative of all garages, the general setup and number of tools in the garage is not unusual for a suburban household where residents may dabble in automotive repair, home improvement, or woodworking.
 
@@
 #| message: false
-knitr::kable(cutting_surfaces)
+knitr::kable(cutting_surfaces %>% rename(Type = type, `Length (cm)` = length, `Cutting Surfaces` = surfaces, `# Tools` = tools, `Total Cutting Length (cm)` = total_cutting_length))
 ```
@@
 comparisons <- cutting_surfaces %>%
-  mutate(digital = (length*10/0.0000625 - 2/0.0000625+1)*surfaces*tools,
-         analog = (length*10/.2 - 2/.2 + 1)) %>%
+  mutate(digital = (length*10*mm_to_mum/resolution - wire_diam*mm_to_mum/resolution + 1)*surfaces*tools,
+         analog = (length*10/examiner_increment - wire_diam/examiner_increment + 1)) %>%
   group_by(type) %>%
-  summarize(digital = sum(digital), analog = sum(analog)) %>%
+  summarize(digital = sum(round(digital)), analog = sum(round(analog))) %>%
   rename(`Tool type` = type, Algorithm = digital, Examiner = analog)
 
-knitr::kable(comparisons)
+compare_tbl <- bind_rows(comparisons, tibble(`Tool type` = "Total", Algorithm = sum(comparisons$Algorithm), Examiner = sum(comparisons$Examiner)))
+knitr::kable(compare_tbl, format.args = list(big.mark=','))
 ```
@@
 The total length of all cutting surfaces of relevant tools in the author's garage is `r total_cutting_surface` cm; primary contributions to the total length are the 50 pack of utility knife blades and the 14 pairs of full-size scissors which could be located throughout the house.
-Of course, a thorough search of the garage and house would come up with far more cutting surfaces to compare against during an actual crime investigation.
+Of course, a thorough search of the garage and house, as one would expect from professionals investigating an actual crime, would yield far more cutting surfaces to compare against.
 
-[^2]: Reasonably is defined as we have used a tool like this to cut wires during electrical work in the past.
+[^2]: Reasonably is defined as we (the authors) have used a tool like this to cut wires while doing home improvement projects in the past.
 
 As in Scenario 1, the wire fragment recovered from the un-detonated explosive would be compared to blade surface test impressions generated from each cutting surface in the garage.
-However, unlike in Scenario 1, we have `r total_cutting_surface` cm of cutting surfaces to work with.
+However, unlike in Scenario 1, we have `r format(total_cutting_surface, big.mark=',')` cm of cutting surfaces to work with.
 We can apply our comparison formula on a by-surface basis to yield the digital and analog comparison estimates shown in @tbl-shop-comparisons.
-Overall, our algorithm would make `r sum(comparisons$Algorithm)` comparisons and an examiner doing a visual inspection would make `r sum(comparisons$Examiner)` comparisons under the assumptions we made in Scenario 1.
-XXX fill this in with proper formatting XXX
+Overall, our algorithm would make `r format(sum(comparisons$Algorithm), big.mark=',')` (dependent) comparisons and an examiner doing a visual inspection would make `r format(sum(comparisons$Examiner), big.mark=',')` comparisons under the assumptions we made in Scenario 1.
 
 ### Multiple Suspects
+
+To fully understand the magnitude of this problem, suppose that we have a group of 4 additional individuals who may have contributed to building the un-detonated bomb.
+These individuals are each associated with the author and have hobbies that include building electronic devices, home improvement, welding, and woodworking (thus, they each have sets of tools).
+As part of the investigation, each individuals' tools are confiscated and examined; blade cuts are made of each cutting surface to be matched to the wire in evidence.
+@tbl-garage-comparisons shows the number of comparisons which would need to be performed to ensure that all wire-cutting tools in each suspect's garage are compared to the evidence from the scene.
+
+```{r}
+#| label: tbl-garage-comparisons
+#| tbl-cap: "Total cutting length of easily accessible wire-cutting tools for 4 individuals associated with the author."
+#| echo: false
+#| message: false
+set.seed(42309427)
+
+n_tools <- c(40, 20, 30)
+n_blades <- sample(5:75, size = 3)
+
+tool_options <- select(bind_rows(cutting_surfaces, dad_surfaces), 1:3) %>% unique()
+
+blades <- filter(tool_options, type %in% c("Utility knife blades", "Razor blades"))
+
+other_tools <- filter(tool_options, !type %in% c("Utility knife blades", "Razor blades"))
+
+dad_tools <- dad_surfaces %>% 
+  mutate(suspect = 4) %>% select(-total_cutting_length)
+
+our_tools <- cutting_surfaces %>% mutate(suspect = 0) %>% select(-total_cutting_length)
+
+sim_garages <- tibble(suspect = 1:3, n_tools = n_tools, n_blades = n_blades) %>%
+  mutate(data = map2(
+    n_tools, n_blades, 
+    ~bind_rows(
+      slice_sample(other_tools, n = .x, replace = T), 
+      slice_sample(blades, n = .y, replace = T)) 
+  )) %>%
+  unnest(data) %>%
+  group_by(suspect, type, length, surfaces) %>% 
+  summarize(tools = n()) %>%
+  ungroup()
+
+garages <- bind_rows(sim_garages, dad_tools, our_tools) %>%
+  group_by(suspect) %>%
+   mutate(digital = (length*10*mm_to_mum/resolution - wire_diam*mm_to_mum/resolution + 1)*surfaces*tools,
+         analog = (length*10/examiner_increment - wire_diam/examiner_increment + 1)*surfaces*tools)
+
+garage_sum <- garages %>%
+  summarize(
+    total_cutting_length = sum(length * surfaces * tools), 
+    digital = sum(digital), 
+    analog = sum(analog)) %>%
+  mutate(suspect = as.character(suspect))
+
+garage_sum_tot <- bind_rows(garage_sum, tibble(suspect = "Total", total_cutting_length = sum(garage_sum$total_cutting_length), digital = sum(garage_sum$digital), analog = sum(garage_sum$analog)))
+
+garage_sum_tot %>%
+  rename(Individual = suspect, `Total Cutting Length (cm)` = total_cutting_length, "Algorithm" = digital, "Examiner" = analog) %>%
+  knitr::kable(format.args = list(big.mark = ','))
+```
+
+If investigators were to attempt to determine which of the suspected individuals cut the wire in evidence, they would need to do `r format(sum(garage_sum$analog), big.mark = ',')` distinct visual comparisons.
+Even if there is an incredibly small false positive error rate, the odds of a false identification are much more likely in this scenario than either of the previous two scenarios.
+But how likely?
+In the next section, we describe two different scenarios for calculating the overall likelihood of a false identification using simple probability calculations and estimated coincidental match probabilities as well as a theoretical approach designed to mimic the approach used with database searches that return the $N$ most similar results.
 

### 2023-11-17 Preparation/effective_obs.qmd
Updates

@@
 ```
+
 No matter how small our resolution, the real-world surface is never more than 1.383 mm in length and the underlying striations are essentially fixed - what changes is what level of detail we can observe.
@@
 ```
+
 The null hypothesis of the Durbin-Watson test is that there is no autocorrelation; we can conclude that there is evidence of autocorrelation in our data.

### 2023-11-20 LPR/index.qmd
Add extra sections of the intro

@@
 We begin with a hypothetical set of three scenarios involving the collection of tool mark evidence and then assess the statistical issues involved with each scenario and explore possible resolutions to those issues.
+
+### Examination Process
+
+### Comparison counts
+
+When assessing the total number of comparisons, we will consider two extremes: first, a set of completely independent comparisons made by dividing the blade cut up into a series of non-overlapping segments the same size as the wire.
+XXX add picture XXX The second extreme is a series of comparisons made by sliding the wire along the blade cut in increments equal to the resolution of the scan or optical viewer (depending on whether this comparison is done automatically or visually).
+Of course, the reality is that an algorithm or examiner will likely make a number of comparisons between these two extremes, however, rather than consider all possible comparison numbers within this span, it is more efficient to explore what happens at each endpoint first.
 

### 2023-11-20 LPR/index.qmd
single cut

--- 2023-11-17 LPR/index.qmd
@@
 
-### Single Tool Comparisons
+### Single Tool, Single Cut Comparisons
 
 Let us first consider the situation where a single tool is recovered from the suspect's house that is viewed as likely to have made the wire cut.
-Then, a forensic examiner might take the tool and use it to cut a piece of metal, known as a blade, in order to record the markings of the entire cutting surface simultaneously.
+Then, a forensic examiner might take the tool and use it to cut a piece of metal, known as a blade cut, in order to record the markings of the entire cutting surface simultaneously.
 This process might be repeated at multiple angles in order to record the effect of different contact surfaces on the resulting striations.
-For simplicity, let us assume that angles are not a factor.
+For simplicity, let us for now assume that only a single cut under one set angle was made.
 Let us further define the blade to be of length $\ell$ and the wire to have diameter $d$ which is fully covered with striations that are untouched by the blast.

### 2023-11-20 LPR/index.qmd
Make comparison counts more general

@@
 ### Examination Process
+
+Let us consider the process of toolmark examination - an examiner is given a tool recovered from a suspect as well as a wire recovered from the scene.
+The examiner will typically take the tool and use it to cut a piece of metal, known as a blade cut, in order to record the markings of the entire cutting surface simultaneously.
+This process may be repeated at multiple angles in order to record the effect of different contact surfaces on the resulting striations.
+The resulting blade cut will be compared to the wire using either visual inspection or perhaps in the future a numerical comparison algorithm similar to those developed for bullet striations [@hare2017automatic] and cartridge cases [@zemmels2023].
+
+The examiner will compare blade cuts from all cutting surfaces during this process.
+Tools used for wire cuts typically have either 2 or 4 cutting surfaces.
+Examiners may make multiple cuts at multiple angles; this will increase the number of available blade cuts for comparison.
+
+For simplicity, let us for now assume that only a single cut under one controlled angle was made.
+
+### Counting Comparisons
+
+Let us consider a single blade cut from a single cutting surface.
+Define the blade cut to be of length $\ell$ and the wire to have diameter $d$; we will also assume that the wire is fully covered with striations suitable for comparison.
+
+At minimum, there must be $\ell/d$ comparisons in order to assess whether the tool matches the wire; in this case, the comparisons are essentially independent of each other (conditional on the striation patterns) because no overlapping data is used.
+Of course, in this situation, we are overlooking the potential for partial matches and assuming that everything is lined up perfectly to produce this minimum number of comparisons; in practice, of course, the number of comparisons in this situation would likely be much higher.
+
+At a maximum, both the blade and the wire are either digitally scanned at the same resolution of $r$ mm, or visually examined using a microscope with a digital view of a certain resolution which can be translated into an equivalent resolution to the digital scan.
+In this case, there may be as many as $\ell/r- d/r + 1$ comparisons performed in order to find the optimal alignment between the two samples[^1].
+Of course, these comparisons are highly dependent, as each sequential comparison shares much of the same data with the previous and next comparisons.
+While it is likely possible to model this autocorrelation and compute an effective number of comparisons accounting for the autocorrelation, developing this model is outside the scope of this paper.
+
+[^1]: This assumes we restrict the wire to be completely contained within the bounds of the blade surface.
+If we allow for partial matches and restrict these matches to require at least $k$ points of overlap, then this becomes $\ell/r - d/r + k + 1$.
+
+Thus, we will consider the two extremes: $\ell/d$ completely independent comparisons and $\ell/r - d/r + 1$ dependent comparisons.
+The true number of visual or numerical comparisons lies somewhere between these two values.
+
+To make this more concrete, consider the pair of pliers shown in @fig-wire-cutter, which has a 1.5 cm cutting surface machined on both sides to produce a peaked cross-section.
+When used to cut a wire, the wire rests against a rectangular surface used to hold it in place; the blade is pushed into the wire, producing striae resulting from imperfections in side A on one half of the wire and striae from imperfections and wear in side B on the other.
+
+![Wire cutter with 1.5 cm cutting surface beveled on both sides](../fig/wire_cutter.png){#fig-wire-cutter}
+
+```{r}
+resolution <- 0.645 # res in mum
+length_blade <- 15 # length in mm
+wire_diam <- 2 # diameter in mm
+mm_to_mum <- 1000 # conversion
+
+obs_blade <- round(length_blade * mm_to_mum/resolution)
+obs_wire <- round(wire_diam * mm_to_mum/resolution)
+alg_compare <- obs_blade - obs_wire + 1
+
+examiner_increment <- 0.2 # increment in mm
+examiner_compare <- (length_blade - wire_diam)/examiner_increment + 1
+```
+
+A scan of a cut made by the blade at $`r resolution`\mu m$ resolution will produce approximately $`r length_blade` \times `r mm_to_mum`/`r resolution` = `r format(obs_blade, big.mark = ',')`$ observations; a scan of a 2mm wire would produce $`r wire_diam` \times `r mm_to_mum`/`r resolution` = `r format(obs_wire, big.mark = ',')`$ observations, and computationally aligning these two data sequences using maximum cross-correlation would require $`r obs_blade` - `r obs_wire` + 1 = `r format(alg_compare, big.mark = ',')`$ (obviously dependent) comparisons.
+Under an optical microscope, an examiner might move the wire at 0.2mm increments along the blade cut surface for a total of $`r length_blade`/`r examiner_increment` - `r wire_diam`/`r examiner_increment` + 1 = `r format(examiner_compare, big.mark = ',')`$ distinct comparisons.
+Then, we must consider that the examiner or algorithm must compare the wire to both side A and side B; we then double the number of total comparisons made to $`r format(alg_compare*2, big.mark = ',')`$ and $`r examiner_compare*2`$ for the algorithm and the examiner, respectively.
 
@@
 Let us first consider the situation where a single tool is recovered from the suspect's house that is viewed as likely to have made the wire cut.
-Then, a forensic examiner might take the tool and use it to cut a piece of metal, known as a blade cut, in order to record the markings of the entire cutting surface simultaneously.
-This process might be repeated at multiple angles in order to record the effect of different contact surfaces on the resulting striations.
-For simplicity, let us for now assume that only a single cut under one set angle was made.
-Let us further define the blade to be of length $\ell$ and the wire to have diameter $d$ which is fully covered with striations that are untouched by the blast.
-
-At minimum, the examiner will have to do $\ell/d$ comparisons in order to assess whether the tool matches the wire; when a comparison algorithm is used, and both the blade and the wire are scanned at the same resolution of $r$ mm, there may be as many as $\ell/r- d/r + 1$ comparisons performed automatically in order to find the optimal alignment between the two samples[^1].
-
-[^1]: This assumes we restrict the wire to be completely contained within the bounds of the blade surface.
-If we allow for partial matches and restrict these matches to require at least $k$ points of overlap, then this becomes $\ell/r - d/r + k + 1$.
-
-To make this more concrete, consider the pair of pliers shown in @fig-wire-cutter, which has a 1.5 cm cutting surface machined on both sides to produce a peaked cross-section.
-When used to cut a wire, the wire rests against a rectangular surface used to hold it in place; the blade is pushed into the wire, producing striae resulting from imperfections in side A on one half of the wire and striae from imperfections and wear in side B on the other.
-
-![Wire cutter with 1.5 cm cutting surface beveled on both sides](../fig/wire_cutter.png){#fig-wire-cutter}
-
-```{r}
-resolution <- 0.645 # res in mum
-length_blade <- 15 # length in mm
-wire_diam <- 2 # diameter in mm
-mm_to_mum <- 1000 # conversion
-
-obs_blade <- round(length_blade * mm_to_mum/resolution)
-obs_wire <- round(wire_diam * mm_to_mum/resolution)
-alg_compare <- obs_blade - obs_wire + 1
-
-examiner_increment <- 0.2 # increment in mm
-examiner_compare <- (length_blade - wire_diam)/examiner_increment + 1
-```
-
-A scan of a cut made by the blade at $`r resolution`\mu m$ resolution will produce approximately $`r length_blade` \times `r mm_to_mum`/`r resolution` = `r format(obs_blade, big.mark = ',')`$ observations; a scan of a 2mm wire would produce $`r wire_diam` \times `r mm_to_mum`/`r resolution` = `r format(obs_wire, big.mark = ',')`$ observations, and computationally aligning these two data sequences using maximum cross-correlation would require $`r obs_blade` - `r obs_wire` + 1 = `r format(alg_compare, big.mark = ',')`$ (obviously dependent) comparisons.
-Under an optical microscope, an examiner might move the wire at 0.2mm increments along the blade cut surface for a total of $`r length_blade`/`r examiner_increment` - `r wire_diam`/`r examiner_increment` + 1 = `r format(examiner_compare, big.mark = ',')`$ distinct comparisons.
-Then, we must consider that the examiner or algorithm must compare the wire to both side A and side B; we then double the number of total comparisons made to $`r format(alg_compare*2, big.mark = ',')`$ and $`r examiner_compare*2`$ for the algorithm and the examiner, respectively.
-
 Ultimately, in order to compare tool marks in this scenario, the examiner or algorithm completes the equivalent to a database search resulting from a single piece of evidence.

### 2023-11-20 LPR/index.qmd
Pictures and more text

@@
 
-For simplicity, let us for now assume that only a single cut under one controlled angle was made.
+<!-- For simplicity, let us for now assume that only a single cut under one controlled angle was made. -->
+
+<!-- ### Blade Configurations -->
+
+<!-- There are 3 major types of blade configurations we observed -->
 
@@
 The true number of visual or numerical comparisons lies somewhere between these two values.
-
-To make this more concrete, consider the pair of pliers shown in @fig-wire-cutter, which has a 1.5 cm cutting surface machined on both sides to produce a peaked cross-section.
-When used to cut a wire, the wire rests against a rectangular surface used to hold it in place; the blade is pushed into the wire, producing striae resulting from imperfections in side A on one half of the wire and striae from imperfections and wear in side B on the other.
-
-![Wire cutter with 1.5 cm cutting surface beveled on both sides](../fig/wire_cutter.png){#fig-wire-cutter}
-
-```{r}
-resolution <- 0.645 # res in mum
-length_blade <- 15 # length in mm
-wire_diam <- 2 # diameter in mm
-mm_to_mum <- 1000 # conversion
-
-obs_blade <- round(length_blade * mm_to_mum/resolution)
-obs_wire <- round(wire_diam * mm_to_mum/resolution)
-alg_compare <- obs_blade - obs_wire + 1
-
-examiner_increment <- 0.2 # increment in mm
-examiner_compare <- (length_blade - wire_diam)/examiner_increment + 1
-```
-
-A scan of a cut made by the blade at $`r resolution`\mu m$ resolution will produce approximately $`r length_blade` \times `r mm_to_mum`/`r resolution` = `r format(obs_blade, big.mark = ',')`$ observations; a scan of a 2mm wire would produce $`r wire_diam` \times `r mm_to_mum`/`r resolution` = `r format(obs_wire, big.mark = ',')`$ observations, and computationally aligning these two data sequences using maximum cross-correlation would require $`r obs_blade` - `r obs_wire` + 1 = `r format(alg_compare, big.mark = ',')`$ (obviously dependent) comparisons.
-Under an optical microscope, an examiner might move the wire at 0.2mm increments along the blade cut surface for a total of $`r length_blade`/`r examiner_increment` - `r wire_diam`/`r examiner_increment` + 1 = `r format(examiner_compare, big.mark = ',')`$ distinct comparisons.
-Then, we must consider that the examiner or algorithm must compare the wire to both side A and side B; we then double the number of total comparisons made to $`r format(alg_compare*2, big.mark = ',')`$ and $`r examiner_compare*2`$ for the algorithm and the examiner, respectively.
-
-### Comparison counts
-
-When assessing the total number of comparisons, we will consider two extremes: first, a set of completely independent comparisons made by dividing the blade cut up into a series of non-overlapping segments the same size as the wire.
-XXX add picture XXX The second extreme is a series of comparisons made by sliding the wire along the blade cut in increments equal to the resolution of the scan or optical viewer (depending on whether this comparison is done automatically or visually).
-Of course, the reality is that an algorithm or examiner will likely make a number of comparisons between these two extremes, however, rather than consider all possible comparison numbers within this span, it is more efficient to explore what happens at each endpoint first.
 
@@
 
-Consider, for example a situation where someone builds a bomb, and a wire fragment is recovered from un-detonated explosive.
+Consider, for example a situation where someone builds a bomb (badly), and a wire fragment is recovered from the un-detonated explosive device.
 We will examine how the forensic examination might proceed in three different cases:
@@
 Examiners identify a suspect and locate a garage full of tools, one of which may have been used to construct the bomb.
-Examiners would like to identify which tool(s) were used to cut the wires during bomb construction.
+Examiners would like to identify which tool(s) were used to cut the wire during bomb construction.
 
 3.  Examiners identify several suspect, each of whom has a set of tools, and wish to determine which suspect(s) may have constructed the bomb based off of the toolmark comparisons.
-
-We will consider wires ranging in diameter from 2mm (solid 12 gauge wire) to 0.0116 mm (stranded 12 gauge wire may be composed of 49 strands of 0.0116 mm diameter wire) for the purposes of this thought experiment; these values represent types of wire easily obtainable at a local hardware store.
 
@@
 
-Let us first consider the situation where a single tool is recovered from the suspect's house that is viewed as likely to have made the wire cut.
-Ultimately, in order to compare tool marks in this scenario, the examiner or algorithm completes the equivalent to a database search resulting from a single piece of evidence.
-The dangers of such searches are well known and have been cited in forensic missteps such as the Madrid bombing case, as discussed in @li2023; @newman2007; @mayfield2006report [Pg 137].
+To make the calculations of number of comparisons more concrete, consider the pair of pliers shown in @fig-wire-cutter, which has a 1.5 cm cutting surface machined on both sides to produce a peaked cross-section.
+When used to cut a wire, the wire rests against a rectangular surface used to hold it in place; the blade is pushed into the wire, producing striae resulting from imperfections in side A on one half of the wire and striae from imperfections and wear in side B on the other.
 
-While $`r examiner_compare*2`$ visual comparisons might not seem so terrible, this is the simplest of the 3 scenarios we consider; each scenario presents a progressively more complex problem.
+![Wire cutter with 1.5 cm cutting surface beveled on both sides](../fig/wire_cutter.png){#fig-wire-cutter alt="Wire cutter with 1.5 cm cutting surface beveled on both sides"}
+
+```{r}
+resolution <- 0.645 # res in mum
+length_blade <- 15 # length in mm
+wire_diam <- 2 # diameter in mm
+mm_to_mum <- 1000 # conversion
+
+obs_blade <- round(length_blade * mm_to_mum/resolution)
+obs_wire <- round(wire_diam * mm_to_mum/resolution)
+alg_compare <- obs_blade - obs_wire + 1
+
+examiner_increment <- wire_diam # increment in mm
+examiner_compare <- (length_blade - wire_diam)/examiner_increment + 1
+```
+
+A scan of a cut made by the blade at $`r resolution`\mu m$ resolution will produce a sequence of approximately $`r length_blade` \times `r mm_to_mum`/`r resolution` = `r format(obs_blade, big.mark = ',')`$ observations.
+A scan of a 2 mm wire would produce $`r wire_diam` \times `r mm_to_mum`/`r resolution` = `r format(obs_wire, big.mark = ',')`$ observations.
+
+In the minimal comparisons case, we would have `r examiner_compare` comparisons for each side of the blade's cutting surface for a total of `r 2*examiner_compare` comparisons.
+
+In the extreme case, computationally aligning the wire sequence to one blade cut using maximum cross-correlation would require $`r obs_blade` - `r obs_wire` + 1 = `r format(alg_compare, big.mark = ',')`$ (obviously dependent) comparisons.
+As there are two or more blade cuts for each cutting tool, there are at most $2(`r obs_blade` - `r obs_wire` + 1) = `r format(2*alg_compare, big.mark = ',')`$ comparisons including the full wire surface.
+
+While $`r examiner_compare*2`$ comparisons might not seem so terrible, this is the simplest of the 3 scenarios we consider; each scenario presents a progressively more complex problem.
 
@@
 
+Let us first consider the simplest situation where a single tool is recovered from the suspect's house that is viewed as likely to have made the wire cut. 
+In order to compare tool marks in this scenario, the examiner or algorithm completes the equivalent to a database search resulting from comparing a single wire in evidence to a single tool. 
+This multiple comparison problem becomes more severe at each step of our thought experiment - considering a single shop of tools adds multiple orders of magnitude to the number of comparisons.
+
+In addition, we might expect that if the guilty party in our thought experiment had succeeded in exploding the bomb, the evidence might be reduced in quality - the proportion of the wire with usable striae might be significantly smaller, which would compound the problem even further.
+
+The dangers of database searches in forensics are well known; database searches have been implicated in forensic missteps such as the Madrid bombing case, as discussed in @li2023; @newman2007; @mayfield2006report [Pg 137].
+In this section, we illustrate the problems with database searches in two different ways: first, considering the implications for random match probabilities under multiple comparisons, and second, considering the impact of "best $N$ matches" search queries on the distribution of match scores.
+These issues are separate, but related, and represent different ways searches and databases (formal or informal) might be used in practice.
+While law enforcement sanctioned wire or toolmark databases used for forensic intelligence do not currently exist to the best of our knowledge, similar systems exist for bullets and cartridge cases (IBIS), fingerprints (AFIS), and footwear (NFS, a UK database).
+It seems reasonable to consider both the informal databases created from an extensive investigation and the formal databases which might be created in the future.
+
 ### Random Match Probability

### 2023-11-21 LPR/index.qmd
Updates

@@
 
-Let us consider the process of toolmark examination - an examiner is given a tool recovered from a suspect as well as a wire recovered from the scene.
+When an examiner conducts a toolmark assessment of wire cut evidence, they are typically working with a tool recovered from the suspect as well as a wire recovered from the scene.
 The examiner will typically take the tool and use it to cut a piece of metal, known as a blade cut, in order to record the markings of the entire cutting surface simultaneously.
@@
 
-![Wire cutter with 1.5 cm cutting surface beveled on both sides](../fig/wire_cutter.png){#fig-wire-cutter alt="Wire cutter with 1.5 cm cutting surface beveled on both sides"}
+![Wire cutter with 1.5 cm cutting surface beveled on both sides](../fig/wire_cutter.png){#fig-wire-cutter}
 
@@
 
-### Random Match Probability
+### Probability of a False Identification
+
+Let us briefly consider a non-exhaustive list of ways an examiner or algorithm might arrive at a false identification:
+
+-   class characteristics match + decider doesn't recognize class characteristics as such
+
+-   coincidental similarity between two items + decider doesn't see/recognize dissimilarities
+
+-   the items randomly match on identifying characteristics
+
+All of these situations may occur during white box studies (though the final situation is very improbable).
+As a result, perhaps the best estimate of the probability of a false identification that we can obtain is the false identification rate of one or more white box studies.
+We have previously published a paper detailing the error rates of different firearms and toolmark studies; to our knowledge there are no similar white box studies for wire cuts.
+However, if we confine ourselves to error rates for striated materials (bullets, aperture shear, ejectors, screwdrivers) we may be able to get a reasonable estimate for the error rate of wire cuts.
+We assembled 3 black-box error rate studies of striated toolmarks (bullets and firing pins) [@bajic2020; @mattijssen2021; @best2022]; these studies estimate the false positive rate (e.g.
+the number of different-source identifications relative to the total number of different-source comparisons) rate to be between 0.07235 [@mattijssen2021, firing pins only] and 0.004545 [@best2022].
+While there are a number of problems with these studies (which we have detailed in previous papers, i.e.
+@inconclusives), our goal here is to use the best estimates we can get of false positive error rates in particular, as the false positive rate is the main driver of the problem with multiple comparisons.
+If we calculate a pooled false positive rate from all studies, we arrive at a false positive rate of $\frac{20 + 56 + 1}{2842 + 774 + 220} = `r sprintf('%0.02f', 77/(2842+774+220))`$, that is, about 2%.
+
+```{r}
+comparisons <- examiner_compare*2
+fpr = (20+56+1)/(2842+774+220) 
+fprformat = sprintf("%0.04f", fpr)
+```
+
+If we start out by considering the first scenario (a single tool with a 1.5 cm blade compared to a single 2mm wire) and completely independent non-overlapping comparisons (7.5 comparisons) with only 2 cutting surfaces, we would effectively have 15 independent comparisons.
+Using this information, we can calculate the probability of at least one false positive error to be `r sprintf("%0.04f%%", 1-(1-fpr)^comparisons * 100)` as shown in Equation \ref{eq:fpr}.
+While far more comparisons are likely to be performed, these comparisons may have some dependencies (at which point our formula would need to account for the covariance between successive comparisons).
+Ultimately, the number of comparisons cannot drop below `r comparisons` without leaving some evidence unexamined, which means this overall false positive rate is the lower bound.
+
+This is the simplest situation of the three different scenarios we have considered, and we still have an unacceptably high rate of at least one false positive conclusion.
+If the wire size decreases or the blade length increases, the overall chance of at least one false positive error will increase because the number of comparisons increases.
+If examiners consider cuts made at multiple angles, the same conclusion holds - the chance of at least one false positive error will increase from this point.
+
+\begin{align*}
+P(\text{False positive error in one comparison}) &= `r fprformat`\\
+P(\text{No false positive error in one comparison}) &= `r sprintf('%0.04f', 1-fpr)`\\
+P(\text{no false positive errors in } N \text{ comparisons}) &= `r sprintf('%0.04f', 1-fpr)`^N\\
+P(\text{at least one false positive error in } N \text{ comparisons}) &= 1 - `r sprintf('%0.04f', 1-fpr)`^N  \tag{A}\label{eq:fpr}\\
+&= `r sprintf("%.04f", 1-(1-fpr)^comparisons)` %>% 
+\end{align*}
+
+It then becomes relevant to consider how low the error rate would need to be for striated comparisons in order to make wire cut examination plausible.
+If we control the overall error rate amid $N$ independent comparisons, our estimated FPR for a single comparison can be calculated by solving Equation \ref{eq:fpr-control} for $r$.
+
+Let $F_n:=\{\text{at least one false positive in }n\text{ comparisons}\}$
+
+\begin{align*}
+P(F_1) &= r\\
+P(\overline{F_1}) &= 1-r\\
+P(F_n) &= \left[1-r\right]^n\\
+P(\overline{F_n}) &= 1 - \left[P(\overline{F_1})\right]^n = 1 - \left[1 - r\right]^n\tag{B}\label{eq:fpr-control}\\
+r &= 1 - \left[1 - P(F_n)\right]^{1/n}
+\end{align*}
+
+We can explore what this means visually by considering different values of $N$ and $r$ to see how the overall false positive error rate behaves, as shown in @fig-control-overall-fpr.
+The astonishing conclusion is that even at the smallest error rate for striated comparisons, 0.45% [@bajic2020], we have about a 3% chance of at least one false positive conclusion when 10 comparisons are made and about a 37.5% chance of at least one false positive conclusion when 100 comparisons are made!
+These results suggest that under current estimates of false positive error rates, it is nearly impossible for toolmark comparisons of wires to provide compelling, "beyond a reasonable doubt" levels of evidence.
+
+```{r control-overall-fpr}
+#| echo: false
+#| message: false
+#| fig-cap: Exploration of overall error rate estimates for different values of $r$, the false positive rate for a single comparison.
+Dotted vertical lines show false positive error rate estimates (from left to right, @bajic2020; @best2022; @mattijssen2021).
+#| label: fig-control-overall-fpr
+library(ggplot2)
+library(tidyr)
+single_rmp = exp(seq(-1000, 0, .1))
+N = c(10, 100, 1000, 10000)
+
+res = crossing(single_rmp, N) %>%
+  mutate(overall = 1 - (1 - single_rmp)^N)
+
+ggplot(res, aes(x = single_rmp, y = overall, group = factor(N), color = factor(N))) + 
+  geom_line() +
+  coord_cartesian(xlim = c(0, 0.07), ylim = c(0, 1)) + 
+  ylab("Probability of at least one false positive\nin N comparisons") + 
+  xlab("Probability of false positive for 1 comparison") + 
+  ggtitle("Wire Comparison: Controlling Overall False Positive Rate") + 
+  geom_vline(xintercept = 0.02, linetype = "dashed", color = "grey40") + 
+  geom_vline(xintercept = c(0.00704, 0.0724, 0.004545), linetype = "dotted", color = "grey50") + 
+  annotate(geom = "text", x = 0.021, y = 0.125, label = "Striated toolmark error rate estimate\nused in this paper", color = "grey40", hjust=0) + 
+  scale_color_discrete("# comparisons") + 
+  theme_bw()
+
+```
 
@@
 
-We need to have an explicit relationship between random match probability and amount of signal (assessed by length, number of features, overall striae depth, etc.).
+We need to have an explicit relationship between false positive rate and amount of signal (assessed by length, number of features, overall striae depth, etc.).
 Until we have this kind of relationship, it will be hard to use error rates (even if they were known for toolmarks) from algorithms or black-box studies in practical situations.

### 2023-11-21 PNAS/index.Rnw
updates and intro

@@
 \author[a,1,2]{Susan Vanderplas}
-\author[b,1]{Heike Hofmann}
-\author[b]{Alicia Carriquiry}
+\author[b, c]{Alicia Carriquiry}
+\author[b, c, 1]{Heike Hofmann}
 
 \affil[a]{Statistics Department, University of Nebraska Lincoln. 350 Hardin Hall, 3310 Holdrege North Wing, Lincoln, NE 68503}
-\affil[b]{Affiliation Two}
-\affil[c]{Affiliation Three}
+\affil[b]{Department of Statistics, Iowa State University. 1121 Snedecor Hall, 2438 Osborn Dr, Ames, IA 50011}
+\affil[c]{Center for Statistics and Applications in Forensic Evidence. 195 Durham Center, 613 Morrill Road, Ames, Iowa 50011}
 
 % Please give the surname of the lead author for the running footer
-\leadauthor{Lead author last name}
+\leadauthor{Vanderplas}
 
@@
 \authordeclaration{Please declare any competing interests here.}
-\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
+\equalauthors{\textsuperscript{1}SVP (Author One) contributed equally to this work with HH (Author Two).}
 \correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: susan.vanderplas@unl.edu}
@@
 % At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol
-\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}
+\keywords{Forensic Evidence $|$ Statistics $|$ Wire cuts $|$ Toolmark analysis}
 
@@
 
-Note: please start your introduction without including the word ``Introduction'' as a section heading (except for math articles in the Physical Sciences section); this heading is implied in the first paragraph.
+Multiple comparisons are a persistent problem for statistical methods and for science more generally.
+XXX cite XXX In forensic statistics, we have been focusing on the necessity of building up research databases which can be used to develop objective algorithms for evidence comparisons and explore closest non-matches as suggested in \cite{pcast}.
+While these databases (and open data more generally) are critically important, we also need to be careful about how our methods are affected by the multiple comparison problem.
 
-\section*{Guide to using this template on Overleaf}
+As we develop objective algorithms and standardize laboratory procedures in forensic pattern evidence, we need to be aware of the fact that we often are conducting many comparisons, either implicitly or explicitly.
+Algorithms in pattern evidence often utilize cross-correlation \cite{vorburgerApplicationsCrosscorrelationFunctions2011} to align different data sources, implicitly making comparisons along the entire length or surface of the data.
+Similarly, forensic examiners visually rotate or align two pieces of evidence under comparison microscopes.
+When we conduct database searches, these searches often return the $N$ best matches from the database, but this requires making comparisons to all of the database entries which meet the search criteria.
 
-Please note that whilst this template provides a preview of the typeset manuscript for submission, to help in this preparation, it will not necessarily be the final publication layout.
+Partial matches in fingerprint evidence have been shown to be problematic XXX cite XXX, and the AFIS database search conducted while investigating the Madrid bombing in 2004 was one factor cited in the inquiry into the false accusation of Brandon Mayfield.
+As forensic intelligence databases in other disciplines are implemented (such as IBIS in firearms), we must confront the issues with multiple comparisons and false positive rates before more embarrasing false accusations lead the public to question the validity of forensic science.
+
+\section*{Calculating the Number of Comparisons}
+
+% Note: please start your introduction without including the word ``Introduction'' as a section heading (except for math articles in the Physical Sciences section); this heading is implied in the first paragraph.
+
+% \section*{Guide to using this template on Overleaf}
+%
+% Please note that whilst this template provides a preview of the typeset manuscript for submission, to help in this preparation, it will not necessarily be the final publication layout.
 For more detailed information please see the \href{ https://www.pnas.org/page/authors/format}{PNAS Information for Authors}.
-
-If you have a question while using this template on Overleaf, please use the help menu (``?'') on the top bar to search for \href{https://www.overleaf.com/help}{help and tutorials}.
+%
+% If you have a question while using this template on Overleaf, please use the help menu (``?'') on the top bar to search for \href{https://www.overleaf.com/help}{help and tutorials}.
 You can also \href{https://www.overleaf.com/contact}{contact the Overleaf support team} at any time with specific questions about your manuscript or feedback on the template.
-
-\subsection*{Author Affiliations}
-
-Include department, institution, and complete address, with the ZIP/postal code, for each author.
+%
+% \subsection*{Author Affiliations}
+%
+% Include department, institution, and complete address, with the ZIP/postal code, for each author.
 Use lower case letters to match authors with institutions, as shown in the example.
@@
 For proper authentication, authors must provide their ORCID at submission and are not permitted to add ORCIDs on proofs.
-
-\subsection*{Submitting Manuscripts}
-
-All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}.
+%
+% \subsection*{Submitting Manuscripts}
+%
+% All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}.
 If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.
-
-\subsection*{Format}
-
-Many authors find it useful to organize their manuscripts with the following order of sections: title, author line and affiliations, keywords, abstract, introduction, results, discussion, materials and methods, acknowledgments, and references.
+%
+% \subsection*{Format}
+%
+% Many authors find it useful to organize their manuscripts with the following order of sections: title, author line and affiliations, keywords, abstract, introduction, results, discussion, materials and methods, acknowledgments, and references.
 Other orders and headings are permitted.
-
-\subsection*{Manuscript Length}
-
-Brief Reports are limited to 3 pages, which is approximately 1,600 words (including the manuscript text, title page, abstract, and figure legends), and 15 references.
+%
+% \subsection*{Manuscript Length}
+%
+% Brief Reports are limited to 3 pages, which is approximately 1,600 words (including the manuscript text, title page, abstract, and figure legends), and 15 references.
 Supporting information (SI) is limited to extended methods, essential supporting datasets, and videos (no additional tables or figures).
-
-\subsection*{References}
-
-References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g.
+%
+%
+% \subsection*{References}
+%
+% References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g.
 \cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric,phdthesis,masterthesis}.
 All references cited in the main text should be included in the main manuscript file.
-
-\subsection*{Data Archival}
-
-PNAS must be able to archive the data essential to a published article.
+%
+% \subsection*{Data Archival}
+%
+% PNAS must be able to archive the data essential to a published article.
 Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the \href{https://www.pnas.org/author-center/editorial-and-journal-policies#materials-and-data-availability}{Information for Authors}, is acceptable.
-
-\subsection*{Language-Editing Services}
-Prior to submission, authors who believe their manuscripts would benefit from professional editing are encouraged to use a language-editing service (see list at https://www.pnas.org/author-center/language-editing).
+%
+% \subsection*{Language-Editing Services}
+% Prior to submission, authors who believe their manuscripts would benefit from professional editing are encouraged to use a language-editing service (see list at https://www.pnas.org/author-center/language-editing).
 PNAS does not take responsibility for or endorse these services, and their use has no bearing on acceptance of a manuscript for publication.
-
-\begin{figure}[t!]
-\centering
-\includegraphics[width=.8\linewidth]{frog}
-\caption{Placeholder image of a frog with a long example legend to show justification setting.}
-\label{fig:frog}
-\end{figure}
-
-\begin{SCfigure*}[\sidecaptionrelwidth][t]
-\centering
-\includegraphics[width=11.4cm,height=11.4cm]{frog}
-\caption{This legend would be placed at the side of the figure, rather than below it.}\label{fig:side}
-\end{SCfigure*}
-
-\begin{table}[t!]
-\centering
-\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
-\begin{tabular}{lrrr}
-Species & CBS & CV & G3 \\
-\midrule
-1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
-2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
-3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
-\bottomrule
-\end{tabular}
-
-\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
-\end{table}
-
-\subsection*{Digital Figures}
-
-EPS, high-resolution PDF, and PowerPoint are preferred formats for figures that will be used in the main manuscript.
+%
+% \begin{figure}[t!]
+% \centering
+% \includegraphics[width=.8\linewidth]{frog}
+% \caption{Placeholder image of a frog with a long example legend to show justification setting.}
+% \label{fig:frog}
+% \end{figure}
+%
+%
+% \begin{SCfigure*}[\sidecaptionrelwidth][t]
+% \centering
+% \includegraphics[width=11.4cm,height=11.4cm]{frog}
+% \caption{This legend would be placed at the side of the figure, rather than below it.}\label{fig:side}
+% \end{SCfigure*}
+%
+% \begin{table}[t!]
+% \centering
+% \caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
+% \begin{tabular}{lrrr}
+% Species & CBS & CV & G3 \\
+% \midrule
+% 1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
+% 2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
+% 3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
+% \bottomrule
+% \end{tabular}
+%
+% \addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
+% \end{table}
+%
+%
+% \subsection*{Digital Figures}
+%
+% EPS, high-resolution PDF, and PowerPoint are preferred formats for figures that will be used in the main manuscript.
 Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
@@
 Include the font files for any text.
-
-Images must be provided at final size, preferably 1 column width (8.7cm).
+%
+% Images must be provided at final size, preferably 1 column width (8.7cm).
 Figures wider than 1 column should be sized to 11.4cm or 17.8cm wide.
 Numbers, letters, and symbols should be no smaller than 6 points (2mm) and no larger than 12 points (6mm) after reduction and must be consistent.
-
-Figures and tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.
-
-Figure \ref{fig:frog} shows an example of how to insert a column-wide figure.
+%
+% Figures and tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.
+%
+% Figure \ref{fig:frog} shows an example of how to insert a column-wide figure.
 To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment.
@@
 Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side legends.
-
-\subsection*{Tables}
-Tables should be included in the main manuscript file and should not be uploaded separately.
-
-\subsection*{Single column equations}
-
-Authors may use 1- or 2-column equations in their article, according to their preference.
-
-To allow an equation to span both columns, use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures.
-
-Note that the use of the \verb|widetext| environment for equations is not recommended, and should not be used.
-
-\begin{figure*}[bt!]
-\begin{align*}
-(x+y)^3&=(x+y)(x+y)^2\\
-       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
-       &=x^3+3x^2y+3xy^3+x^3.
-\end{align*}
-\end{figure*}
-
-\subsection*{Supporting Information (SI)}
-
-Authors are limited to the following types of SI: extended methods, datasets, videos, and 3D figures. Extended discussion is not permitted.
-
-\subsection*{SI Appendix}
-
-Supply extended materials and methods information in a separate SI Appendix PDF file.
+%
+% \subsection*{Tables}
+% Tables should be included in the main manuscript file and should not be uploaded separately.
+%
+% \subsection*{Single column equations}
+%
+% Authors may use 1- or 2-column equations in their article, according to their preference.
+%
+% To allow an equation to span both columns, use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures.
+%
+% Note that the use of the \verb|widetext| environment for equations is not recommended, and should not be used.
+%
+% \begin{figure*}[bt!]
+% \begin{align*}
+% (x+y)^3&=(x+y)(x+y)^2\\
+%        &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
+%        &=x^3+3x^2y+3xy^3+x^3.
+% \end{align*}
+% \end{figure*}
+%
+%
+%
+% \subsection*{Supporting Information (SI)}
+%
+% Authors are limited to the following types of SI: extended methods, datasets, videos, and 3D figures. Extended discussion is not permitted.
+%
+% \subsection*{SI Appendix}
+%
+% Supply extended materials and methods information in a separate SI Appendix PDF file.
 Include SI movie legends.
@@
 Supplementary figures and tables are not allowed.
-
-\subsubsection*{SI Datasets}
-
-Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be published in raw format and will not be edited or composed.
-
-\subsubsection*{SI Movies}
-
-Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and and include a brief legend for each movie in the main manuscript file.
+%
+%
+% \subsubsection*{SI Datasets}
+%
+% Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be published in raw format and will not be edited or composed.
+%
+%
+% \subsubsection*{SI Movies}
+%
+% Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and and include a brief legend for each movie in the main manuscript file.
 All movies should be submitted at the desired reproduction size and length.
 Movies should be no more than 10 MB in size.
-
+%
+%
+%
+%
 \matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required.
@@
 \bibsplit[3]
-%Use \bibsplit to split the references from the body of the text. Value "[3]" represents the number of reference in the left column (Note: Please avoid single column figures & tables on this page.)
+% %Use \bibsplit to split the references from the body of the text. Value "[3]" represents the number of reference in the left column (Note: Please avoid single column figures & tables on this page.)
 
 % Bibliography
-\bibliography{pnas-sample}
+% \bibliography{pnas-sample}
 
+\bibliography{../references}
 \end{document}

### 2023-11-21 LPR/index.qmd
compare wire to blade cut, not the other way round

@@
 We have spent considerable time assembling large data sets of tool marks, bullets, and cartridge cases to support our research as well as create public resources to accelerate future research [@csafeDataPortal].
-However, we are wary about the implications of database searches across forensic disciplines, for reasons also mentioned in the PCAST report: the necessity of understanding the random match probability for searches against these databases, in particular when using degraded evidence often found in crime scenes[@pcast, pg 52].
+However, we are wary about the implications of database searches across forensic disciplines, for reasons also mentioned in the PCAST report: the necessity of understanding the random match probability for searches against these databases, in particular when using degraded evidence often found in crime scenes [@pcast, pg 52].
 
-This issue has been raised in the past with regard to DNA [@thompson2003] and fingerprints [@fine2006]; one of the many issues identified in the aftermath of the 2004 Madrid bombing case was that the IAFIS database used to locate similar prints is extremely large and thus it is possible for the database to locate unusually similar non-matches.
+This issue has been raised in the past with regard to DNA [@thompson2003] and fingerprints [@fine2006]; one of the many issues identified in the aftermath of the 2004 Madrid bombing case was that the IAFIS database used to locate similar prints is extremely large and thus it is possible for the database to locate 'unusually' similar non-matches.
 This issue will almost certainly become an issue in other pattern disciplines as databases of firearms evidence such as NIBIN @nibin2021 become more commonly utilized by law enforcement.
@@
 When an examiner conducts a toolmark assessment of wire cut evidence, they are typically working with a tool recovered from the suspect as well as a wire recovered from the scene.
-The examiner will typically take the tool and use it to cut a piece of metal, known as a blade cut, in order to record the markings of the entire cutting surface simultaneously.
+The examiner will typically take the tool and use it to cut a piece of metal, known as a blade cut, to record the markings of the entire cutting surface simultaneously.
 This process may be repeated at multiple angles in order to record the effect of different contact surfaces on the resulting striations.
-The resulting blade cut will be compared to the wire using either visual inspection or perhaps in the future a numerical comparison algorithm similar to those developed for bullet striations [@hare2017automatic] and cartridge cases [@zemmels2023].
+Markings on the wire will then be compared to these blade cuts using either visual inspection or perhaps in the future a numerical comparison algorithm similar to those developed for bullet striations [@hare2017automatic] and cartridge cases [@zemmels2023].
+\hh{let's not just cite CSAFE we are not the only ones.
+}
 

### 2023-11-21 PNAS/index.Rnw

@@
 
-\firstpage{12}
+\firstpage[4]{6}
 % Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting.
@@
 If this is the case, add \parshape=0 to the end of the list environment.
-\dropcap{T}his PNAS journal template is provided to help you write your work in the correct journal format. Instructions for use are provided below.
+% \dropcap{T}his PNAS journal template is provided to help you write your work in the correct journal format. Instructions for use are provided below.
 
-Multiple comparisons are a persistent problem for statistical methods and for science more generally.
+\dropcap{M}ultiple comparisons are a persistent problem for statistical methods and for science more generally.
 XXX cite XXX In forensic statistics, we have been focusing on the necessity of building up research databases which can be used to develop objective algorithms for evidence comparisons and explore closest non-matches as suggested in \cite{pcast}.
@@
 Partial matches in fingerprint evidence have been shown to be problematic XXX cite XXX, and the AFIS database search conducted while investigating the Madrid bombing in 2004 was one factor cited in the inquiry into the false accusation of Brandon Mayfield.
-As forensic intelligence databases in other disciplines are implemented (such as IBIS in firearms), we must confront the issues with multiple comparisons and false positive rates before more embarrasing false accusations lead the public to question the validity of forensic science.
+As forensic intelligence databases in other disciplines are implemented (such as IBIS in firearms), we must confront the issues with multiple comparisons and false positive rates before more embarrassing false accusations lead the public to question the validity of forensic science.
+
+In this paper, we consider the multiple comparisons problems that arise from a relatively simple toolmark comparison: matching a cut wire to a wire-cutting tool.
+We describe the approach and calculate the number of comparisons performed, and then assess the likelihood of a false positive error under multiple comparisons using rates from published black-box studies.
+
+\section*{Wire Cut Forensics}
+
+A forensics examiner tasked with determining whether a wire in evidence was cut by a specific tool will create one or more blade cuts which can be compared to the cut surface of the wire.
+These cuts are made in a sheet of material similar to the wire's composition, and may be performed at multiple angles, as the angle of the tool to the substrate can affect which striations are recorded on the substrate surface.
+The blade cuts will then be compared to the wire under a comparison microscope; eventually, automatic comparison algorithms may also be validated for lab use.
+Each side of each blade cut will be compared to each side of the wire; depending on the tool design, there can be between 2 and 4 cutting surfaces in contact with the substrate.
 
 \section*{Calculating the Number of Comparisons}
+
+In order to calculate the number of comparisons, we define $\ell$ to be the length of the blade cut, and $d$ to be the diameter of the wire.
+Without loss of generality, we will assume that the wire is fully covered with striations suitable for comparison; if this is not the case, it is equivalent to changing the value of $d$.
+At minimum, there must be $\ell/d$ comparisons in order to assess whether one side of the blade cut matches one side of the wire; in this case, the comparisons are independent of each other (conditional on striation patterns) because no overlapping data is used.
+Of course, this overlooks the possibility of partial matches and assumes perfect alignment between the diameter of the wire and the portion of the blade used to cut the wire, which is unrealistic.
+
+At a maximum, both the blade and the wire are either digitally scanned at resolution $r$ mm, or visually examined using a microscope with a digital resolution that can be expressed as $r$ equivalent to the digital scan.
+In this case, there may be as many as $\ell/r - d/r + 1$ full comparisons performed in order to find the optimal alignment between the two samples.
+These comparisons are of course highly dependent, as each sequential comparison shares much of the same data with the previous and next comparisons; such dependency could be expressed via autocorrelation, but we will defer that derivation to a later paper.
+In this paper, let us consider the number of comparisons to lie somewhere between the minimum and maximum number of comparisons.
+
+Then, we must consider the number of surfaces which must be compared: the wire may have between one and two sets of striae (depending on tool configuration) and there may be between 2 and 4 different blade cut surfaces to examine.
+This results in a multiplier of between 2 and 8 comparisons to be conducted to compare the striations on the wire to the striations on the blade cuts, even before we consider the problem of cutting angles.
+
+\subsection*{A concrete example}
+Let us consider a wire-cutting tool with a 1.5 cm razor blade that meets a cast surface; the wire is held against this rectangular cast surface as the blade is pushed into the wire, splitting it in two.
+This is a minimal scenario - the wire will have striations from one side of the blade, while the blade itself has two cutting edges, which we will call side A and side B.
+The blade cuts of a sheet of aluminum that is 1.5 cm in width thus produce two striated edges corresponding to side A and side B which can be compared to other cuts to assess similarity.
+
+In addition, we have a 12 gauge aluminum wire (2 mm diameter) which may have been cut  by the wire-cutting tool described above.
+Class characteristics, which are shared by all tools of similar manufacture, appear to match: there is a flat impression on one side of the wire corresponding to the cast metal backstop of the tool, and the wire is cut such that the blade and the backstop appear to be perpendicular (that is, the wire appears to have been cut with a blade similar to the blade configuration of the tool in question).
+
+<<simple-problem-setup, echo = F, warning = F, message = F>>=
+ell <- 15
+d <- 2
+res <- 0.645/1000
+ncomp_min <- ell/d
+ncomp_max <- ell/res - d/res + 1
+@
+
+In this example, $\ell = \Sexpr{ell} mm$, $d = \Sexpr{2} mm$, and the minimal number of comparisons between the wire and one blade cut is $\ell/d = \Sexpr{ncomp_min}$.
+As there are two blade cuts (side A and side B), the minimal number of comparisons overall is \Sexpr{ncomp_min*2}, and these comparisons are independent.
+
+If the blade cuts and the wire edge are scanned at a resolution of $\Sexpr{res*1000} \mu m$, the resulting digital files will require $\Sexpr{ncomp_max}$ comparisons per blade cut, or $\Sexpr{ncomp_max*2}$ comparisons overall.
+These comparisons are required in order to find the optimal alignment between the wire and the blade cut; they are implicit in the calculation of cross-correlation which is either the first step or the only step used to objectively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
+This problem is not unique to algorithms, however; and examiner would need to physically align the wire and the blade cut by searching along the length of the cut to visually match striations, performing the same process physically that the algorithm performs computationally.
+While these sequential comparisons are of course highly autocorrelated, and we cannot assume sequential independence when calculating the probability of an error, they serve as an upper bound on the number of comparisons which could be performed.
 

### 2023-11-22 LPR/index.qmd
doesn't compile without error because of the \hh
Phrasing tweaks and updates

@@
 
-When an examiner conducts a toolmark assessment of wire cut evidence, they are typically working with a tool recovered from the suspect as well as a wire recovered from the scene.
+When an examiner conducts a tool mark assessment of wire cut evidence, they are typically working with a tool recovered from the suspect as well as a wire recovered from the scene.
 The examiner will typically take the tool and use it to cut a piece of metal, known as a blade cut, to record the markings of the entire cutting surface simultaneously.
 This process may be repeated at multiple angles in order to record the effect of different contact surfaces on the resulting striations.
-Markings on the wire will then be compared to these blade cuts using either visual inspection or perhaps in the future a numerical comparison algorithm similar to those developed for bullet striations [@hare2017automatic] and cartridge cases [@zemmels2023].
-\hh{let's not just cite CSAFE we are not the only ones.
-}
+Markings on the wire will then be compared to these blade cuts using either visual inspection or perhaps in the future a numerical comparison algorithm similar to those developed for bullet striations [@hare2017automatic; @mattijssen2021] and cartridge cases [@zemmels2023; @tong2014; @geradts2001].
 
@@
 
-All of these situations may occur during white box studies (though the final situation is very improbable).
+All of these situations may occur during black box studies (though the final situation is very improbable).
 As a result, perhaps the best estimate of the probability of a false identification that we can obtain is the false identification rate of one or more white box studies.
@@
 
+```{=tex}
 \begin{align*}
@@
 \end{align*}
-
+```
 It then becomes relevant to consider how low the error rate would need to be for striated comparisons in order to make wire cut examination plausible.
@@
 \end{align*}
-
+```
 We can explore what this means visually by considering different values of $N$ and $r$ to see how the overall false positive error rate behaves, as shown in @fig-control-overall-fpr.
-The astonishing conclusion is that even at the smallest error rate for striated comparisons, 0.45% [@bajic2020], we have about a 3% chance of at least one false positive conclusion when 10 comparisons are made and about a 37.5% chance of at least one false positive conclusion when 100 comparisons are made!
+The astonishing conclusion is that even at the smallest error rate for striated comparisons, 0.45% [@best2022], we have about a 3% chance of at least one false positive conclusion when 10 comparisons are made and about a 37.5% chance of at least one false positive conclusion when 100 comparisons are made!
 These results suggest that under current estimates of false positive error rates, it is nearly impossible for toolmark comparisons of wires to provide compelling, "beyond a reasonable doubt" levels of evidence.

### 2023-11-22 PNAS/index.Rnw
Phrasing tweaks and updates

@@
 % Use the lineno option to display guide line numbers if required.
-
+\usepackage{cleveref}
 \templatetype{pnasbriefreport} % Choose template
@@
 \doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+<<setup, include = F>>=
+library(knitr)
+opts_chunk$set(error=F, warning=F, message = F, dpi=300, echo = F)
+@
 
@@
 
-\dropcap{M}ultiple comparisons are a persistent problem for statistical methods and for science more generally.
-XXX cite XXX In forensic statistics, we have been focusing on the necessity of building up research databases which can be used to develop objective algorithms for evidence comparisons and explore closest non-matches as suggested in \cite{pcast}.
-While these databases (and open data more generally) are critically important, we also need to be careful about how our methods are affected by the multiple comparison problem.
+\dropcap{M}ultiple comparisons are a persistent problem for statistical methods and for science.
+In forensic statistics, researchers have been focusing on the necessity of building up research databases which can be used to develop objective algorithms for evidence comparisons and explore closest non-matches \citep{pcast}.
+While these databases (and open data) are critically important in forensics, we also need to be careful about how our methods are affected by the multiple comparison problem.
 
-As we develop objective algorithms and standardize laboratory procedures in forensic pattern evidence, we need to be aware of the fact that we often are conducting many comparisons, either implicitly or explicitly.
-Algorithms in pattern evidence often utilize cross-correlation \cite{vorburgerApplicationsCrosscorrelationFunctions2011} to align different data sources, implicitly making comparisons along the entire length or surface of the data.
+As we develop objective algorithms and standardize laboratory procedures in forensic pattern evidence, we must remain cognizant that we often are conducting many comparisons, either implicitly or explicitly.
+Algorithms in pattern evidence often utilize cross-correlation \citep{vorburgerApplicationsCrosscorrelationFunctions2011} to align different data sources, implicitly making comparisons along the entire length or surface of the data.
 Similarly, forensic examiners visually rotate or align two pieces of evidence under comparison microscopes.
@@
 While these sequential comparisons are of course highly autocorrelated, and we cannot assume sequential independence when calculating the probability of an error, they serve as an upper bound on the number of comparisons which could be performed.
+
+\section*{Probability of False Identifications}
+
+There are two components to a false positive rate: coincidental matches (that is, two pieces of evidence that happen to have similar characteristics but are not from the same source) and human/technical failures \citep[p 50]{pcast}.
+In objective disciplines with matching rules, such as DNA, these sources can be distinguished, and when coincidental match rates are very low, human/technical failures make up a large component of the error rate.
+However, in toolmark examination, we do not have an objective matching rule; examiners testify based on subjective, individual rules for how much similarity is enough for an identification.
+As a result, we cannot separate coincidental match rates from human/technical failures.
+<<fpr, include = F>>=
+errors <- c(20, 56, 1)
+knm <- c(2842, 774, 220)
+
+fpr <- sum(errors)/sum(knm)
+@
+
+Compounding this problem, open-set studies of striated comparisons (bullets, firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest the false positive rate for these comparisons is between 0.004545\citep{best2022} and 0.07235\citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of is \Sexpr{sprintf("%.02f%%", fpr*100)}, which is high relative to the consequences of a false identification of evidence from different sources.
+
+If we begin by considering the multiple comparison problem relative to the $\ell/d$ independent comparisons between a blade cut and a wire, we can calculate the overall false positive rate for all comparisons using $r$, the false positive rate for one comparison, and simple probability.
+Let $F_n:=\{\text{at least one false positive in }n\text{ comparisons}\}$.
+Then,
+\begin{align}
+P(F_1) &= r\nonumber\\
+P(\overline{F_1}) &= 1-r\nonumber\\
+P(F_n) &= \left[1-r\right]^n\nonumber\\
+P(\overline{F_n}) &= 1 - \left[P(\overline{F_1})\right]^n = 1 - \left[1 - r\right]^n.\label{eq:fpr-control}
+\end{align}
+Using this relationship between $F_n$ and the false positive rate $r$, we can determine what $r$ must be in order to limit $P(F_n)$ to an acceptable level by solving for $r$: $r = 1 - \left[1 - P(F_n)\right]^{1/n}$.
+
+\Cref{fig:control-overall-fpr} explores the implications of this relationship.
+The astonishing conclusion is that even at the smallest error rate for striated comparisons, we have about a 3\% chance of at least one false positive conclusion when 10 comparisons are made and about a 37.5\% chance of at least one false positive conclusion when 100 comparisons are made.
+These results suggest that using current estimates of false positive error rates, it is nearly impossible for wire comparisons to provide compelling, 'beyond a reasonable doubt' levels of evidence.
+<<control-overall-fpr, fig.cap="Exploration of overall error rate estimates for different values of $r$, the false positive rate for a single comparison.
+Dotted vertical lines show false positive error rate estimates (from left to right, \\citep{best2022,bajic2020,mattijssen2021}).", fig.width = 8, fig.height = 4>>=
+library(ggplot2)
+library(tidyr)
+library(dplyr)
+library(latex2exp)
+library(geomtextpath)
+single_rmp = exp(seq(-1000, 0, .1))
+N = c(10, 100, 1000, 10000)
+
+res = crossing(single_rmp, N) %>%
+  mutate(overall = 1 - (1 - single_rmp)^N)
+
+studies <- tibble(single_rmp = c( 0.004545,0.00704, 0.0724), study = c("Best (2022)","Bajic (2020)",  "Mattijssen (2021)"), y = c(-0.025, .15, 0.6), hjust = c(0, 0, 1))
+ggplot(res, aes(x = single_rmp, y = overall, group = factor(N), color = factor(N))) +
+  geom_line() +
+  coord_cartesian(xlim = c(0, 0.07), ylim = c(0, 1)) +
+  ylab("Probability of a coincidental match (N comparisons)") +
+  xlab("Probability of a false positive (1 comparison)") +
+  scale_y_continuous(breaks = seq(0, 1, .1), minor_breaks = seq(0, 1, .05)) +
+  ggtitle("Wire Comparison: Controlling Overall False Positive Rate") +
+  geom_vline(xintercept = 0.02, linetype = "dashed", color = "grey40") +
+  geom_vline(xintercept = studies$single_rmp, linetype = "51", color = "grey50") +
+  geom_label(data = studies[1,], aes(x = single_rmp, y = y, label = study), inherit.aes = F, hjust = studies$hjust[1], vjust = 0) +
+  geom_label(data = studies[2,], aes(x = single_rmp, y = y, label = study), inherit.aes = F, hjust = studies$hjust[2], vjust = 0) +
+  geom_label(data = studies[3,], aes(x = single_rmp, y = y, label = study), inherit.aes = F, hjust = studies$hjust[3], vjust = 0) +
+  annotate(geom = "label", x = 0.02, y = 0.1, label = "Pooled error rate \nused in this paper", color = "grey40", hjust=0) +
+  scale_color_discrete("# comparisons") +
+  theme_bw() +
+  theme(legend.position = c(1, 0), legend.justification = c(1, 0), legend.background = element_rect(color = "black"))
+@
+
+If we look at this data differently, we can assess the probability that $N$ comparisons are made accurately (e.g.
+there are no coincidental matches).
+It is clear from \Cref{fig:Accuracy} that with existing false positive error rate estimates between 0.07 and 0.0045, the accuracy of the aggregate comparison between a wire and a tool becomes questionable, as the number of candidate alignments that must be examined is too high.
+Even the most innocuous example we could come up with (small blade, only 2 cutting surfaces, and a relatively large wire) involved a minimum of \Sexpr{2*ncomp_min} completely independent comparisons.
+As a result, it seems improbable that a wire comparison made under current protocols (with corresponding error rates) would have an overall accuracy rate which would be reliable enough to be presented at trial.
+When we then consider that examiners would make cuts at multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons even in the simple case we have presented here, the probability of a coincidental match increases
+
+<<Accuracy, fig.cap="Probability of accurate decisions for $N$ comparisons for different error rates.", fig.width = 8, fig.height = 4>>=
+r_vals <- tibble(r = c(.0005, 0.001, 0.005, .01, 0.05),
+                 rlab = sprintf("P(F[1])==%0.05f", r))
+res2 <- crossing(r = r_vals$r,
+                 N = c(seq(1, 11, .01),seq(11.1, 109.9, .1), seq(110, 1100, 1), seq(1110, 10000, 10))) %>%
+  mutate(overall = (1 - r)^N) %>%
+  left_join(r_vals)
+
+breakpt <- .9
+
+error_res <- filter(res2, lead(overall, 1) <= breakpt)
+res2 <- filter(res2, overall >= breakpt)
+
+res2sum <- res2 %>% group_by(r) %>% filter(overall == min(overall))
+
+ggplot(data = res2, aes(x = N, y = overall, color = factor(r), label = rlab)) +
+  geom_textline(parse = T, gap = F, straight = F, padding = unit(1, "mm"), vjust = 0, linewidth = 0.75) +
+  geom_line(data = error_res, linetype = 2, linewidth = 0.75) +
+  xlab("Number of comparisons") +
+  scale_x_log10() +
+  geom_hline(yintercept = breakpt, linewidth = 1) +
+  geom_label(data = res2sum, aes(x = N, y = breakpt, label = paste0("N == ", round(N))), parse = T, vjust = 1, hjust = 1) +
+  ylab("Probability of no coincidental matches (N comparisons)") +
+  ggtitle("# Comparisons and Overall Accuracy") +
+  scale_color_discrete("FP Rate for single comparison", guide = 'none') +
+  theme_bw() +
+  coord_cartesian(ylim = c(0.75, 1.0)) +
+  theme(axis.text.y = element_text(hjust = 0.5, vjust = 0.5, angle = 90))
+
+@
 
@@
 %
-\matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required.
-
-\subsection*{Subsection for Method}
-Example text for subsection.
-}
-
-\showmatmethods{} % Display the Materials and Methods section
+% \matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required.
+%
+% \subsection*{Subsection for Method}
+% Example text for subsection.
+% }
+%
+% \showmatmethods{} % Display the Materials and Methods section
 
@@
 
-\bibsplit[3]
+% \bibsplit[3]
 % %Use \bibsplit to split the references from the body of the text. Value "[3]" represents the number of reference in the left column (Note: Please avoid single column figures & tables on this page.)

### 2023-11-22 PNAS/index.Rnw
Tweaking the introduction

@@
 Similarly, forensic examiners visually rotate or align two pieces of evidence under comparison microscopes.
-When we conduct database searches, these searches often return the $N$ best matches from the database, but this requires making comparisons to all of the database entries which meet the search criteria.
+Similarly, when we conduct database searches, these searches often return the $N$ best matches from the database, but this requires making comparisons to all of the database entries which meet the search criteria.
 
-Partial matches in fingerprint evidence have been shown to be problematic XXX cite XXX, and the AFIS database search conducted while investigating the Madrid bombing in 2004 was one factor cited in the inquiry into the false accusation of Brandon Mayfield.
-As forensic intelligence databases in other disciplines are implemented (such as IBIS in firearms), we must confront the issues with multiple comparisons and false positive rates before more embarrassing false accusations lead the public to question the validity of forensic science.
+The multiple comparison problems in forensics are distinct from the familywise error rate and false discovery rate problems that have been previously addressed in major ways: our tests are sequential, not concurrent, with dependencies depending on both the testing procedure and the data.
+In addition, an identification might be sufficient to end the search; as a result, this problem is more akin to a ``search until you find" problem than a situation where a family-wise error rate or false discovery rate is applicable - we are interested in the probability of at least one error in a series of sequential comparisons.
 
-In this paper, we consider the multiple comparisons problems that arise from a relatively simple toolmark comparison: matching a cut wire to a wire-cutting tool.
-We describe the approach and calculate the number of comparisons performed, and then assess the likelihood of a false positive error under multiple comparisons using rates from published black-box studies.
+The multiple testing problem is closely related to issues which have caused embarrassing false accusations and mistrust of forensic pattern evidence: partial matches and database searches.
+As we assemble research databases for explicitly making many comparisons and tackle problems which implicitly involve multiple comparisons, we must confront these issues before more embarrassing missteps lead the public to question the validity of forensic science.
+
+In this paper, we consider the multiple comparisons problem that arises from a relatively simple toolmark comparison: matching a cut wire to a wire-cutting tool.
+We describe the approach and calculate the number of comparisons performed, and then assess the likelihood of a false positive error considering the multiple comparisons performed, using error rates derived from published black-box studies.
 

### 2023-11-22 PNAS/index.Rnw
Add FDR/FWER paper
Tweaking formatting of first page, add acknowledgements.
Tweak first page, add acknowledgements.

@@
 
-\firstpage[4]{6}
+\firstpage[8]{7}
 % Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting.
@@
 
-\dropcap{M}ultiple comparisons are a persistent problem for statistical methods and for science.
+\dropcap{M}ultiple comparisons are a persistent problem when developing statistical methods for scientific problems \citep{benjaminiControllingFalseDiscovery1995}.
 In forensic statistics, researchers have been focusing on the necessity of building up research databases which can be used to develop objective algorithms for evidence comparisons and explore closest non-matches \citep{pcast}.
@@
 
-\section*{Wire Cut Forensics}
+\section*{Examination Process}
 
@@
 
-\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}
+\acknow{This work was funded (or partially funded) by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreements 70NANB15H176 and 70NANB20H019 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, Duke University, University of California Irvine, University of Virginia, West Virginia University, University of Pennsylvania, Swarthmore College and University of Nebraska, Lincoln.}
 

### 2023-11-22 PNAS/index.Rnw
not sure it compiles

--- 2023-11-22 PNAS/index.Rnw
@@
 
-As we develop objective algorithms and standardize laboratory procedures in forensic pattern evidence, we must remain cognizant that we often are conducting many comparisons, either implicitly or explicitly.
-Algorithms in pattern evidence often utilize cross-correlation \citep{vorburgerApplicationsCrosscorrelationFunctions2011} to align different data sources, implicitly making comparisons along the entire length or surface of the data.
+As we develop objective algorithms and standardize laboratory procedures in forensic pattern evidence, we must remain cognizant that a single conclusion is often based on  many comparisons, either implicitly or explicitly.
+\hh{Algorithms, in particular, make a large number of comparisons out of a user's sight.
+The cross-correlation function, for example, has been suggested \citep{vorburgerApplicationsCrosscorrelationFunctions2011} as one of the first quantitative measures in response to the NAS report (cite!!!), and is used in most pattern searching algorithms to 'find the best' alignment between two sources and also evaluate the overall similarity.
+}  %Algorithms in pattern evidence often utilize cross-correlation \citep{vorburgerApplicationsCrosscorrelationFunctions2011} to align different data sources, implicitly making comparisons along the entire length or surface of the data.
+\hh{'Finding the best' requires implicitly to slide one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source and keep track of the similarity measure.
+}
+
 Similarly, forensic examiners visually rotate or align two pieces of evidence under comparison microscopes.
@@
 
-Compounding this problem, open-set studies of striated comparisons (bullets, firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest the false positive rate for these comparisons is between 0.004545\citep{best2022} and 0.07235\citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of is \Sexpr{sprintf("%.02f%%", fpr*100)}, which is high relative to the consequences of a false identification of evidence from different sources.
+Compounding this problem, open-set studies of striated comparisons (bullets, firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest the false positive rate for these comparisons is between 0.004545 \citep{best2022} and 0.07235 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of is \Sexpr{sprintf("%.2f", fpr)}.
+%, which is high relative to the consequences of a false identification of evidence from different sources.
 

### 2023-11-22 PNAS/index.Rnw

--- 2023-11-22 PNAS/index.Rnw
@@
 
-As we develop objective algorithms and standardize laboratory procedures in forensic pattern evidence, we must remain cognizant that we often are conducting many comparisons, either implicitly or explicitly.
-Algorithms in pattern evidence often utilize cross-correlation \citep{vorburgerApplicationsCrosscorrelationFunctions2011} to align different data sources, implicitly making comparisons along the entire length or surface of the data.
+As we develop objective algorithms and standardize laboratory procedures in forensic pattern evidence, we must remain cognizant that a single conclusion is often based on  many comparisons, either implicitly or explicitly.
+\hh{Algorithms, in particular, make a large number of comparisons out of a user's sight.
+The cross-correlation function, for example, has been suggested \citep{vorburgerApplicationsCrosscorrelationFunctions2011} as one of the first quantitative measures in response to the NAS report (cite!!!), and is used in most pattern searching algorithms to 'find the best' alignment between two sources and also evaluate the overall similarity.
+}  %Algorithms in pattern evidence often utilize cross-correlation \citep{vorburgerApplicationsCrosscorrelationFunctions2011} to align different data sources, implicitly making comparisons along the entire length or surface of the data.
+\hh{'Finding the best' requires implicitly to slide one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source and keep track of the similarity measure.
+}
+
 Similarly, forensic examiners visually rotate or align two pieces of evidence under comparison microscopes.
-Similarly, when we conduct database searches, these searches often return the $N$ best matches from the database, but this requires making comparisons to all of the database entries which meet the search criteria.
+When we conduct database searches, these searches often return the $N$ best matches from the database, but this requires making comparisons to all of the database entries which meet the search criteria.
 

### 2023-11-22 LPR/index.qmd

@@
 If we start out by considering the first scenario (a single tool with a 1.5 cm blade compared to a single 2mm wire) and completely independent non-overlapping comparisons (7.5 comparisons) with only 2 cutting surfaces, we would effectively have 15 independent comparisons.
-Using this information, we can calculate the probability of at least one false positive error to be `r sprintf("%0.04f%%", 1-(1-fpr)^comparisons * 100)` as shown in Equation \ref{eq:fpr}.
+Using this information, we can calculate the probability of at least one false positive to be `r sprintf("%0.04f%%", 1-(1-fpr)^comparisons * 100)` as shown in Equation \ref{eq:fpr}.
 While far more comparisons are likely to be performed, these comparisons may have some dependencies (at which point our formula would need to account for the covariance between successive comparisons).

### 2023-11-22 PNAS/index.Rnw
Add colors and discussion
Add NAS report and fix class option clash

@@
 \usepackage{cleveref}
+\usepackage[dvipsnames]{xcolor} % colors
+\newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
+\newcommand{\hh}[1]{{\textcolor{Green}{#1}}}
 \templatetype{pnasbriefreport} % Choose template
@@
 
-\firstpage[8]{7}
+\firstpage[3]{8}
 % Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting.
@@
 \hh{Algorithms, in particular, make a large number of comparisons out of a user's sight.
-The cross-correlation function, for example, has been suggested \citep{vorburgerApplicationsCrosscorrelationFunctions2011} as one of the first quantitative measures in response to the NAS report (cite!!!), and is used in most pattern searching algorithms to 'find the best' alignment between two sources and also evaluate the overall similarity.
-}  %Algorithms in pattern evidence often utilize cross-correlation \citep{vorburgerApplicationsCrosscorrelationFunctions2011} to align different data sources, implicitly making comparisons along the entire length or surface of the data.
-\hh{'Finding the best' requires implicitly to slide one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source and keep track of the similarity measure.
+The cross-correlation function, for example, has been suggested \citep{vorburgerApplicationsCrosscorrelationFunctions2011} as one of the first quantitative measures in response to the NAS report\citep{nas2009}, and is used in most pattern searching algorithms to find the best alignment between two sources and also evaluate the overall similarity.
 }
+\hh{Finding the best} \svp{alignment} \hh{requires implicitly to slide one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source and keep track of the similarity measure.}
 
@@
 @
+
+\section*{Discussion}
 

### 2023-11-22 PNAS/index.Rnw
Updates

@@
 }
-\hh{Finding the best} \svp{alignment} \hh{requires implicitly to slide one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source and keep track of the similarity measure.}
+\hh{Finding the best} \svp{alignment} \hh{requires implicitly} \svp{sliding} \hh{one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source and} \svp{keeping} \hh{track of the similarity measure.}
 

### 2023-11-22 PNAS/index.Rnw
competing interest

@@
 \authorcontributions{Please provide details of author contributions here.}
-\authordeclaration{Please declare any competing interests here.}
+\authordeclaration{The authors have no competing interests to declare.}
 \equalauthors{\textsuperscript{1}SVP (Author One) contributed equally to this work with HH (Author Two).}
@@
 \hh{Finding the best} \svp{alignment} \hh{requires implicitly} \svp{sliding} \hh{one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source and} \svp{keeping} \hh{track of the similarity measure.}
+\svp{This mirrors the forensic examination process, as examiners visually}
+rotate or align two pieces of evidence under comparison microscopes.
 
-Similarly, forensic examiners visually rotate or align two pieces of evidence under comparison microscopes.
-When we conduct database searches, these searches often return the $N$ best matches from the database, but this requires making comparisons to all of the database entries which meet the search criteria.
+An explicit search, in constrast, occurs when we conduct database searches.
+These searches often return the $N$ best matches from the database, but this requires making comparisons to all of the database entries which meet the search criteria.
 

### 2023-11-22 PNAS/index.Rnw
Add koehler ref and add newlines in a few paragraphs so it's easier to edit/merge

@@
 \Cref{fig:control-overall-fpr} explores the implications of this relationship.
-The astonishing conclusion is that even at the smallest error rate for striated comparisons, we have about a 3\% chance of at least one false positive conclusion when 10 comparisons are made and about a 37.5\% chance of at least one false positive conclusion when 100 comparisons are made.
+The astonishing conclusion is that even at the smallest false positive error rate for striated comparisons, we have about a 3\% chance of at least one false positive conclusion when 10 comparisons are made and about a 37.5\% chance of at least one false positive conclusion when 100 comparisons are made.
 These results suggest that using current estimates of false positive error rates, it is nearly impossible for wire comparisons to provide compelling, 'beyond a reasonable doubt' levels of evidence.
+
 <<control-overall-fpr, fig.cap="Exploration of overall error rate estimates for different values of $r$, the false positive rate for a single comparison.
@@
 
-If we look at this data differently, we can assess the probability that $N$ comparisons are made accurately (e.g.
-there are no coincidental matches).
+If we look at this data differently, we can assess the probability that $N$ comparisons are made accurately (e.g. there are no coincidental matches or examiner errors).
 It is clear from \Cref{fig:Accuracy} that with existing false positive error rate estimates between 0.07 and 0.0045, the accuracy of the aggregate comparison between a wire and a tool becomes questionable, as the number of candidate alignments that must be examined is too high.

### 2023-11-22 PNAS/index.Rnw
color

--- 2023-11-22 PNAS/index.Rnw
@@
 \newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
-\newcommand{\hh}[1]{{\textcolor{Green}{#1}}}
+\newcommand{\hh}[1]{{\textcolor{ForestGreen}{#1}}}
 \templatetype{pnasbriefreport} % Choose template
@@
 }
-\hh{Finding the best} \svp{alignment} \hh{requires implicitly} \svp{sliding} \hh{one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source and} \svp{keeping} \hh{track of the similarity measure.}
+\hh{Finding the best} \svp{alignment} \hh{requires implicitly to slide one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source and keep track of the similarity measure.}
 

### 2023-11-22 PNAS/index.Rnw
conclusion and slight wording tweak

--- 2023-11-22 PNAS/index.Rnw
@@
 
-Compounding this problem, open-set studies of striated comparisons (bullets, firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest the false positive rate for these comparisons is between 0.004545 \citep{best2022} and 0.07235 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of is \Sexpr{sprintf("%.2f", fpr)}.
+Compounding this problem, open-set studies of striated comparisons (bullets, firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest the false positive rate for these comparisons is between 0.004545 \citep{best2022} and 0.07235 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
 %, which is high relative to the consequences of a false identification of evidence from different sources.
@@
 
-\section*{Discussion}
+\section*{Discussion \& Conclusions}
+
+In this short example, we have shown that the implicit comparisons performed during the forensic analysis of wire cuts have the potential to vastly increase the overall chance that at least one false positive error occurs.
+Our example is the simplest possible case for a wire cut analysis, but the conclusions apply to many other situations where we make multiple comparisons against a set of accumulated information, either in a database or in a series of test samples created during the examination process.
+As a result, it is critically important that forensic science grapple with this multiple comparison problem and modify lab procedures and testimony accordingly.
+We have three primary suggestions for improvements targeted at the problem of multiple comparisons.
+
+First, examiners should report (or defense attorneys should request) the overall length or area of surfaces generated from blade cuts, test fires, and equivalent procedures intended to assess the marks made by a tool submitted for comparison with evidence.
+Examiners should also report the total consecutive length or area of the recovered evidence.
+These pieces of information will allow for calculations of minimum and maximum number of comparisons (as we have demonstrated in this paper), and will facilitate calculation of overall error rates.
+
+Second, researchers in this field should conduct studies relating the length/area of comparison surface to the error rate.
+For instance, here, we have combined studies looking at bullet striations with studies on firing pin shear marks - the striated surfaces are of orders of magnitude different lengths, but they represent our best estimate of the error rate for striated comparisons.
+This was necessary, because we could not find black-box studies of wire cuts.
+These studies should be Type II studies as defined by Koehler \citep{koehlerIntuitiveErrorRate2017a} in that they should be designed to assess the actual error rate when examiners are making difficult comparisons.
+
+Finally, when databases are used at any stage of the forensic evidence evaluation process (from suitability assessment and triage to reports which will be used at trial), the number of database items searched and the number of results resturned must be reported.
+Furthermore, the number of results used for futher manual comparison should also be reported.
+This will help make the multiple comparison issue visible for everyone involved in evaluating the forensic evidence, from examiners themselves to jurors and judges.
+
+We believe that some of the multiple comparison problems we have assessed here can be overcome with targeted research and changes in how evidence is assessed, but the first step in this process has to be making the number of comparisons performed more visible to everyone involved in the process.
 

### 2023-11-22 PNAS/index.Rnw
Updates

@@
 In this short example, we have shown that the implicit comparisons performed during the forensic analysis of wire cuts have the potential to vastly increase the overall chance that at least one false positive error occurs.
+While often a forensic conclusion is presented as a single decision, underlying that decision are many more comparisons which are not discussed or reported upon.
+This omission is not necessarily the result of malice; it is the result of a protocol which focuses on finding the best possible match between two items and then evaluating that match.
+
 Our example is the simplest possible case for a wire cut analysis, but the conclusions apply to many other situations where we make multiple comparisons against a set of accumulated information, either in a database or in a series of test samples created during the examination process.
 As a result, it is critically important that forensic science grapple with this multiple comparison problem and modify lab procedures and testimony accordingly.
-We have three primary suggestions for improvements targeted at the problem of multiple comparisons.
+We have three primary suggestions for improvements targeted at making the multiple comparisons visible to everyone involved in the process of evidence evaluation.
 

### 2023-11-22 PNAS/index.Rnw
Title
first sentence

--- 2023-11-22 PNAS/index.Rnw
@@
 
-\dropcap{M}ultiple comparisons are a persistent problem when developing statistical methods for scientific problems \citep{benjaminiControllingFalseDiscovery1995}.
-In forensic statistics, researchers have been focusing on the necessity of building up research databases which can be used to develop objective algorithms for evidence comparisons and explore closest non-matches \citep{pcast}.
+\dropcap{I}n forensic evidence a single conclusion is often based on  many comparisons, either implicitly or explicitly.
+Multiple comparisons are a persistent problem when developing statistical methods for scientific problems \citep{benjaminiControllingFalseDiscovery1995}.
+In forensic statistics, researchers have been focusing on building up research databases to develop algorithms for objective evidence comparisons.
+\hh{In practice, these algorithms then allow searches across (vast) databases and report the best $N$ matches to a particular piece of evidence, raising the critically important problem of close non-matches }\citep{pcast}.
+
+%and explore closest non-matches .
 While these databases (and open data) are critically important in forensics, we also need to be careful about how our methods are affected by the multiple comparison problem.
 
-As we develop objective algorithms and standardize laboratory procedures in forensic pattern evidence, we must remain cognizant that a single conclusion is often based on  many comparisons, either implicitly or explicitly.
-\hh{Algorithms, in particular, make a large number of comparisons out of a user's sight.
-The cross-correlation function, for example, has been suggested \citep{vorburgerApplicationsCrosscorrelationFunctions2011} as one of the first quantitative measures in response to the NAS report\citep{nas2009}, and is used in most pattern searching algorithms to find the best alignment between two sources and also evaluate the overall similarity.
+This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{@fine2006}: one of the issues leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
+
+ \hh{Algorithms, in particular, make a large number of implicit comparisons out of a user's sight.
+The cross-correlation function, for example, has been suggested \citep{vorburgerApplicationsCrosscorrelationFunctions2011} as one of the first quantitative measures in response to the NAS report \citep{nas2009}, and is used in most pattern searching algorithms to find the best alignment between two sources and also evaluate the overall similarity.
 }
-\hh{Finding the best} \svp{alignment} \hh{requires implicitly to slide one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source and keep track of the similarity measure.}
+\hh{Finding the best} \svp{alignment} \hh{requires implicitly} \svp{sliding} \hh{one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source and} \svp{keeping} \hh{track of the similarity measure.}
+\svp{This mirrors the forensic examination process, as examiners visually}
+rotate or align two pieces of evidence under comparison microscopes.
 
-Similarly, forensic examiners visually rotate or align two pieces of evidence under comparison microscopes.
-When we conduct database searches, these searches often return the $N$ best matches from the database, but this requires making comparisons to all of the database entries which meet the search criteria.
+%An explicit search, in constrast, occurs when we conduct database searches.
+%These searches often return the $N$ best matches from the database, but this requires making comparisons to all of the database entries which meet the search criteria.
 

### 2023-11-22 PNAS/index.Rnw

--- 2023-11-22 PNAS/index.Rnw
@@
 
-\title{Template for preparing your Brief Report submission to PNAS using Overleaf}
+\title{Mirror, mirror on the wall. Who's the best match of them all?}
 
@@
 
-\dropcap{M}ultiple comparisons are a persistent problem when developing statistical methods for scientific problems \citep{benjaminiControllingFalseDiscovery1995}.
-In forensic statistics, researchers have been focusing on the necessity of building up research databases which can be used to develop objective algorithms for evidence comparisons and explore closest non-matches \citep{pcast}.
+\dropcap{I}n forensic evidence a single conclusion is often based on  many comparisons, either implicitly or explicitly.
+Multiple comparisons are a persistent problem when developing statistical methods for scientific problems \citep{benjaminiControllingFalseDiscovery1995}.
+In forensic statistics, researchers have been focusing on building up research databases to develop algorithms for objective evidence comparisons.
+\hh{In practice, these algorithms then allow searches across (vast) databases and report the best $N$ matches to a particular piece of evidence, raising the critically important problem of close non-matches }\citep{pcast}.
+
+%and explore closest non-matches .
 While these databases (and open data) are critically important in forensics, we also need to be careful about how our methods are affected by the multiple comparison problem.
 
-As we develop objective algorithms and standardize laboratory procedures in forensic pattern evidence, we must remain cognizant that a single conclusion is often based on  many comparisons, either implicitly or explicitly.
-\hh{Algorithms, in particular, make a large number of comparisons out of a user's sight.
-The cross-correlation function, for example, has been suggested \citep{vorburgerApplicationsCrosscorrelationFunctions2011} as one of the first quantitative measures in response to the NAS report\citep{nas2009}, and is used in most pattern searching algorithms to find the best alignment between two sources and also evaluate the overall similarity.
+This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{@fine2006}: one of the issues leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
+
+ \hh{Algorithms, in particular, make a large number of implicit comparisons out of a user's sight.
+The cross-correlation function, for example, has been suggested \citep{vorburgerApplicationsCrosscorrelationFunctions2011} as one of the first quantitative measures in response to the NAS report \citep{nas2009}, and is used in most pattern searching algorithms to find the best alignment between two sources and also evaluate the overall similarity.
 }
@@
 
-An explicit search, in constrast, occurs when we conduct database searches.
-These searches often return the $N$ best matches from the database, but this requires making comparisons to all of the database entries which meet the search criteria.
+%An explicit search, in constrast, occurs when we conduct database searches.
+%These searches often return the $N$ best matches from the database, but this requires making comparisons to all of the database entries which meet the search criteria.
 

### 2023-11-22 PNAS/index.Rnw
added citation, partial matches

@@
 \newcommand{\hh}[1]{{\textcolor{ForestGreen}{#1}}}
-\templatetype{pnasbriefreport} % Choose template
+\templatetype{pnasbriefreport}
+%\templatetype{pnasbriefreport} % Choose template
 % {pnasresearcharticle} = Template for a two-column research article
@@
 
-This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{@fine2006}: one of the issues leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
+This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehlerFingerprintErrorRate2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
+%
+\hh{In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.}
+%
+\hh{In order to avoid embarassing false accusation and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.}
+%
+\hh{Database searches are an example where the number of searches is explicitly available as the number of objects meeting the search criteria. }
+\hh{Another source of (implicit) multiple comparisons are partial matches. }
+
+\hh{XXX director's notes: XXX what is different between full and partial matches is that error rates in the literature are generally reported based on full length comparisons, where macro features are used to anchor comparisons (align along grooves, then zoom in and only 'wiggle' a bit to the left and right to adjust at the micro-scale).
+These kind of adjustments are 'baked' into published error rates.
+}
 
@@
 
-The multiple comparison problems in forensics are distinct from the familywise error rate and false discovery rate problems that have been previously addressed in major ways: our tests are sequential, not concurrent, with dependencies depending on both the testing procedure and the data.
+%The multiple comparison problems in forensics are distinct from the familywise error rate and false discovery rate problems that have been previously addressed in major ways: our tests are sequential, not concurrent, with dependencies depending on both the testing procedure and the data.
 In addition, an identification might be sufficient to end the search; as a result, this problem is more akin to a ``search until you find" problem than a situation where a family-wise error rate or false discovery rate is applicable - we are interested in the probability of at least one error in a series of sequential comparisons.
 
-The multiple testing problem is closely related to issues which have caused embarrassing false accusations and mistrust of forensic pattern evidence: partial matches and database searches.
-As we assemble research databases for explicitly making many comparisons and tackle problems which implicitly involve multiple comparisons, we must confront these issues before more embarrassing missteps lead the public to question the validity of forensic science.
+%The multiple testing problem is closely related to issues which have caused embarrassing false accusations and mistrust of forensic pattern evidence: partial matches and database searches.
+%As we assemble research databases for explicitly making many comparisons and tackle problems which implicitly involve multiple comparisons, we must confront these issues before more embarrassing missteps lead the public to question the validity of forensic science.
 
-In this paper, we consider the multiple comparisons problem that arises from a relatively simple toolmark comparison: matching a cut wire to a wire-cutting tool.
+Here, we consider the multiple comparisons problem that arises from a relatively simple toolmark comparison: matching a cut wire to a wire-cutting tool.
 We describe the approach and calculate the number of comparisons performed, and then assess the likelihood of a false positive error considering the multiple comparisons performed, using error rates derived from published black-box studies.

### 2023-11-22 PNAS/index.Rnw
partials (maybe) in the conclusion

@@
 \hh{Database searches are an example where the number of searches is explicitly available as the number of objects meeting the search criteria. }
-\hh{Another source of (implicit) multiple comparisons are partial matches. }
 
-\hh{XXX director's notes: XXX what is different between full and partial matches is that error rates in the literature are generally reported based on full length comparisons, where macro features are used to anchor comparisons (align along grooves, then zoom in and only 'wiggle' a bit to the left and right to adjust at the micro-scale).
+%\hh{XXX director's notes: XXX what is different between full and partial matches is that error rates in the literature are generally reported based on full length comparisons, where macro features are used to anchor comparisons (align along grooves, then zoom in and only 'wiggle' a bit to the left and right to adjust at the micro-scale).
 These kind of adjustments are 'baked' into published error rates.

### 2023-11-22 PNAS/index.Rnw
Fix first page
out with the wrong bib file

@@
 
-\firstpage[3]{8}
+\firstpage[3]{12}
 % Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting.
@@
 Here, we consider the multiple comparisons problem that arises from a relatively simple toolmark comparison: matching a cut wire to a wire-cutting tool.
-We describe the approach and calculate the number of comparisons performed, and then assess the likelihood of a false positive error considering the multiple comparisons performed, using error rates derived from published black-box studies.
+We describe the comparison approach, estimate the (minimal) number of comparisons needed, and discuss how the false discovery rate changes with the number of comparisons involved, using error rates derived from published black-box studies.
 

### 2023-11-22 PNAS/index.Rnw
Fix yaxis label

--- 2023-11-22 PNAS/index.Rnw
@@
 
-\firstpage[3]{8}
+\firstpage[4]{7}
 % Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting.
@@
   coord_cartesian(xlim = c(0, 0.07), ylim = c(0, 1)) +
-  ylab("Probability of a coincidental match (N comparisons)") +
+  ylab("Probability of a false positive (N comparisons)") +
   xlab("Probability of a false positive (1 comparison)") +

### 2023-11-22 PNAS/index.Rnw
references, images

--- 2023-11-22 PNAS/index.Rnw
@@
 
-\firstpage[3]{12}
+\firstpage[11]{6}
 % Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting.
@@
 
-This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehlerFingerprintErrorRate2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
+This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehler2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
 %
@@
 ggplot(res, aes(x = single_rmp, y = overall, group = factor(N), color = factor(N))) +
-  geom_line() +
   coord_cartesian(xlim = c(0, 0.07), ylim = c(0, 1)) +
-  ylab("Probability of a coincidental match (N comparisons)") +
-  xlab("Probability of a false positive (1 comparison)") +
-  scale_y_continuous(breaks = seq(0, 1, .1), minor_breaks = seq(0, 1, .05)) +
+  ylab("Probability of a false positive\n(N comparisons)") +
+  xlab("Probability of a false positive\n(1 comparison)") +
   ggtitle("Wire Comparison: Controlling Overall False Positive Rate") +
@@
   annotate(geom = "label", x = 0.02, y = 0.1, label = "Pooled error rate \nused in this paper", color = "grey40", hjust=0) +
-  scale_color_discrete("# comparisons") +
+  scale_color_discrete("# comparisons  ") +
+  geom_line(linewidth=1) +
   theme_bw() +
-  theme(legend.position = c(1, 0), legend.justification = c(1, 0), legend.background = element_rect(color = "black"))
+  theme(legend.position = c(1, 0),
+        legend.justification = c(1, 0),
+        legend.background = element_rect(color = "grey40", fill=alpha("white", 0.1))) +
+  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = seq(0, 1, .05)) +
+  scale_x_continuous(breaks = c(0, 0.02, 0.04, 0.06, studies$single_rmp),
+                     labels=c(0, 0.02, 0.04, 0.06, "0.004545", "\n0.00704","0.0724"))
+#  geom_hline(yintercept = 0.1, colour="black")
 @
@@
 ggplot(data = res2, aes(x = N, y = overall, color = factor(r), label = rlab)) +
-  geom_textline(parse = T, gap = F, straight = F, padding = unit(1, "mm"), vjust = 0, linewidth = 0.75) +
+  geom_textline(parse = T, gap = F, straight = F,
+                padding = unit(1, "mm"), vjust = 0, linewidth = 1) +
+  xlab("Number of comparisons") +
+  scale_x_log10(
+    breaks = c(1, 1000, 10000, res2sum$N),
+    minor_breaks = c((1:9)[-2],10*(3:9), 100*3:9, 1000*1:9)
+  ) +
+  geom_hline(yintercept = breakpt, linewidth = 1) +
+#  geom_label(
+#    data = res2sum,
+#    aes(x = N, y = breakpt, label = paste0("N == ", round(N))),
+#    parse = T, vjust = 1, hjust = 1) +
+  geom_segment(data = res2sum,
+     aes(x = N, xend=N, y = -Inf, yend=0.9), colour="grey30", linewidth=0.5) +
   geom_line(data = error_res, linetype = 2, linewidth = 0.75) +
-  xlab("Number of comparisons") +
-  scale_x_log10() +
-  geom_hline(yintercept = breakpt, linewidth = 1) +
-  geom_label(data = res2sum, aes(x = N, y = breakpt, label = paste0("N == ", round(N))), parse = T, vjust = 1, hjust = 1) +
-  ylab("Probability of no coincidental matches (N comparisons)") +
+  ylab("Probability false positive\n(N comparisons)") +
   ggtitle("# Comparisons and Overall Accuracy") +
@@
   theme_bw() +
-  coord_cartesian(ylim = c(0.75, 1.0)) +
+  coord_cartesian(ylim = c(0.775, 1.0)) +
   theme(axis.text.y = element_text(hjust = 0.5, vjust = 0.5, angle = 90))

### 2023-11-22 PNAS/index.Rnw
Fixing intro

--- 2023-11-22 PNAS/index.Rnw
@@
 
-\firstpage[4]{7}
+\firstpage{6}
 % Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting.
@@
 This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehlerFingerprintErrorRate2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
+\svp{Very close non-matches are a direct consequence of a large number of comparisons; they will occur more frequently as databases increase in size and algorithms become more powerful.}
+% In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.
 %
-\hh{In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.}
+In order to avoid embarassing false accusations and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.
 %
-\hh{In order to avoid embarassing false accusation and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.}
-%
-\hh{Database searches are an example where the number of searches is explicitly available as the number of objects meeting the search criteria. }
+% \hh{Database searches are an example where the number of searches is explicitly available as the number of objects meeting the search criteria.
+} % This doesn't seem super-relevant and distracts from the continuity between the algorithm discussion above and the next paragraph
 
@@
 
- \hh{Algorithms, in particular, make a large number of implicit comparisons out of a user's sight.
-The cross-correlation function, for example, has been suggested \citep{vorburgerApplicationsCrosscorrelationFunctions2011} as one of the first quantitative measures in response to the NAS report \citep{nas2009}, and is used in most pattern searching algorithms to find the best alignment between two sources and also evaluate the overall similarity.
-}
-\hh{Finding the best} \svp{alignment} \hh{requires implicitly} \svp{sliding} \hh{one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source and} \svp{keeping} \hh{track of the similarity measure.}
+Algorithms make a large number of comparisons \svp{that are not obvious to the user}
+%out of a user's sight.
+The cross-correlation function, for example, \svp{was} \citep{vorburgerApplicationsCrosscorrelationFunctions2011} one of the first quantitative measures \svp{proposed} in response to the NAS report \citep{nas2009}, and is used in \svp{many} pattern searching algorithms to find the best alignment between two sources and also evaluate the overall similarity.
+Finding the best \svp{alignment} requires implicitly sliding one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source \svp{ while keeping} track of \svp{the value of a} similarity measure.
 \svp{This mirrors the forensic examination process, as examiners visually}

### 2023-11-22 PNAS/index.Rnw

--- 2023-11-22 PNAS/index.Rnw
@@
 
-This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehler2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
+This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehlerFingerprintErrorRate2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
+\svp{Very close non-matches are a direct consequence of a large number of comparisons; they will occur more frequently as databases increase in size and algorithms become more powerful.}
+% In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.
+
+In order to avoid embarassing false accusations and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.
 %
-\hh{In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.}
-%
-\hh{In order to avoid embarassing false accusation and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.}
-%
-\hh{Database searches are an example where the number of searches is explicitly available as the number of objects meeting the search criteria. }
+% \hh{Database searches are an example where the number of searches is explicitly available as the number of objects meeting the search criteria.
+} % This doesn't seem super-relevant and distracts from the continuity between the algorithm discussion above and the next paragraph
 

### 2023-11-22 PNAS/index.Rnw
label

--- 2023-11-22 PNAS/index.Rnw
@@
   geom_segment(data = res2sum,
-     aes(x = N, xend=N, y = -Inf, yend=0.9), colour="grey30", linewidth=0.5) +
-  geom_line(data = error_res, linetype = 2, linewidth = 0.75) +
-  ylab("Probability false positive\n(N comparisons)") +
+     aes(x = N, xend=N, y = -Inf, yend=0.9, color=factor(r)),
+     linewidth=0.5) +
+  geom_line(data = error_res, linetype = 2, linewidth = 0.75, alpha =0.5) +
+  ylab("1 - Probability of false positive\n(N comparisons)") +
   ggtitle("# Comparisons and Overall Accuracy") +

### 2023-11-22 PNAS/index.Rnw
Fix paragraphs and rearrange one bit of intro
round x

--- 2023-11-22 PNAS/index.Rnw
@@
     breaks = c(1, 1000, 10000, res2sum$N),
+    labels = c("1", "1000", "10000", round(res2sum$N)),
     minor_breaks = c((1:9)[-2],10*(3:9), 100*3:9, 1000*1:9)

### 2023-11-24 PNAS/index.Rnw
Add figures and newlines
Not sure how to get the figure to be the right size... ugh
Updates

@@
 
-This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehlerFingerprintErrorRate2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
+This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehler2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
 \svp{Very close non-matches are a direct consequence of a large number of comparisons; they will occur more frequently as databases increase in size and algorithms become more powerful.}
@@
 
-In order to avoid embarassing false accusations and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.
 %
@@
 rotate or align two pieces of evidence under comparison microscopes.
-
+In order to avoid embarassing false accusations and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.
 %An explicit search, in constrast, occurs when we conduct database searches.
@@
 These cuts are made in a sheet of material similar to the wire's composition, and may be performed at multiple angles, as the angle of the tool to the substrate can affect which striations are recorded on the substrate surface.
-The blade cuts will then be compared to the wire under a comparison microscope; eventually, automatic comparison algorithms may also be validated for lab use.
+The blade cuts will then be compared to the wire under a comparison microscope in a process shown in \Cref{fig:wire-blade}, though eventually, automatic comparison algorithms may also be validated for lab use.
 Each side of each blade cut will be compared to each side of the wire; depending on the tool design, there can be between 2 and 4 cutting surfaces in contact with the substrate.
@@
 Of course, this overlooks the possibility of partial matches and assumes perfect alignment between the diameter of the wire and the portion of the blade used to cut the wire, which is unrealistic.
+
+\begin{figure}[t]
+\includegraphics[width=\columnwidth]{../fig/wire-blade-comparison.png}
+\caption{A comparison between a wire and a blade cut requires sliding the wire along the entire blade cut length to determine the best match (or whether there is a match).}\label{fig:wire-blade}
+\end{figure}
 
@@
 \subsection*{A concrete example}
-Let us consider a wire-cutting tool with a 1.5 cm razor blade that meets a cast surface; the wire is held against this rectangular cast surface as the blade is pushed into the wire, splitting it in two.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=\columnwidth]{../fig/wire_cutter_wires.png}
+\caption{RJ45 Crimp tool with a 1.5cm razor blade used for cutting. 1mm and 2mm diameter aluminum wires cut with the pliers are shown in a box in the top right corner.}\label{fig:tool-pic}
+\end{figure}
+
+Let us consider a wire-cutting tool with a 1.5 cm razor blade that meets a cast surface (one such tool is shown in \Cref{fig:tool-pic}); the wire is held against this rectangular cast surface as the blade is pushed into the wire, splitting it in two.
 This is a minimal scenario - the wire will have striations from one side of the blade, while the blade itself has two cutting edges, which we will call side A and side B.
@@
 In addition, we have a 12 gauge aluminum wire (2 mm diameter) which may have been cut  by the wire-cutting tool described above.
-Class characteristics, which are shared by all tools of similar manufacture, appear to match: there is a flat impression on one side of the wire corresponding to the cast metal backstop of the tool, and the wire is cut such that the blade and the backstop appear to be perpendicular (that is, the wire appears to have been cut with a blade similar to the blade configuration of the tool in question).
+Class characteristics, which are shared by all tools of similar manufacture, appear to match:
+there is a flat impression on one side of the wire corresponding to the cast metal backstop of the tool,
+and the wire is cut such that the blade and the backstop appear to be perpendicular
+(that is, the wire appears to have been cut with a blade similar to the blade configuration of the tool in question).
 

### 2023-11-29 PNAS/index.Rnw
Fix lingering issues.

@@
 opts_chunk$set(error=F, warning=F, message = F, dpi=300, echo = F)
+options(digits = 4)
 @
@@
 In forensic statistics, researchers have been focusing on building up research databases to develop algorithms for objective evidence comparisons.
-\hh{In practice, these algorithms then allow searches across (vast) databases and report the best $N$ matches to a particular piece of evidence, raising the critically important problem of close non-matches }\citep{pcast}.
+In practice, these algorithms then allow searches across (vast) databases and report the best $N$ matches to a particular piece of evidence, raising the critically important problem of close non-matches\citep{pcast}.
 
@@
 This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehler2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
-\svp{Very close non-matches are a direct consequence of a large number of comparisons; they will occur more frequently as databases increase in size and algorithms become more powerful.}
+Very close non-matches are a direct consequence of a large number of comparisons; they will occur more frequently as databases increase in size and algorithms become more powerful.
 % In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.
@@
 %out of a user's sight.
-The cross-correlation function, for example, \svp{was} \citep{vorburgerApplicationsCrosscorrelationFunctions2011} one of the first quantitative measures \svp{proposed} in response to the NAS report \citep{nas2009}, and is used in \svp{many} pattern searching algorithms to find the best alignment between two sources and also evaluate the overall similarity.
-Finding the best \svp{alignment} requires implicitly sliding one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source \svp{ while keeping} track of \svp{the value of a} similarity measure.
-\svp{This mirrors the forensic examination process, as examiners visually}
-rotate or align two pieces of evidence under comparison microscopes.
-In order to avoid embarassing false accusations and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.
+The cross-correlation function\citep{vorburgerApplicationsCrosscorrelationFunctions2011}, for example, was one of the first quantitative measures proposed in response to the NAS report \citep{nas2009}, and is used in many pattern searching algorithms to find the best alignment between two sources and also evaluate the overall similarity.
+Finding the best alignment requires implicitly sliding one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source while keeping track of the value of a similarity measure.
+This mirrors the forensic examination process, as examiners visually rotate or align two pieces of evidence under comparison microscopes.
+In order to avoid embarrassing false accusations and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.
 %An explicit search, in constrast, occurs when we conduct database searches.
@@
 These cuts are made in a sheet of material similar to the wire's composition, and may be performed at multiple angles, as the angle of the tool to the substrate can affect which striations are recorded on the substrate surface.
-The blade cuts will then be compared to the wire under a comparison microscope in a process shown in \Cref{fig:wire-blade}, though eventually, automatic comparison algorithms may also be validated for lab use.
+The blade cuts will then be compared to the wire under a comparison microscope, though eventually, automatic comparison algorithms may also be validated for lab use.
 Each side of each blade cut will be compared to each side of the wire; depending on the tool design, there can be between 2 and 4 cutting surfaces in contact with the substrate.
@@
 Without loss of generality, we will assume that the wire is fully covered with striations suitable for comparison; if this is not the case, it is equivalent to changing the value of $d$.
+\svp{An illustration of the sliding comparison process is shown in \Cref{fig:wire-blade}.}
 At minimum, there must be $\ell/d$ comparisons in order to assess whether one side of the blade cut matches one side of the wire; in this case, the comparisons are independent of each other (conditional on striation patterns) because no overlapping data is used.
@@
 \begin{figure}[t]
-\includegraphics[width=\columnwidth]{../fig/wire-blade-comparison.png}
-\caption{A comparison between a wire and a blade cut requires sliding the wire along the entire blade cut length to determine the best match (or whether there is a match).}\label{fig:wire-blade}
+\includegraphics[width=\columnwidth]{../fig/wire-blade-comparison-label.png}
+\caption{A comparison between a wire and a blade cut requires sliding the wire along the entire blade cut length to determine the best match (or whether there is a match).
+\svp{Surfaces shown are rendered 2D topographical scans of a wire and blade cut taken with a confocal light microscope.}}\label{fig:wire-blade}
 \end{figure}
@@
 
-Compounding this problem, open-set studies of striated comparisons (bullets, firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest the false positive rate for these comparisons is between 0.004545 \citep{best2022} and 0.07235 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
+Compounding this problem, open-set studies of striated comparisons (bullets, firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest the false positive rate for these comparisons is between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
 %, which is high relative to the consequences of a false identification of evidence from different sources.
@@
 Let $F_n:=\{\text{at least one false positive in }n\text{ comparisons}\}$.
-Then,
-\begin{align}
-P(F_1) &= r\nonumber\\
-P(\overline{F_1}) &= 1-r\nonumber\\
-P(F_n) &= \left[1-r\right]^n\nonumber\\
-P(\overline{F_n}) &= 1 - \left[P(\overline{F_1})\right]^n = 1 - \left[1 - r\right]^n.\label{eq:fpr-control}
-\end{align}
+Then, $P(F_1) = r$, $P(\overline{F_1}) = 1-r$, $P(F_n) = [1-r]^n$, and $P(\overline{F_n}) = 1 - \left[P(\overline{F_1})\right]^n$.
 Using this relationship between $F_n$ and the false positive rate $r$, we can determine what $r$ must be in order to limit $P(F_n)$ to an acceptable level by solving for $r$: $r = 1 - \left[1 - P(F_n)\right]^{1/n}$.
@@
 As a result, it seems improbable that a wire comparison made under current protocols (with corresponding error rates) would have an overall accuracy rate which would be reliable enough to be presented at trial.
-When we then consider that examiners would make cuts at multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons even in the simple case we have presented here, the probability of a coincidental match increases
+When we then consider that examiners would make cuts at multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons even in the simple case we have presented here, the probability of a coincidental match increases further still.
 

### 2023-11-29 PNAS/index.Rnw
Fix sig figs and other minor issues

@@
 
-If the blade cuts and the wire edge are scanned at a resolution of $\Sexpr{res*1000} \mu m$, the resulting digital files will require $\Sexpr{ncomp_max}$ comparisons per blade cut, or $\Sexpr{ncomp_max*2}$ comparisons overall.
+If the blade cuts and the wire edge are scanned at a resolution of $\Sexpr{res*1000} \mu m$, the resulting digital files will require \svp{approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons per blade cut; $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall}.
 These comparisons are required in order to find the optimal alignment between the wire and the blade cut; they are implicit in the calculation of cross-correlation which is either the first step or the only step used to objectively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.

### 2023-11-29 PNAS/index.Rnw
Removing fig 3 and adding abstract, tweaking phrasing, etc. to try to get down to 3 pages.

@@
 \begin{abstract}
-Please provide an abstract of no more than 250 words in a single paragraph.
-Abstracts should explain to the general reader the major contributions of the article.
-References in the abstract must be cited in full within the abstract itself and cited in the text.
+Wires have striations left from the cutting process; as in other forms of toolmark analysis, these striae are used to connect the wire to the tool used.
+In this paper, we show that the process of wire comparison introduces complexities not present in better-investigated forensic toolmark areas such as bullets, as wire comparison inherently requires multiple distinct comparisons, increasing the expected false positive rate.
+We call attention to the multiple comparison problem in wire examination and relate it to other situations in forensics that involve multiple comparisons, such as database searches.
 \end{abstract}
@@
 
-\firstpage[11]{6}
+\firstpage[8]{6}
 % Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting.
@@
 Multiple comparisons are a persistent problem when developing statistical methods for scientific problems \citep{benjaminiControllingFalseDiscovery1995}.
-In forensic statistics, researchers have been focusing on building up research databases to develop algorithms for objective evidence comparisons.
-In practice, these algorithms then allow searches across (vast) databases and report the best $N$ matches to a particular piece of evidence, raising the critically important problem of close non-matches\citep{pcast}.
+Researchers have been focusing on building up databases to develop algorithms for objective evidence comparisons.
+\svp{Then, the algorithms facilitate} searches across (vast) databases, reporting the best $N$ matches to a particular piece of evidence.
+This raises the critically important problem of close non-matches\citep{pcast}.
 
@@
 
-Algorithms make a large number of comparisons \svp{that are not obvious to the user}
-%out of a user's sight.
-The cross-correlation function\citep{vorburgerApplicationsCrosscorrelationFunctions2011}, for example, was one of the first quantitative measures proposed in response to the NAS report \citep{nas2009}, and is used in many pattern searching algorithms to find the best alignment between two sources and also evaluate the overall similarity.
+\svp{Compounding this issue,} algorithms make a large number of comparisons that are not obvious to the user.
+The cross-correlation function\citep{vorburgerApplicationsCrosscorrelationFunctions2011} was one of the first quantitative measures proposed in response to the NAS report \citep{nas2009}, and is used in many pattern searching algorithms to find the best alignment between two sources and also evaluate the overall similarity.
 Finding the best alignment requires implicitly sliding one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source while keeping track of the value of a similarity measure.
@@
 In order to avoid embarrassing false accusations and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.
-%An explicit search, in constrast, occurs when we conduct database searches.
-%These searches often return the $N$ best matches from the database, but this requires making comparisons to all of the database entries which meet the search criteria.
-
-%The multiple comparison problems in forensics are distinct from the familywise error rate and false discovery rate problems that have been previously addressed in major ways: our tests are sequential, not concurrent, with dependencies depending on both the testing procedure and the data.
-In addition, an identification might be sufficient to end the search; as a result, this problem is more akin to a ``search until you find" problem than a situation where a family-wise error rate or false discovery rate is applicable - we are interested in the probability of at least one error in a series of sequential comparisons.
-
-%The multiple testing problem is closely related to issues which have caused embarrassing false accusations and mistrust of forensic pattern evidence: partial matches and database searches.
-%As we assemble research databases for explicitly making many comparisons and tackle problems which implicitly involve multiple comparisons, we must confront these issues before more embarrassing missteps lead the public to question the validity of forensic science.
 
@@
 At a maximum, both the blade and the wire are either digitally scanned at resolution $r$ mm, or visually examined using a microscope with a digital resolution that can be expressed as $r$ equivalent to the digital scan.
-In this case, there may be as many as $\ell/r - d/r + 1$ full comparisons performed in order to find the optimal alignment between the two samples.
-These comparisons are of course highly dependent, as each sequential comparison shares much of the same data with the previous and next comparisons; such dependency could be expressed via autocorrelation, but we will defer that derivation to a later paper.
-In this paper, let us consider the number of comparisons to lie somewhere between the minimum and maximum number of comparisons.
+In this case, there may be as many as $\ell/r - d/r + 1$ comparisons performed in order to find the optimal alignment between the two samples.
+These comparisons are of course highly dependent, as each sequential comparison shares much of the same data with the previous and next comparisons.
+% ; such dependency could be expressed via autocorrelation, but we will defer that derivation to a later paper.
+\svp{For simplicity}, let us consider the number of comparisons to lie somewhere between \svp{these two estimates}.
 
 Then, we must consider the number of surfaces which must be compared: the wire may have between one and two sets of striae (depending on tool configuration) and there may be between 2 and 4 different blade cut surfaces to examine.
-This results in a multiplier of between 2 and 8 comparisons to be conducted to compare the striations on the wire to the striations on the blade cuts, even before we consider the problem of cutting angles.
+This results in a multiplier of between 2 and 8 comparisons to be conducted to compare the striations on the wire to the striations on the blade cuts.
 
@@
 If the blade cuts and the wire edge are scanned at a resolution of $\Sexpr{res*1000} \mu m$, the resulting digital files will require \svp{approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons per blade cut; $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall}.
-These comparisons are required in order to find the optimal alignment between the wire and the blade cut; they are implicit in the calculation of cross-correlation which is either the first step or the only step used to objectively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
-This problem is not unique to algorithms, however; and examiner would need to physically align the wire and the blade cut by searching along the length of the cut to visually match striations, performing the same process physically that the algorithm performs computationally.
+These comparisons are required in order to find the optimal alignment between the wire and the blade cut; they are implicit in the calculation of cross-correlation which is either the first or the only step used to objectively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
+This problem is not unique to algorithms; an examiner would need to physically align the wire and the blade cut by searching along the length of the cut to visually match striations, performing the same process physically that the algorithm performs computationally.
 While these sequential comparisons are of course highly autocorrelated, and we cannot assume sequential independence when calculating the probability of an error, they serve as an upper bound on the number of comparisons which could be performed.
@@
 
-Compounding this problem, open-set studies of striated comparisons (bullets, firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest the false positive rate for these comparisons is between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
-%, which is high relative to the consequences of a false identification of evidence from different sources.
+Compounding this problem, open-set studies of striated comparisons (bullets and firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest \svp{that a starting point for estimating the} false positive rate for these comparisons is between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
 
 If we begin by considering the multiple comparison problem relative to the $\ell/d$ independent comparisons between a blade cut and a wire, we can calculate the overall false positive rate for all comparisons using $r$, the false positive rate for one comparison, and simple probability.
-Let $F_n:=\{\text{at least one false positive in }n\text{ comparisons}\}$.
-Then, $P(F_1) = r$, $P(\overline{F_1}) = 1-r$, $P(F_n) = [1-r]^n$, and $P(\overline{F_n}) = 1 - \left[P(\overline{F_1})\right]^n$.
-Using this relationship between $F_n$ and the false positive rate $r$, we can determine what $r$ must be in order to limit $P(F_n)$ to an acceptable level by solving for $r$: $r = 1 - \left[1 - P(F_n)\right]^{1/n}$.
+Let $$F_n:=\{\text{at least one false pos.
+in }n\text{ disjoint comparisons}\}.$$ Then, $P(F_1) = r$, $P(\overline{F_1}) = 1-r$, $P(F_n) = [1-r]^n$, and $P(\overline{F_n}) = 1 - \left[P(\overline{F_1})\right]^n$.
+% Using this relationship between $F_n$ and the false positive rate $r$, we can determine what $r$ must be in order to limit $P(F_n)$ to an acceptable level by solving for $r$: $r = 1 - \left[1 - P(F_n)\right]^{1/n}$.
+\svp{\Cref{fig:Accuracy} shows $P(\overline{F_n})$ on the y-axis; if we set 90\% as the acceptable probability, we can make 2 comparisons when the single-comparison FPR is 0.05 and 10 comparisons at 0.01.
+Thus, at an error rate of \Sexpr{sprintf("%.2f", fpr)}, we can conduct approximately 5 comparisons while keeping the FPR for all comparisons under 10\%.}
 
-\Cref{fig:control-overall-fpr} explores the implications of this relationship.
-The astonishing conclusion is that even at the smallest false positive error rate for striated comparisons, we have about a 3\% chance of at least one false positive conclusion when 10 comparisons are made and about a 37.5\% chance of at least one false positive conclusion when 100 comparisons are made.
-These results suggest that using current estimates of false positive error rates, it is nearly impossible for wire comparisons to provide compelling, 'beyond a reasonable doubt' levels of evidence.
+\svp{It is clear that under these constraints, the} accuracy of the aggregate comparison between a wire and a tool becomes questionable, as the number of candidate alignments that must be examined is too high.
+Even the most innocuous example we could come up with (small blade, only 2 cutting surfaces, and a relatively large wire) involved a minimum of \Sexpr{2*ncomp_min} completely independent comparisons.
+As a result, it seems improbable that a wire comparison made under current protocols (with corresponding error rates) would have an overall accuracy rate which would be reliable enough to be presented at trial.
+When we then consider that examiners would make cuts at multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons even in the simple case we have presented here, the probability of a false positive error increases further still.
 
 <<control-overall-fpr, fig.cap="Exploration of overall error rate estimates for different values of $r$, the false positive rate for a single comparison.
-Dotted vertical lines show false positive error rate estimates (from left to right, \\citep{best2022,bajic2020,mattijssen2021}).", fig.width = 8, fig.height = 4>>=
+Dotted vertical lines show false positive error rate estimates (from left to right, \\citep{best2022,bajic2020,mattijssen2021}).", fig.width = 8, fig.height = 4, include = F>>=
 library(ggplot2)
@@
 
-If we look at this data differently, we can assess the probability that $N$ comparisons are made accurately (e.g. there are no coincidental matches or examiner errors).
-It is clear from \Cref{fig:Accuracy} that with existing false positive error rate estimates between 0.07 and 0.0045, the accuracy of the aggregate comparison between a wire and a tool becomes questionable, as the number of candidate alignments that must be examined is too high.
-Even the most innocuous example we could come up with (small blade, only 2 cutting surfaces, and a relatively large wire) involved a minimum of \Sexpr{2*ncomp_min} completely independent comparisons.
-As a result, it seems improbable that a wire comparison made under current protocols (with corresponding error rates) would have an overall accuracy rate which would be reliable enough to be presented at trial.
-When we then consider that examiners would make cuts at multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons even in the simple case we have presented here, the probability of a coincidental match increases further still.
-
 <<Accuracy, fig.cap="Probability of accurate decisions for $N$ comparisons for different error rates.", fig.width = 8, fig.height = 4>>=
@@
 res2 <- crossing(r = r_vals$r,
-                 N = c(seq(1, 11, .01),seq(11.1, 109.9, .1), seq(110, 1100, 1), seq(1110, 10000, 10))) %>%
+                 N = c(seq(1, 11, .01),seq(11.1, 109.9, .1), seq(110, 1100, 1))) %>%
   mutate(overall = (1 - r)^N) %>%
@@
 
-We believe that some of the multiple comparison problems we have assessed here can be overcome with targeted research and changes in how evidence is assessed, but the first step in this process has to be making the number of comparisons performed more visible to everyone involved in the process.
+We believe that the multiple comparison problems we have discussed here can be overcome with research and changes in evidence assessment procedures, but the first step is making the number of comparisons performed more visible to everyone involved in the process.
 

### 2023-12-06 PNAS/index.Rnw
Alicia updates

@@
 \newcommand{\hh}[1]{{\textcolor{ForestGreen}{#1}}}
+\newcommand{\ac}[1]{{\textcolor{Purple}{#1}}}
 \templatetype{pnasbriefreport}
@@
 
-\title{Mirror, mirror on the wall. Who's the best match of them all?}
+\title{Mirror, mirror on the wall. Which's the best match of them all?}
 
@@
 \begin{abstract}
-Wires have striations left from the cutting process; as in other forms of toolmark analysis, these striae are used to connect the wire to the tool used.
-In this paper, we show that the process of wire comparison introduces complexities not present in better-investigated forensic toolmark areas such as bullets, as wire comparison inherently requires multiple distinct comparisons, increasing the expected false positive rate.
+
+\ac{When wires are cut, the tool produces striations on the cut surfaced}; as in other forms of toolmark analysis, these striae are used to connect the wire to the tool that was used \ac{in the cutting process.
+Here, we argue that the practice} of comparing two wire cut surfaces introduces complexities not present in better-investigated forensic examination  of toolmars such as those observed on bullets, as wire comparisons inherently require multiple distinct comparisons, increasing the expected false positive rate.
 We call attention to the multiple comparison problem in wire examination and relate it to other situations in forensics that involve multiple comparisons, such as database searches.
@@
 
-\dropcap{I}n forensic evidence a single conclusion is often based on  many comparisons, either implicitly or explicitly.
-Multiple comparisons are a persistent problem when developing statistical methods for scientific problems \citep{benjaminiControllingFalseDiscovery1995}.
-Researchers have been focusing on building up databases to develop algorithms for objective evidence comparisons.
-\svp{Then, the algorithms facilitate} searches across (vast) databases, reporting the best $N$ matches to a particular piece of evidence.
-This raises the critically important problem of close non-matches\citep{pcast}.
+\dropcap{I}n forensic \ac{evaluations}, a single conclusion often \ac{relies} on  many comparisons, either implicitly or explicitly.
+Multiple comparisons arise persistently when developing statistical methods to address scientific problems \citep{benjaminiControllingFalseDiscovery1995}, and can significantly increase the probability of false "discoveries".
+Yet, this potentially significant issue is often ignored.
+\ac{Now that vast databases and efficient algorithms are routinely used in forensic evaluations to find "matches" to crime scene items, the problems of close non-matches\citep{pcast}, and of the erosion of confidence due to multiple comparisons, become critically important.}
+%Researchers have been focusing on building up databases to develop algorithms for objective evidence comparisons.
+%\svp{Then, the algorithms facilitate} searches across (vast) databases, reporting the best $N$ matches to a particular piece of evidence.
+%This raises the critically important problem of .
 
@@
 
-This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehler2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
-Very close non-matches are a direct consequence of a large number of comparisons; they will occur more frequently as databases increase in size and algorithms become more powerful.
+A similar concern was raised in the past with regard to DNA \citep{thompson2003} and latent print evaluations  \citep{koehler2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid train bombing case was that the large size of the IAFIS database used to search for similar prints made it possible to locate `unusually' close non-matches.
+\ac{The probability of finding  close non-matches for an item of evidence is directly proportional to the size of the database available for searching;} false "matches" occur more frequently as databases increase in size and algorithms become more powerful.
 % In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.
@@
 
-\svp{Compounding this issue,} algorithms make a large number of comparisons that are not obvious to the user.
-The cross-correlation function\citep{vorburgerApplicationsCrosscorrelationFunctions2011} was one of the first quantitative measures proposed in response to the NAS report \citep{nas2009}, and is used in many pattern searching algorithms to find the best alignment between two sources and also evaluate the overall similarity.
-Finding the best alignment requires implicitly sliding one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source while keeping track of the value of a similarity measure.
-This mirrors the forensic examination process, as examiners visually rotate or align two pieces of evidence under comparison microscopes.
-In order to avoid embarrassing false accusations and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.
+\svp{Compounding this issue,} the use of novel algorithms typically results in a large number of comparisons that may not obvious to the user.
+For example, the cross-correlation function\citep{vorburgerApplicationsCrosscorrelationFunctions2011} was one of the first measures proposed for quantifying the similarity  between two patterns in response to the \ac{ 2009 NRC} report \citep{nas2009}, and continues to be used in many pattern searching algorithms to find the best alignment between two images and also to quantify their overall similarity.
+Finding the best alignment often consists in sliding one surfaceacross the whole length (for one-dimensional patterns, such as striations) or area (for two dimensional sources, such as impression marks) of the other item while keeping track of the value of a similarity measure.
+This mirrors the forensic examination process, where examiners visually rotate or align two surfaces under comparison microscopes.
+In order to avoid \ac{problematic} false accusations and a further erosion of the public's trust in science, it is important to address the problem of multiple comparisons by controlling their effect on false discovery rates.
 
-Here, we consider the multiple comparisons problem that arises from a relatively simple toolmark comparison: matching a cut wire to a wire-cutting tool.
-We describe the comparison approach, estimate the (minimal) number of comparisons needed, and discuss how the false discovery rate changes with the number of comparisons involved, using error rates derived from published black-box studies.
+Here, we consider the multiple comparisons problem that arises from a relatively simple toolmark examination: matching a cut wire to a wire-cutting tool.
+We describe the comparison approach, estimate the (minimal) number of comparisons that are needed to carry out the examination, and discuss how the false discovery rate changes with the number of comparisons involved, using error rates derived from published black-box studies.
 
@@
 
-A forensics examiner tasked with determining whether a wire in evidence was cut by a specific tool will create one or more blade cuts which can be compared to the cut surface of the wire.
+A forensics examiner tasked with determining whether a wire in evidence was cut by a specific tool will create one or more blade cuts \ac{using the putative tool,} which can be compared to the cut surface of the wire.
 These cuts are made in a sheet of material similar to the wire's composition, and may be performed at multiple angles, as the angle of the tool to the substrate can affect which striations are recorded on the substrate surface.
 The blade cuts will then be compared to the wire under a comparison microscope, though eventually, automatic comparison algorithms may also be validated for lab use.
-Each side of each blade cut will be compared to each side of the wire; depending on the tool design, there can be between 2 and 4 cutting surfaces in contact with the substrate.
+Each side of each blade cut will be compared to each side of the wire; depending on the tool design, there can be between two and four cutting surfaces in contact with the substrate.
 
@@
 
-In order to calculate the number of comparisons, we define $\ell$ to be the length of the blade cut, and $d$ to be the diameter of the wire.
-Without loss of generality, we will assume that the wire is fully covered with striations suitable for comparison; if this is not the case, it is equivalent to changing the value of $d$.
-\svp{An illustration of the sliding comparison process is shown in \Cref{fig:wire-blade}.}
-At minimum, there must be $\ell/d$ comparisons in order to assess whether one side of the blade cut matches one side of the wire; in this case, the comparisons are independent of each other (conditional on striation patterns) because no overlapping data is used.
+In order to calculate the number of possible comparisons to be carried out in the course of the examination, we define $\ell$ to be the length of the blade cut, and $d$ to be the diameter of the wire.
+Without loss of generality, we assume that the wire is fully covered with striations suitable for comparison; if this is not the case, we adjust the value of $d$.
+\svp{An illustration of the sliding comparison process is shown in \Cref{fig:wire-blade}.} Imagine that you start from the left end of the blade and move the wire surface \ac{$d$ units at a time} along the blade.
+At minimum, there must be $\ell/d$ comparisons in order to assess whether one side of the blade cut matches one side of the wire; in this case, the comparisons are independent of each other (conditional on striation patterns) because \ac{each comparison involves different surfaces.}
+%no overlapping data is used.
 Of course, this overlooks the possibility of partial matches and assumes perfect alignment between the diameter of the wire and the portion of the blade used to cut the wire, which is unrealistic.
@@
 In this case, there may be as many as $\ell/r - d/r + 1$ comparisons performed in order to find the optimal alignment between the two samples.
-These comparisons are of course highly dependent, as each sequential comparison shares much of the same data with the previous and next comparisons.
+These comparisons are \ac{no longer independent}, as each sequential comparison shares much of the same data with the previous and next comparisons.
 % ; such dependency could be expressed via autocorrelation, but we will defer that derivation to a later paper.
@@
 
-Then, we must consider the number of surfaces which must be compared: the wire may have between one and two sets of striae (depending on tool configuration) and there may be between 2 and 4 different blade cut surfaces to examine.
+\ac{Finally}, we must consider the number of surfaces which must be compared: the wire may have between one and two sets of striae (depending on tool configuration) and there may be between 2 and 4 different blade cut surfaces to examine.
 This results in a multiplier of between 2 and 8 comparisons to be conducted to compare the striations on the wire to the striations on the blade cuts.
@@
 Let us consider a wire-cutting tool with a 1.5 cm razor blade that meets a cast surface (one such tool is shown in \Cref{fig:tool-pic}); the wire is held against this rectangular cast surface as the blade is pushed into the wire, splitting it in two.
-This is a minimal scenario - the wire will have striations from one side of the blade, while the blade itself has two cutting edges, which we will call side A and side B.
+This is a minimal scenario - the wire will acquire striations from one side of the blade, while the blade itself has two cutting edges, which we will call side A and side B.
 The blade cuts of a sheet of aluminum that is 1.5 cm in width thus produce two striated edges corresponding to side A and side B which can be compared to other cuts to assess similarity.
@@
 In this example, $\ell = \Sexpr{ell} mm$, $d = \Sexpr{2} mm$, and the minimal number of comparisons between the wire and one blade cut is $\ell/d = \Sexpr{ncomp_min}$.
-As there are two blade cuts (side A and side B), the minimal number of comparisons overall is \Sexpr{ncomp_min*2}, and these comparisons are independent.
+As there are two blade cuts (side A and side B), the minimal number of comparisons overall is \Sexpr{ncomp_min*2}, and these comparisons are \ac{non-overlapping and} independent.
 
 If the blade cuts and the wire edge are scanned at a resolution of $\Sexpr{res*1000} \mu m$, the resulting digital files will require \svp{approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons per blade cut; $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall}.
-These comparisons are required in order to find the optimal alignment between the wire and the blade cut; they are implicit in the calculation of cross-correlation which is either the first or the only step used to objectively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
+These comparisons are required in order to find the optimal alignment between the wire and the blade cut; they are implicit in the calculation of cross-correlation which is the first \ac{and often} the only step used to objectively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
 This problem is not unique to algorithms; an examiner would need to physically align the wire and the blade cut by searching along the length of the cut to visually match striations, performing the same process physically that the algorithm performs computationally.
-While these sequential comparisons are of course highly autocorrelated, and we cannot assume sequential independence when calculating the probability of an error, they serve as an upper bound on the number of comparisons which could be performed.
+While these sequential comparisons are highly autocorrelated, and we cannot assume sequential independence when calculating the probability of an error, they serve as an upper bound on the number of comparisons which could be performed.
 
@@
 
-There are two components to a false positive rate: coincidental matches (that is, two pieces of evidence that happen to have similar characteristics but are not from the same source) and human/technical failures \citep[p 50]{pcast}.
+There are \ac{at least }two \ac{contributors} to a false positive rate: coincidental matches (that is, two pieces of evidence that happen to have similar characteristics but are not from the same source) and human/technical failures \citep[p 50]{pcast}.
 In objective disciplines with matching rules, such as DNA, these sources can be distinguished, and when coincidental match rates are very low, human/technical failures make up a large component of the error rate.
@@
 
-Compounding this problem, open-set studies of striated comparisons (bullets and firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest \svp{that a starting point for estimating the} false positive rate for these comparisons is between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
+%Compounding this problem, o
+Open-set studies of striated comparisons (bullets and firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest \svp{that a starting point for estimating the} false positive rate for these comparisons is between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
 
-If we begin by considering the multiple comparison problem relative to the $\ell/d$ independent comparisons between a blade cut and a wire, we can calculate the overall false positive rate for all comparisons using $r$, the false positive rate for one comparison, and simple probability.
+If we begin by considering the multiple comparison problem relative to the $\ell/d$ independent comparisons between a blade cut and a wire, we can compute the overall false positive rate for all comparisons using $r$, the false positive rate for one comparison, and simple probability \ac{calculations}.
 Let $$F_n:=\{\text{at least one false pos.
@@
 % Using this relationship between $F_n$ and the false positive rate $r$, we can determine what $r$ must be in order to limit $P(F_n)$ to an acceptable level by solving for $r$: $r = 1 - \left[1 - P(F_n)\right]^{1/n}$.
-\svp{\Cref{fig:Accuracy} shows $P(\overline{F_n})$ on the y-axis; if we set 90\% as the acceptable probability, we can make 2 comparisons when the single-comparison FPR is 0.05 and 10 comparisons at 0.01.
-Thus, at an error rate of \Sexpr{sprintf("%.2f", fpr)}, we can conduct approximately 5 comparisons while keeping the FPR for all comparisons under 10\%.}
+\svp{\Cref{fig:Accuracy} shows $P(\overline{F_n})$ on the y-axis; if we set 90\% as the acceptable} \ac{examination-wide false positive rate (FPR)}, we can make 2 comparisons when the single-comparison FPR is 0.05 and 10 comparisons at 0.01.
+Thus, at an error rate of \Sexpr{sprintf("%.2f", fpr)}, we can conduct approximately 5 comparisons \ac{in the course of an examination without exceeding the examination-wide FPR of 10\%.}
 
-\svp{It is clear that under these constraints, the} accuracy of the aggregate comparison between a wire and a tool becomes questionable, as the number of candidate alignments that must be examined is too high.
+\svp{It is clear that under these constraints, the} accuracy of the \ac{forensic evaluation that involves multiple comparisons} between a wire and a tool \ac{is likely to be low}, as the number of candidate alignments that must be examined is too high.
 Even the most innocuous example we could come up with (small blade, only 2 cutting surfaces, and a relatively large wire) involved a minimum of \Sexpr{2*ncomp_min} completely independent comparisons.
-As a result, it seems improbable that a wire comparison made under current protocols (with corresponding error rates) would have an overall accuracy rate which would be reliable enough to be presented at trial.
+As a result, \ac{findings from}  wire comparisons made under current protocols (using error rates recently published in the forensics literature) \ac{are probably not} reliable enough to be presented at trial.
 When we then consider that examiners would make cuts at multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons even in the simple case we have presented here, the probability of a false positive error increases further still.
@@
 
-In this short example, we have shown that the implicit comparisons performed during the forensic analysis of wire cuts have the potential to vastly increase the overall chance that at least one false positive error occurs.
-While often a forensic conclusion is presented as a single decision, underlying that decision are many more comparisons which are not discussed or reported upon.
-This omission is not necessarily the result of malice; it is the result of a protocol which focuses on finding the best possible match between two items and then evaluating that match.
+\ac{Forensic practitioners often report the findings from their examinations in the form of a categorical conclusion reflecting a single decision.
+This can be misleading when the one decision relies on multiple comparisons which are not individually presented in reports or testimony.} In this short contribution, we have shown that the implicit comparisons performed during the forensic analysis of wire cuts have the potential to vastly increase the overall chance that at least one false positive error occurs.
+%While often a forensic conclusion is presented as a single decision, underlying that decision are many more comparisons which are not discussed or reported upon.
+\ac{Omission of the underlying} details is not necessarily \ac{due to} malice; rather, it results from protocols which focus on finding the best possible match between two items and then evaluating that match, \ac{and also from lack of familiarity with basic statistical ideas.}
 
-Our example is the simplest possible case for a wire cut analysis, but the conclusions apply to many other situations where we make multiple comparisons against a set of accumulated information, either in a database or in a series of test samples created during the examination process.
-As a result, it is critically important that forensic science grapple with this multiple comparison problem and modify lab procedures and testimony accordingly.
-We have three primary suggestions for improvements targeted at making the multiple comparisons visible to everyone involved in the process of evidence evaluation.
+In our example \ac{we highlight a simple} case where a wire \ac{ is cut using a two-sided blade}, but the findings apply to many other situations where \ac{a forensic evaluation involves} multiple comparisons against a set of accumulated information.
+\ac{Examples include a database search or a situation similar to the example where multiple comparisons of evidence against reference samples are conducted.}
+%either in a database or in a series of test samples created during the examination process.
+As a result, it is critically important that forensic \ac{practitioners become aware of the impact of carrying out multiple comparisons on their final conclusions}
+%science grapple with this multiple comparison problem and 
+and that they modify lab procedures and testimony accordingly.
+\ac{We suggest three strategies to improve transparency of forensic examinations and to produce more reliable estimates of examination-wide error rates.}
+%We have three primary suggestions for improvements targeted at making the multiple comparisons visible to everyone involved in the process of evidence evaluation.
 
@@
 
-Finally, when databases are used at any stage of the forensic evidence evaluation process (from suitability assessment and triage to reports which will be used at trial), the number of database items searched and the number of results resturned must be reported.
+Finally, when databases are used at any stage of the forensic evidence evaluation process (from suitability assessment and triage to reports which will be used at trial), the number of database items searched and the number of results returned must be reported.
 Furthermore, the number of results used for futher manual comparison should also be reported.
+For example, if a firearms examiner searches a local NIBIN database with 1000 entries, requests the 20 closest matches to her evidence, and then carries out a physical examination of five exemplars from the list of 20, all of those values should be clearly reported to enable calculation of the potential FPR.
 This will help make the multiple comparison issue visible for everyone involved in evaluating the forensic evidence, from examiners themselves to jurors and judges.
 
-We believe that the multiple comparison problems we have discussed here can be overcome with research and changes in evidence assessment procedures, but the first step is making the number of comparisons performed more visible to everyone involved in the process.
+%We believe that the multiple comparison problems we have discussed here can be overcome with research and changes in evidence assessment procedures, but the first step is making the number of comparisons performed more visible to everyone involved in the process.
 

### 2023-12-12 PNAS/index.Rnw
dependency

@@
 
-\dropcap{I}n forensic \ac{evaluations}, a single conclusion often \ac{relies} on  many comparisons, either implicitly or explicitly.
+\dropcap{I}n forensic evaluations, a single conclusion often relies on  many comparisons, either implicitly or explicitly.
 Multiple comparisons arise persistently when developing statistical methods to address scientific problems \citep{benjaminiControllingFalseDiscovery1995}, and can significantly increase the probability of false "discoveries".
 Yet, this potentially significant issue is often ignored.
-\ac{Now that vast databases and efficient algorithms are routinely used in forensic evaluations to find "matches" to crime scene items, the problems of close non-matches\citep{pcast}, and of the erosion of confidence due to multiple comparisons, become critically important.}
+Now that vast databases and efficient algorithms are routinely used in forensic evaluations to find "matches" to crime scene items, the problems of close non-matches \citep{pcast}, and of the erosion of confidence due to multiple comparisons, become critically important.
 %Researchers have been focusing on building up databases to develop algorithms for objective evidence comparisons.
@@
 
-A similar concern was raised in the past with regard to DNA \citep{thompson2003} and latent print evaluations  \citep{koehler2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid train bombing case was that the large size of the IAFIS database used to search for similar prints made it possible to locate `unusually' close non-matches.
-\ac{The probability of finding  close non-matches for an item of evidence is directly proportional to the size of the database available for searching;} false "matches" occur more frequently as databases increase in size and algorithms become more powerful.
+These concerns were raised in the past with regard to DNA \citep{thompson2003} and latent print evaluations  \citep{koehler2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid train bombing case was that the large size of the IAFIS database used to search for similar prints made it possible to locate `unusually' close non-matches.
+The probability of finding  close non-matches for an item of evidence is directly linked to the size of the database available for searching: false "matches" occur more frequently as databases increase in size and algorithms are more powerful.
 % In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.
@@
 
-\svp{Compounding this issue,} the use of novel algorithms typically results in a large number of comparisons that may not obvious to the user.
-For example, the cross-correlation function\citep{vorburgerApplicationsCrosscorrelationFunctions2011} was one of the first measures proposed for quantifying the similarity  between two patterns in response to the \ac{ 2009 NRC} report \citep{nas2009}, and continues to be used in many pattern searching algorithms to find the best alignment between two images and also to quantify their overall similarity.
-Finding the best alignment often consists in sliding one surfaceacross the whole length (for one-dimensional patterns, such as striations) or area (for two dimensional sources, such as impression marks) of the other item while keeping track of the value of a similarity measure.
-This mirrors the forensic examination process, where examiners visually rotate or align two surfaces under comparison microscopes.
-In order to avoid \ac{problematic} false accusations and a further erosion of the public's trust in science, it is important to address the problem of multiple comparisons by controlling their effect on false discovery rates.
+Compounding this issue, algorithmic results are typically based on a large number of comparisons that may not obvious to the user.
+For example, a typical task in the forensic examination process is to find the best alignment between two items of evidence.
+For one-dimensional patterns, such as striation marks, this consists of sliding one item along the full length of the other item.
+For two-dimensional patterns, such as impression marks on an area, an additional rotation is involved.
+In practice, examiners visually rotate or align two surfaces under comparison microscopes.
+Algorithmically this visual search is mirrored by finding a maximum of the (area) cross-correlation function.
+The cross-correlation function was one of the first measures proposed \citep{vorburgerApplicationsCrosscorrelationFunctions2011} for quantifying the similarity  between two patterns in response to the  2009 NRC report \citep{nas2009}, and continues to be used in many pattern searching algorithms to find the best alignment between two images and  quantify their overall similarity.
+
+In order to avoid problematic false accusations and a further erosion of the public's trust in science, it is important to address the problem of multiple comparisons by controlling their effect on false discovery rates.
 
@@
 
-A forensics examiner tasked with determining whether a wire in evidence was cut by a specific tool will create one or more blade cuts \ac{using the putative tool,} which can be compared to the cut surface of the wire.
+A forensics examiner tasked with determining whether a wire in evidence was cut by a specific tool will create one or more blade cuts using the putative tool which can be compared to the cut surface of the wire.
+
 These cuts are made in a sheet of material similar to the wire's composition, and may be performed at multiple angles, as the angle of the tool to the substrate can affect which striations are recorded on the substrate surface.
@@
 In order to calculate the number of possible comparisons to be carried out in the course of the examination, we define $\ell$ to be the length of the blade cut, and $d$ to be the diameter of the wire.
-Without loss of generality, we assume that the wire is fully covered with striations suitable for comparison; if this is not the case, we adjust the value of $d$.
-\svp{An illustration of the sliding comparison process is shown in \Cref{fig:wire-blade}.} Imagine that you start from the left end of the blade and move the wire surface \ac{$d$ units at a time} along the blade.
-At minimum, there must be $\ell/d$ comparisons in order to assess whether one side of the blade cut matches one side of the wire; in this case, the comparisons are independent of each other (conditional on striation patterns) because \ac{each comparison involves different surfaces.}
+Here, we assume that striations  suitable for comparison are visible across the full width $d$ of the wire; if this is not the case, we adjust the value of $d$.
+An illustration of the sliding comparison process is shown in \Cref{fig:wire-blade}.
+Imagine that a comparison starts from the left end of the blade and moves the wire surface $d$ units at a time along the blade.
+At minimum, there must be $\ell/d$ comparisons in order to assess whether one side of the blade cut matches one side of the wire.
+We can generally assume that these comparisons are independent  because each comparison involves different surfaces.
+\hh{We need to further assume that the length $d$ is wide enough to cover some striations; this ensures that the width of $d$ is wider than any significant autocorrelation  observed for the striations on the blade}.
 %no overlapping data is used.
-Of course, this overlooks the possibility of partial matches and assumes perfect alignment between the diameter of the wire and the portion of the blade used to cut the wire, which is unrealistic.
+\hh{Note, that this is a rather crude assessment of the bare minimal number of comparisons.
+}% Of course, this overlooks the possibility of partial matches and assumes perfect alignment between the diameter of the wire and the portion of the blade used to cut the wire, which is unrealistic.
 
@@
 \caption{A comparison between a wire and a blade cut requires sliding the wire along the entire blade cut length to determine the best match (or whether there is a match).
-\svp{Surfaces shown are rendered 2D topographical scans of a wire and blade cut taken with a confocal light microscope.}}\label{fig:wire-blade}
+Surfaces shown are rendered 2D topographical scans of a wire and blade cut taken with a confocal light microscope.}\label{fig:wire-blade}
 \end{figure}
 
-At a maximum, both the blade and the wire are either digitally scanned at resolution $r$ mm, or visually examined using a microscope with a digital resolution that can be expressed as $r$ equivalent to the digital scan.
+At a maximum, both the blade and the wire are either digitally scanned at resolution of $r$ pixel per mm, or visually examined using a microscope with a digital resolution that can be expressed as equivalent to the digital scan.
 In this case, there may be as many as $\ell/r - d/r + 1$ comparisons performed in order to find the optimal alignment between the two samples.
-These comparisons are \ac{no longer independent}, as each sequential comparison shares much of the same data with the previous and next comparisons.
+These comparisons are no longer independent, as each sequential comparison shares much of the same data with the previous and next comparisons.
 % ; such dependency could be expressed via autocorrelation, but we will defer that derivation to a later paper.
-\svp{For simplicity}, let us consider the number of comparisons to lie somewhere between \svp{these two estimates}.
+For simplicity, let us consider the number of comparisons to lie somewhere between these two estimates.
 
-\ac{Finally}, we must consider the number of surfaces which must be compared: the wire may have between one and two sets of striae (depending on tool configuration) and there may be between 2 and 4 different blade cut surfaces to examine.
+Finally, we must consider the number of surfaces which must be compared: the wire may have between one and two sets of striae (depending on tool configuration) and there may be between 2 and 4 different blade cut surfaces to examine.
 This results in a multiplier of between 2 and 8 comparisons to be conducted to compare the striations on the wire to the striations on the blade cuts.
@@
 
-Let us consider a wire-cutting tool with a 1.5 cm razor blade that meets a cast surface (one such tool is shown in \Cref{fig:tool-pic}); the wire is held against this rectangular cast surface as the blade is pushed into the wire, splitting it in two.
-This is a minimal scenario - the wire will acquire striations from one side of the blade, while the blade itself has two cutting edges, which we will call side A and side B.
+Let us consider a wire-cutting tool with a 1.5 cm razor blade that meets a cast surface (one such tool is shown in \Cref{fig:tool-pic}); the wire is held against this rectangular cast surface as the blade is pushed into the wire, cutting it in two.
+This is a minimal scenario -- the wire will acquire striations from one side of the blade, while the blade itself has two cutting edges, which we will call side A and side B.
 The blade cuts of a sheet of aluminum that is 1.5 cm in width thus produce two striated edges corresponding to side A and side B which can be compared to other cuts to assess similarity.

### 2023-12-14 PNAS/index.Rnw
resolution in microns per pixel, independence

@@
 
-\firstpage[8]{6}
+\firstpage[8]{5}
 % Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting.
@@
 
-If the blade cuts and the wire edge are scanned at a resolution of $\Sexpr{res*1000} \mu m$, the resulting digital files will require \svp{approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons per blade cut; $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall}.
+If the blade cuts and the wire edge are scanned at a lateral resolution of $\Sexpr{res*1000} \mu m$ per pixel, the resulting digital files will require \svp{approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons per blade cut; $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall}.
 These comparisons are required in order to find the optimal alignment between the wire and the blade cut; they are implicit in the calculation of cross-correlation which is the first \ac{and often} the only step used to objectively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.

### 2023-12-15 PNAS/index.Rnw
Updates

--- 2023-12-06 PNAS/index.Rnw
@@
 
-\firstpage[8]{6}
+\firstpage[3]{6}
 % Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting.
@@
 Yet, this potentially significant issue is often ignored.
-\ac{Now that vast databases and efficient algorithms are routinely used in forensic evaluations to find "matches" to crime scene items, the problems of close non-matches\citep{pcast}, and of the erosion of confidence due to multiple comparisons, become critically important.}
+\ac{Now that vast databases and efficient algorithms are routinely used in forensic evaluations to find "matches" to crime scene items, the problems of close non-matches\citep{pcast} and the erosion of confidence due to multiple comparisons become critically important.}
 %Researchers have been focusing on building up databases to develop algorithms for objective evidence comparisons.

### 2024-01-05 PNAS/index.Rnw
Remove coloration

@@
 
-A forensics examiner tasked with determining whether a wire in evidence was cut by a specific tool will create one or more blade cuts \ac{using the putative tool,} which can be compared to the cut surface of the wire.
+A forensics examiner tasked with determining whether a wire in evidence was cut by a specific tool will create one or more blade cuts using the putative tool, which can be compared to the cut surface of the wire.
 These cuts are made in a sheet of material similar to the wire's composition, and may be performed at multiple angles, as the angle of the tool to the substrate can affect which striations are recorded on the substrate surface.
@@
 Without loss of generality, we assume that the wire is fully covered with striations suitable for comparison; if this is not the case, we adjust the value of $d$.
-\svp{An illustration of the sliding comparison process is shown in \Cref{fig:wire-blade}.} Imagine that you start from the left end of the blade and move the wire surface \ac{$d$ units at a time} along the blade.
-At minimum, there must be $\ell/d$ comparisons in order to assess whether one side of the blade cut matches one side of the wire; in this case, the comparisons are independent of each other (conditional on striation patterns) because \ac{each comparison involves different surfaces.}
+An illustration of the sliding comparison process is shown in \Cref{fig:wire-blade}.
+Imagine that you start from the left end of the blade and move the wire surface $d$ units at a time along the blade.
+At minimum, there must be $\ell/d$ comparisons in order to assess whether one side of the blade cut matches one side of the wire; in this case, the comparisons are independent of each other (conditional on striation patterns) because each comparison involves different surfaces.
 %no overlapping data is used.
@@
 \caption{A comparison between a wire and a blade cut requires sliding the wire along the entire blade cut length to determine the best match (or whether there is a match).
-\svp{Surfaces shown are rendered 2D topographical scans of a wire and blade cut taken with a confocal light microscope.}}\label{fig:wire-blade}
+Surfaces shown are rendered 2D topographical scans of a wire and blade cut taken with a confocal light microscope.}\label{fig:wire-blade}
 \end{figure}
@@
 In this case, there may be as many as $\ell/r - d/r + 1$ comparisons performed in order to find the optimal alignment between the two samples.
-These comparisons are \ac{no longer independent}, as each sequential comparison shares much of the same data with the previous and next comparisons.
+These comparisons are no longer independent, as each sequential comparison shares much of the same data with the previous and next comparisons.
 % ; such dependency could be expressed via autocorrelation, but we will defer that derivation to a later paper.
-\svp{For simplicity}, let us consider the number of comparisons to lie somewhere between \svp{these two estimates}.
+For simplicity, let us consider the number of comparisons to lie somewhere between these two estimates.
 
-\ac{Finally}, we must consider the number of surfaces which must be compared: the wire may have between one and two sets of striae (depending on tool configuration) and there may be between 2 and 4 different blade cut surfaces to examine.
+Finally, we must consider the number of surfaces which must be compared: the wire may have between one and two sets of striae (depending on tool configuration) and there may be between 2 and 4 different blade cut surfaces to examine.
 This results in a multiplier of between 2 and 8 comparisons to be conducted to compare the striations on the wire to the striations on the blade cuts.
@@
 In this example, $\ell = \Sexpr{ell} mm$, $d = \Sexpr{2} mm$, and the minimal number of comparisons between the wire and one blade cut is $\ell/d = \Sexpr{ncomp_min}$.
-As there are two blade cuts (side A and side B), the minimal number of comparisons overall is \Sexpr{ncomp_min*2}, and these comparisons are \ac{non-overlapping and} independent.
+As there are two blade cuts (side A and side B), the minimal number of comparisons overall is \Sexpr{ncomp_min*2}, and these comparisons are non-overlapping and independent.
 
 If the blade cuts and the wire edge are scanned at a resolution of $\Sexpr{res*1000} \mu m$, the resulting digital files will require \svp{approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons per blade cut; $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall}.
-These comparisons are required in order to find the optimal alignment between the wire and the blade cut; they are implicit in the calculation of cross-correlation which is the first \ac{and often} the only step used to objectively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
+These comparisons are required in order to find the optimal alignment between the wire and the blade cut; they are implicit in the calculation of cross-correlation which is the first and often the only step used to objectively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
 This problem is not unique to algorithms; an examiner would need to physically align the wire and the blade cut by searching along the length of the cut to visually match striations, performing the same process physically that the algorithm performs computationally.
@@
 
-There are \ac{at least }two \ac{contributors} to a false positive rate: coincidental matches (that is, two pieces of evidence that happen to have similar characteristics but are not from the same source) and human/technical failures \citep[p 50]{pcast}.
+There are at least two contributors to a false positive rate: coincidental matches (that is, two pieces of evidence that happen to have similar characteristics but are not from the same source) and human/technical failures \citep[p 50]{pcast}.
 In objective disciplines with matching rules, such as DNA, these sources can be distinguished, and when coincidental match rates are very low, human/technical failures make up a large component of the error rate.
@@
 %Compounding this problem, o
-Open-set studies of striated comparisons (bullets and firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest \svp{that a starting point for estimating the} false positive rate for these comparisons is between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
+Open-set studies of striated comparisons (bullets and firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest that a starting point for estimating the false positive rate for these comparisons is between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
 
-If we begin by considering the multiple comparison problem relative to the $\ell/d$ independent comparisons between a blade cut and a wire, we can compute the overall false positive rate for all comparisons using $r$, the false positive rate for one comparison, and simple probability \ac{calculations}.
+If we begin by considering the multiple comparison problem relative to the $\ell/d$ independent comparisons between a blade cut and a wire, we can compute the overall false positive rate for all comparisons using $r$, the false positive rate for one comparison, and simple probability calculations.
 Let $$F_n:=\{\text{at least one false pos.
@@
 % Using this relationship between $F_n$ and the false positive rate $r$, we can determine what $r$ must be in order to limit $P(F_n)$ to an acceptable level by solving for $r$: $r = 1 - \left[1 - P(F_n)\right]^{1/n}$.
-\svp{\Cref{fig:Accuracy} shows $P(\overline{F_n})$ on the y-axis; if we set 90\% as the acceptable} \ac{examination-wide false positive rate (FPR)}, we can make 2 comparisons when the single-comparison FPR is 0.05 and 10 comparisons at 0.01.
-Thus, at an error rate of \Sexpr{sprintf("%.2f", fpr)}, we can conduct approximately 5 comparisons \ac{in the course of an examination without exceeding the examination-wide FPR of 10\%.}
+\Cref{fig:Accuracy} shows $P(\overline{F_n})$ on the y-axis; if we set 90\% as the acceptable examination-wide false positive rate (FPR), we can make 2 comparisons when the single-comparison FPR is 0.05 and 10 comparisons at 0.01.
+Thus, at an error rate of \Sexpr{sprintf("%.2f", fpr)}, we can conduct approximately 5 comparisons in the course of an examination without exceeding the examination-wide FPR of 10\%.
 
-\svp{It is clear that under these constraints, the} accuracy of the \ac{forensic evaluation that involves multiple comparisons} between a wire and a tool \ac{is likely to be low}, as the number of candidate alignments that must be examined is too high.
+It is clear that under these constraints, the accuracy of the forensic evaluation that involves multiple comparisons between a wire and a tool is likely to be low, as the number of candidate alignments that must be examined is too high.
 Even the most innocuous example we could come up with (small blade, only 2 cutting surfaces, and a relatively large wire) involved a minimum of \Sexpr{2*ncomp_min} completely independent comparisons.
-As a result, \ac{findings from}  wire comparisons made under current protocols (using error rates recently published in the forensics literature) \ac{are probably not} reliable enough to be presented at trial.
+As a result, findings from  wire comparisons made under current protocols (using error rates recently published in the forensics literature) are probably not reliable enough to be presented at trial.
 When we then consider that examiners would make cuts at multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons even in the simple case we have presented here, the probability of a false positive error increases further still.
@@
 
-\ac{Forensic practitioners often report the findings from their examinations in the form of a categorical conclusion reflecting a single decision.
-This can be misleading when the one decision relies on multiple comparisons which are not individually presented in reports or testimony.} In this short contribution, we have shown that the implicit comparisons performed during the forensic analysis of wire cuts have the potential to vastly increase the overall chance that at least one false positive error occurs.
+Forensic practitioners often report the findings from their examinations in the form of a categorical conclusion reflecting a single decision.
+This can be misleading when the one decision relies on multiple comparisons which are not individually presented in reports or testimony.
+In this short contribution, we have shown that the implicit comparisons performed during the forensic analysis of wire cuts have the potential to vastly increase the overall chance that at least one false positive error occurs.
 %While often a forensic conclusion is presented as a single decision, underlying that decision are many more comparisons which are not discussed or reported upon.
-\ac{Omission of the underlying} details is not necessarily \ac{due to} malice; rather, it results from protocols which focus on finding the best possible match between two items and then evaluating that match, \ac{and also from lack of familiarity with basic statistical ideas.}
+\ac{Omission of the underlying} details is not necessarily \ac{due to} malice; rather, it results from protocols which focus on finding the best possible match between two items and then evaluating that match, and also from lack of familiarity with basic statistical ideas.
 
-In our example \ac{we highlight a simple} case where a wire \ac{ is cut using a two-sided blade}, but the findings apply to many other situations where \ac{a forensic evaluation involves} multiple comparisons against a set of accumulated information.
-\ac{Examples include a database search or a situation similar to the example where multiple comparisons of evidence against reference samples are conducted.}
+In our example we highlight a simple} case where a wire  is cut using a two-sided blade, but the findings apply to many other situations where a forensic evaluation involves multiple comparisons against a set of accumulated information.
+Examples include a database search or a situation similar to the example where multiple comparisons of evidence against reference samples are conducted.
 %either in a database or in a series of test samples created during the examination process.
-As a result, it is critically important that forensic \ac{practitioners become aware of the impact of carrying out multiple comparisons on their final conclusions}
+As a result, it is critically important that forensic practitioners become aware of the impact of carrying out multiple comparisons on their final conclusions
 %science grapple with this multiple comparison problem and
 and that they modify lab procedures and testimony accordingly.
-\ac{We suggest three strategies to improve transparency of forensic examinations and to produce more reliable estimates of examination-wide error rates.}
+We suggest three strategies to improve transparency of forensic examinations and to produce more reliable estimates of examination-wide error rates.
 %We have three primary suggestions for improvements targeted at making the multiple comparisons visible to everyone involved in the process of evidence evaluation.

### 2024-01-05 PNAS/index.Rnw
Updates and fix typo

@@
 
-\ac{When wires are cut, the tool produces striations on the cut surfaced}; as in other forms of toolmark analysis, these striae are used to connect the wire to the tool that was used \ac{in the cutting process.
-Here, we argue that the practice} of comparing two wire cut surfaces introduces complexities not present in better-investigated forensic examination  of toolmars such as those observed on bullets, as wire comparisons inherently require multiple distinct comparisons, increasing the expected false positive rate.
+When wires are cut, the tool produces striations on the cut surfaced; as in other forms of toolmark analysis, these striae are used to connect the wire to the tool that was used in the cutting process.
+Here, we argue that the practice of comparing two wire cut surfaces introduces complexities not present in better-investigated forensic examination  of toolmarks such as those observed on bullets, as wire comparisons inherently require multiple distinct comparisons, increasing the expected false positive rate.
 We call attention to the multiple comparison problem in wire examination and relate it to other situations in forensics that involve multiple comparisons, such as database searches.

### 2024-01-05 PNAS/index.Rnw
Remove more colored text

@@
 
-\dropcap{I}n forensic \ac{evaluations}, a single conclusion often \ac{relies} on  many comparisons, either implicitly or explicitly.
+\dropcap{I}n forensic evaluations, a single conclusion often relies on  many comparisons, either implicitly or explicitly.
 Multiple comparisons arise persistently when developing statistical methods to address scientific problems \citep{benjaminiControllingFalseDiscovery1995}, and can significantly increase the probability of false "discoveries".
 Yet, this potentially significant issue is often ignored.
-\ac{Now that vast databases and efficient algorithms are routinely used in forensic evaluations to find "matches" to crime scene items, the problems of close non-matches\citep{pcast} and the erosion of confidence due to multiple comparisons become critically important.}
+Now that vast databases and efficient algorithms are routinely used in forensic evaluations to find "matches" to crime scene items, the problems of close non-matches\citep{pcast} and the erosion of confidence due to multiple comparisons become critically important.
 %Researchers have been focusing on building up databases to develop algorithms for objective evidence comparisons.
@@
 A similar concern was raised in the past with regard to DNA \citep{thompson2003} and latent print evaluations  \citep{koehler2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid train bombing case was that the large size of the IAFIS database used to search for similar prints made it possible to locate `unusually' close non-matches.
-\ac{The probability of finding  close non-matches for an item of evidence is directly proportional to the size of the database available for searching;} false "matches" occur more frequently as databases increase in size and algorithms become more powerful.
+The probability of finding  close non-matches for an item of evidence is directly proportional to the size of the database available for searching; false "matches" occur more frequently as databases increase in size and algorithms become more powerful.
 % In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.
@@
 
-\svp{Compounding this issue,} the use of novel algorithms typically results in a large number of comparisons that may not obvious to the user.
-For example, the cross-correlation function\citep{vorburgerApplicationsCrosscorrelationFunctions2011} was one of the first measures proposed for quantifying the similarity  between two patterns in response to the \ac{ 2009 NRC} report \citep{nas2009}, and continues to be used in many pattern searching algorithms to find the best alignment between two images and also to quantify their overall similarity.
+Compounding this issue, the use of novel algorithms typically results in a large number of comparisons that may not obvious to the user.
+For example, the cross-correlation function\citep{vorburgerApplicationsCrosscorrelationFunctions2011} was one of the first measures proposed for quantifying the similarity  between two patterns in response to the 2009 NRC report \citep{nas2009}, and continues to be used in many pattern searching algorithms to find the best alignment between two images and also to quantify their overall similarity.
 Finding the best alignment often consists in sliding one surfaceacross the whole length (for one-dimensional patterns, such as striations) or area (for two dimensional sources, such as impression marks) of the other item while keeping track of the value of a similarity measure.
 This mirrors the forensic examination process, where examiners visually rotate or align two surfaces under comparison microscopes.
-In order to avoid \ac{problematic} false accusations and a further erosion of the public's trust in science, it is important to address the problem of multiple comparisons by controlling their effect on false discovery rates.
+In order to avoid problematic false accusations and a further erosion of the public's trust in science, it is important to address the problem of multiple comparisons by controlling their effect on false discovery rates.
 
@@
 
-If the blade cuts and the wire edge are scanned at a resolution of $\Sexpr{res*1000} \mu m$, the resulting digital files will require \svp{approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons per blade cut; $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall}.
+If the blade cuts and the wire edge are scanned at a resolution of $\Sexpr{res*1000} \mu m$, the resulting digital files will require approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons per blade cut; $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall.
 These comparisons are required in order to find the optimal alignment between the wire and the blade cut; they are implicit in the calculation of cross-correlation which is the first and often the only step used to objectively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
@@
 library(dplyr)
-library(latex2exp)
+# library(latex2exp)
 library(geomtextpath)
@@
 %While often a forensic conclusion is presented as a single decision, underlying that decision are many more comparisons which are not discussed or reported upon.
-\ac{Omission of the underlying} details is not necessarily \ac{due to} malice; rather, it results from protocols which focus on finding the best possible match between two items and then evaluating that match, and also from lack of familiarity with basic statistical ideas.
+Omission of the underlying details is not necessarily due to malice; rather, it results from protocols which focus on finding the best possible match between two items and then evaluating that match, and also from lack of familiarity with basic statistical ideas.
 
-In our example we highlight a simple} case where a wire  is cut using a two-sided blade, but the findings apply to many other situations where a forensic evaluation involves multiple comparisons against a set of accumulated information.
+In our example we highlight a simple case where a wire  is cut using a two-sided blade, but the findings apply to many other situations where a forensic evaluation involves multiple comparisons against a set of accumulated information.
 Examples include a database search or a situation similar to the example where multiple comparisons of evidence against reference samples are conducted.

### 2024-01-05 PNAS/index.Rnw
abstract

@@
 
-When wires are cut, the tool produces striations on the cut surfaced; as in other forms of toolmark analysis, these striae are used to connect the wire to the tool that was used in the cutting process.
-Here, we argue that the practice of comparing two wire cut surfaces introduces complexities not present in better-investigated forensic examination  of toolmarks such as those observed on bullets, as wire comparisons inherently require multiple distinct comparisons, increasing the expected false positive rate.
+When wires are cut, the tool produces striations on the cut surfaced; as in other forms of forensic analysis, these striation marks are used to connect the evidence to the source that created them.
+Here, we argue that the practice of comparing two wire cut surfaces introduces complexities not present in better-investigated forensic examination  of toolmars such as those observed on bullets, as wire comparisons inherently require multiple distinct comparisons, increasing the expected false positive rate.
 We call attention to the multiple comparison problem in wire examination and relate it to other situations in forensics that involve multiple comparisons, such as database searches.
@@
 
-\firstpage[3]{6}
+\firstpage[8]{5}
 % Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting.
@@
 Yet, this potentially significant issue is often ignored.
-Now that vast databases and efficient algorithms are routinely used in forensic evaluations to find "matches" to crime scene items, the problems of close non-matches\citep{pcast} and the erosion of confidence due to multiple comparisons become critically important.
+Now that vast databases and efficient algorithms are routinely used in forensic evaluations to find "matches" to crime scene items, the problems of close non-matches \citep{pcast}, and of the erosion of confidence due to multiple comparisons, become critically important.
 %Researchers have been focusing on building up databases to develop algorithms for objective evidence comparisons.
@@
 
-A similar concern was raised in the past with regard to DNA \citep{thompson2003} and latent print evaluations  \citep{koehler2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid train bombing case was that the large size of the IAFIS database used to search for similar prints made it possible to locate `unusually' close non-matches.
-The probability of finding  close non-matches for an item of evidence is directly proportional to the size of the database available for searching; false "matches" occur more frequently as databases increase in size and algorithms become more powerful.
+These concerns were raised in the past with regard to DNA \citep{thompson2003} and latent print evaluations  \citep{koehler2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid train bombing case was that the large size of the IAFIS database used to search for similar prints made it possible to locate `unusually' close non-matches.
+The probability of finding  close non-matches for an item of evidence is directly linked to the size of the database available for searching: false "matches" occur more frequently as databases increase in size and algorithms are more powerful.
 % In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.
@@
 
-Compounding this issue, the use of novel algorithms typically results in a large number of comparisons that may not obvious to the user.
-For example, the cross-correlation function\citep{vorburgerApplicationsCrosscorrelationFunctions2011} was one of the first measures proposed for quantifying the similarity  between two patterns in response to the 2009 NRC report \citep{nas2009}, and continues to be used in many pattern searching algorithms to find the best alignment between two images and also to quantify their overall similarity.
-Finding the best alignment often consists in sliding one surfaceacross the whole length (for one-dimensional patterns, such as striations) or area (for two dimensional sources, such as impression marks) of the other item while keeping track of the value of a similarity measure.
-This mirrors the forensic examination process, where examiners visually rotate or align two surfaces under comparison microscopes.
+Compounding this issue, algorithmic results are typically based on a large number of comparisons that may not obvious to the user.
+For example, a typical task in the forensic examination process is to find the best alignment between two items of evidence.
+For one-dimensional patterns, such as striation marks, this consists of sliding one item along the full length of the other item.
+For two-dimensional patterns, such as impression marks on an area, an additional rotation is involved.
+In practice, examiners visually rotate or align two surfaces under comparison microscopes.
+Algorithmically this visual search is mirrored by finding a maximum of the (area) cross-correlation function.
+The cross-correlation function was one of the first measures proposed \citep{vorburgerApplicationsCrosscorrelationFunctions2011} for quantifying the similarity  between two patterns in response to the  2009 NRC report \citep{nas2009}, and continues to be used in many pattern searching algorithms to find the best alignment between two images and  quantify their overall similarity.
+
 In order to avoid problematic false accusations and a further erosion of the public's trust in science, it is important to address the problem of multiple comparisons by controlling their effect on false discovery rates.
@@
 
-A forensics examiner tasked with determining whether a wire in evidence was cut by a specific tool will create one or more blade cuts using the putative tool, which can be compared to the cut surface of the wire.
+A forensics examiner tasked with determining whether a wire in evidence was cut by a specific tool will create one or more blade cuts using the putative tool which can be compared to the cut surface of the wire.
+
 These cuts are made in a sheet of material similar to the wire's composition, and may be performed at multiple angles, as the angle of the tool to the substrate can affect which striations are recorded on the substrate surface.
@@
 In order to calculate the number of possible comparisons to be carried out in the course of the examination, we define $\ell$ to be the length of the blade cut, and $d$ to be the diameter of the wire.
-Without loss of generality, we assume that the wire is fully covered with striations suitable for comparison; if this is not the case, we adjust the value of $d$.
+Here, we assume that striations  suitable for comparison are visible across the full width $d$ of the wire; if this is not the case, we adjust the value of $d$.
 An illustration of the sliding comparison process is shown in \Cref{fig:wire-blade}.
-Imagine that you start from the left end of the blade and move the wire surface $d$ units at a time along the blade.
-At minimum, there must be $\ell/d$ comparisons in order to assess whether one side of the blade cut matches one side of the wire; in this case, the comparisons are independent of each other (conditional on striation patterns) because each comparison involves different surfaces.
+Imagine that a comparison starts from the left end of the blade and moves the wire surface $d$ units at a time along the blade.
+At minimum, there must be $\ell/d$ comparisons in order to assess whether one side of the blade cut matches one side of the wire.
+We can generally assume that these comparisons are independent  because each comparison involves different surfaces.
+\hh{We need to further assume that the length $d$ is wide enough to cover some striations; this ensures that the width of $d$ is wider than any significant autocorrelation  observed for the striations on the blade}.
 %no overlapping data is used.
-Of course, this overlooks the possibility of partial matches and assumes perfect alignment between the diameter of the wire and the portion of the blade used to cut the wire, which is unrealistic.
+\hh{Note, that this is a rather crude assessment of the bare minimal number of comparisons.
+}% Of course, this overlooks the possibility of partial matches and assumes perfect alignment between the diameter of the wire and the portion of the blade used to cut the wire, which is unrealistic.
 
@@
 
-At a maximum, both the blade and the wire are either digitally scanned at resolution $r$ mm, or visually examined using a microscope with a digital resolution that can be expressed as $r$ equivalent to the digital scan.
+At a maximum, both the blade and the wire are either digitally scanned at resolution of $r$ pixel per mm, or visually examined using a microscope with a digital resolution that can be expressed as equivalent to the digital scan.
 In this case, there may be as many as $\ell/r - d/r + 1$ comparisons performed in order to find the optimal alignment between the two samples.
@@
 In this example, $\ell = \Sexpr{ell} mm$, $d = \Sexpr{2} mm$, and the minimal number of comparisons between the wire and one blade cut is $\ell/d = \Sexpr{ncomp_min}$.
-As there are two blade cuts (side A and side B), the minimal number of comparisons overall is \Sexpr{ncomp_min*2}, and these comparisons are non-overlapping and independent.
+As there are two blade cuts (side A and side B), the minimal number of comparisons overall is \Sexpr{ncomp_min*2}, and these comparisons are \ac{non-overlapping and} independent.
 
-If the blade cuts and the wire edge are scanned at a resolution of $\Sexpr{res*1000} \mu m$, the resulting digital files will require approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons per blade cut; $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall.
-These comparisons are required in order to find the optimal alignment between the wire and the blade cut; they are implicit in the calculation of cross-correlation which is the first and often the only step used to objectively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
+If the blade cuts and the wire edge are scanned at a lateral resolution of $\Sexpr{res*1000} \mu m$ per pixel, the resulting digital files will require \svp{approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons per blade cut; $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall}.
+These comparisons are required in order to find the optimal alignment between the wire and the blade cut; they are implicit in the calculation of cross-correlation which is the first \ac{and often} the only step used to objectively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
 This problem is not unique to algorithms; an examiner would need to physically align the wire and the blade cut by searching along the length of the cut to visually match striations, performing the same process physically that the algorithm performs computationally.
@@
 
-There are at least two contributors to a false positive rate: coincidental matches (that is, two pieces of evidence that happen to have similar characteristics but are not from the same source) and human/technical failures \citep[p 50]{pcast}.
+There are \ac{at least }two \ac{contributors} to a false positive rate: coincidental matches (that is, two pieces of evidence that happen to have similar characteristics but are not from the same source) and human/technical failures \citep[p 50]{pcast}.
 In objective disciplines with matching rules, such as DNA, these sources can be distinguished, and when coincidental match rates are very low, human/technical failures make up a large component of the error rate.
@@
 %Compounding this problem, o
-Open-set studies of striated comparisons (bullets and firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest that a starting point for estimating the false positive rate for these comparisons is between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
+Open-set studies of striated comparisons (bullets and firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest \svp{that a starting point for estimating the} false positive rate for these comparisons is between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
 
-If we begin by considering the multiple comparison problem relative to the $\ell/d$ independent comparisons between a blade cut and a wire, we can compute the overall false positive rate for all comparisons using $r$, the false positive rate for one comparison, and simple probability calculations.
+If we begin by considering the multiple comparison problem relative to the $\ell/d$ independent comparisons between a blade cut and a wire, we can compute the overall false positive rate for all comparisons using $r$, the false positive rate for one comparison, and simple probability \ac{calculations}.
 Let $$F_n:=\{\text{at least one false pos.
@@
 % Using this relationship between $F_n$ and the false positive rate $r$, we can determine what $r$ must be in order to limit $P(F_n)$ to an acceptable level by solving for $r$: $r = 1 - \left[1 - P(F_n)\right]^{1/n}$.
-\Cref{fig:Accuracy} shows $P(\overline{F_n})$ on the y-axis; if we set 90\% as the acceptable examination-wide false positive rate (FPR), we can make 2 comparisons when the single-comparison FPR is 0.05 and 10 comparisons at 0.01.
-Thus, at an error rate of \Sexpr{sprintf("%.2f", fpr)}, we can conduct approximately 5 comparisons in the course of an examination without exceeding the examination-wide FPR of 10\%.
+\svp{\Cref{fig:Accuracy} shows $P(\overline{F_n})$ on the y-axis; if we set 90\% as the acceptable} \ac{examination-wide false positive rate (FPR)}, we can make 2 comparisons when the single-comparison FPR is 0.05 and 10 comparisons at 0.01.
+Thus, at an error rate of \Sexpr{sprintf("%.2f", fpr)}, we can conduct approximately 5 comparisons \ac{in the course of an examination without exceeding the examination-wide FPR of 10\%.}
 
-It is clear that under these constraints, the accuracy of the forensic evaluation that involves multiple comparisons between a wire and a tool is likely to be low, as the number of candidate alignments that must be examined is too high.
+\svp{It is clear that under these constraints, the} accuracy of the \ac{forensic evaluation that involves multiple comparisons} between a wire and a tool \ac{is likely to be low}, as the number of candidate alignments that must be examined is too high.
 Even the most innocuous example we could come up with (small blade, only 2 cutting surfaces, and a relatively large wire) involved a minimum of \Sexpr{2*ncomp_min} completely independent comparisons.
-As a result, findings from  wire comparisons made under current protocols (using error rates recently published in the forensics literature) are probably not reliable enough to be presented at trial.
+As a result, \ac{findings from}  wire comparisons made under current protocols (using error rates recently published in the forensics literature) \ac{are probably not} reliable enough to be presented at trial.
 When we then consider that examiners would make cuts at multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons even in the simple case we have presented here, the probability of a false positive error increases further still.
@@
 library(dplyr)
-# library(latex2exp)
+library(latex2exp)
 library(geomtextpath)
@@
 
-Forensic practitioners often report the findings from their examinations in the form of a categorical conclusion reflecting a single decision.
-This can be misleading when the one decision relies on multiple comparisons which are not individually presented in reports or testimony.
-In this short contribution, we have shown that the implicit comparisons performed during the forensic analysis of wire cuts have the potential to vastly increase the overall chance that at least one false positive error occurs.
+\ac{Forensic practitioners often report the findings from their examinations in the form of a categorical conclusion reflecting a single decision.
+This can be misleading when the one decision relies on multiple comparisons which are not individually presented in reports or testimony.} In this short contribution, we have shown that the implicit comparisons performed during the forensic analysis of wire cuts have the potential to vastly increase the overall chance that at least one false positive error occurs.
 %While often a forensic conclusion is presented as a single decision, underlying that decision are many more comparisons which are not discussed or reported upon.
-Omission of the underlying details is not necessarily due to malice; rather, it results from protocols which focus on finding the best possible match between two items and then evaluating that match, and also from lack of familiarity with basic statistical ideas.
+\ac{Omission of the underlying} details is not necessarily \ac{due to} malice; rather, it results from protocols which focus on finding the best possible match between two items and then evaluating that match, \ac{and also from lack of familiarity with basic statistical ideas.}
 
-In our example we highlight a simple case where a wire  is cut using a two-sided blade, but the findings apply to many other situations where a forensic evaluation involves multiple comparisons against a set of accumulated information.
-Examples include a database search or a situation similar to the example where multiple comparisons of evidence against reference samples are conducted.
+In our example \ac{we highlight a simple} case where a wire \ac{ is cut using a two-sided blade}, but the findings apply to many other situations where \ac{a forensic evaluation involves} multiple comparisons against a set of accumulated information.
+\ac{Examples include a database search or a situation similar to the example where multiple comparisons of evidence against reference samples are conducted.}
 %either in a database or in a series of test samples created during the examination process.
-As a result, it is critically important that forensic practitioners become aware of the impact of carrying out multiple comparisons on their final conclusions
+As a result, it is critically important that forensic \ac{practitioners become aware of the impact of carrying out multiple comparisons on their final conclusions}
 %science grapple with this multiple comparison problem and
 and that they modify lab procedures and testimony accordingly.
-We suggest three strategies to improve transparency of forensic examinations and to produce more reliable estimates of examination-wide error rates.
+\ac{We suggest three strategies to improve transparency of forensic examinations and to produce more reliable estimates of examination-wide error rates.}
 %We have three primary suggestions for improvements targeted at making the multiple comparisons visible to everyone involved in the process of evidence evaluation.

### 2024-01-09 PNAS/index.Rnw

--- 2024-01-05 PNAS/index.Rnw
@@
 \begin{abstract}
-
-When wires are cut, the tool produces striations on the cut surfaced; as in other forms of toolmark analysis, these striae are used to connect the wire to the tool that was used in the cutting process.
-Here, we argue that the practice of comparing two wire cut surfaces introduces complexities not present in better-investigated forensic examination  of toolmarks such as those observed on bullets, as wire comparisons inherently require multiple distinct comparisons, increasing the expected false positive rate.
+When wires are cut, the tool produces striations on the cut surfaced; as in other forms of forensic analysis, these striation marks are used to connect the evidence to the source that created them.
+Here, we argue that the practice of comparing two wire cut surfaces introduces complexities not present in better-investigated forensic examination  of toolmars such as those observed on bullets, as wire comparisons inherently require multiple distinct comparisons, increasing the expected false positive rate.
 We call attention to the multiple comparison problem in wire examination and relate it to other situations in forensics that involve multiple comparisons, such as database searches.
@@
 \dropcap{I}n forensic evaluations, a single conclusion often relies on  many comparisons, either implicitly or explicitly.
-Multiple comparisons arise persistently when developing statistical methods to address scientific problems \citep{benjaminiControllingFalseDiscovery1995}, and can significantly increase the probability of false "discoveries".
-Yet, this potentially significant issue is often ignored.
-Now that vast databases and efficient algorithms are routinely used in forensic evaluations to find "matches" to crime scene items, the problems of close non-matches\citep{pcast} and the erosion of confidence due to multiple comparisons become critically important.
+Multiple comparisons arise persistently when developing statistical methods to address scientific problems \citep{benjaminiControllingFalseDiscovery1995}, and greatly increase the probability of false discoveries.
+%Yet, this potentially significant issue is often ignored.
+Now that vast databases and efficient algorithms are routinely used in forensic evaluations to propose matches to crime scene items, the problem of close non-matches~\citep{pcast}  due to multiple comparisons becomes critically important.
+Scientists have the responsibility to address this often ignored issue to prevent a further erosion of the public's confidence in the justice system.
 %Researchers have been focusing on building up databases to develop algorithms for objective evidence comparisons.
@@
 
-A similar concern was raised in the past with regard to DNA \citep{thompson2003} and latent print evaluations  \citep{koehler2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid train bombing case was that the large size of the IAFIS database used to search for similar prints made it possible to locate `unusually' close non-matches.
-The probability of finding  close non-matches for an item of evidence is directly proportional to the size of the database available for searching; false "matches" occur more frequently as databases increase in size and algorithms become more powerful.
+The multiple comparison problem is not new: it has been raised in the past with regard to DNA~\citep{thompson2003} and latent print evaluations~ \citep{koehler2021}.
+One of the root causes~\citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid train bombing case was that the large size of the IAFIS database used to search for similar prints made it possible to locate `unusually' close non-matches.
+The probability of finding  close non-matches for an item of evidence is proportional to the size of the database available for searching; close non-matches occur more frequently as databases increase in size and algorithms become more powerful.
 % In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.
@@
 Compounding this issue, the use of novel algorithms typically results in a large number of comparisons that may not obvious to the user.
-For example, the cross-correlation function\citep{vorburgerApplicationsCrosscorrelationFunctions2011} was one of the first measures proposed for quantifying the similarity  between two patterns in response to the 2009 NRC report \citep{nas2009}, and continues to be used in many pattern searching algorithms to find the best alignment between two images and also to quantify their overall similarity.
-Finding the best alignment often consists in sliding one surfaceacross the whole length (for one-dimensional patterns, such as striations) or area (for two dimensional sources, such as impression marks) of the other item while keeping track of the value of a similarity measure.
-This mirrors the forensic examination process, where examiners visually rotate or align two surfaces under comparison microscopes.
-In order to avoid problematic false accusations and a further erosion of the public's trust in science, it is important to address the problem of multiple comparisons by controlling their effect on false discovery rates.
+For example, the cross-correlation function~\citep{vorburgerApplicationsCrosscorrelationFunctions2011} was one of the first measures proposed for quantifying the similarity  between two patterns in response to the 2009 NRC report~\citep{nas2009}, and continues to be used in many pattern searching algorithms to find the best alignment between two images and to quantify their overall similarity.
+Finding the best alignment often consists in sliding one surface across the whole length (for one-dimensional patterns, such as striations) or area (for two dimensional sources, such as impression marks) of the other item while keeping track of the value of a similarity measure.
+This mirrors the forensic examination process: the examiner visually rotates and shifts items under a comparison microscope to align two surfaces.
+In order to avoid false accusations and a further erosion of the public's trust in science, it is important to address the problem of multiple comparisons by controlling their effect on false discovery rates.
 

### 2024-01-15 PNAS/index.Rnw
Updates to language about statistical independence

@@
 Imagine that you start from the left end of the blade and move the wire surface $d$ units at a time along the blade.
-At minimum, there must be $\ell/d$ comparisons in order to assess whether one side of the blade cut matches one side of the wire; in this case, the comparisons are independent of each other (conditional on striation patterns) because each comparison involves different surfaces.
+At minimum, there must be $\ell/d$ comparisons in order to assess whether one side of the blade cut matches one side of the wire; in this case, the comparisons are physically separate and thus expected to be statistically independent, though empirically there will be nonzero correlations due to the physical similarities between striations.
 %no overlapping data is used.

### 2024-01-19 PNAS/index.Rnw
line 106

@@
 %no overlapping data is used.
-Of course, this overlooks the possibility of partial matches and assumes perfect alignment between the diameter of the wire and the portion of the blade used to cut the wire, which is unrealistic.
+This method of counting the number of comparisons assumes
+perfect alignment between the diameter of the wire and the portion of the blade used to cut the wire, which is unrealistic but provides us with a lower bound.
 

### 2024-01-19 PNAS/index.Rnw
line 102

@@
 
-In order to calculate the number of possible comparisons to be carried out in the course of the examination, we define $\ell$ to be the length of the blade cut, and $d$ to be the diameter of the wire.
-Without loss of generality, we assume that the wire is fully covered with striations suitable for comparison; if this is not the case, we adjust the value of $d$.
+In order to calculate the number of comparisons carried out in the course of one examination, we define $\ell$ to be the length of the blade cut, and $d$ to be the diameter of the wire.
+We assume that the wire is covered with striations suitable for comparison across its full diameter $d$.
+If this is not the case, we reduce the value $d$ correspondingly.
 An illustration of the sliding comparison process is shown in \Cref{fig:wire-blade}.
 Imagine that you start from the left end of the blade and move the wire surface $d$ units at a time along the blade.
+
 At minimum, there must be $\ell/d$ comparisons in order to assess whether one side of the blade cut matches one side of the wire; in this case, the comparisons are physically separate and thus expected to be statistically independent, though empirically there will be nonzero correlations due to the physical similarities between striations.

### 2024-01-19 PNAS/index.Rnw
Rewording

@@
 We assume that the wire is covered with striations suitable for comparison across its full diameter $d$.
-If this is not the case, we reduce the value $d$ correspondingly.
+If this is not the case, we reduce the value $d$.
+Both the blade and the wire are either digitally scanned at resolution $r$ mm per pixel, or visually examined using a microscope with a digital resolution that can be expressed as $r$ equivalent to the digital scan.
 An illustration of the sliding comparison process is shown in \Cref{fig:wire-blade}.
-Imagine that you start from the left end of the blade and move the wire surface $d$ units at a time along the blade.
+Imagine that we move the cut wire along the blade cut in order to assess whether striations on the blade cut match the striations on the wire.
+We can move the wire unit-by-unit, or we can move the wire by its full length, with no overlap to the previous comparison.
 
-At minimum, there must be $\ell/d$ comparisons in order to assess whether one side of the blade cut matches one side of the wire; in this case, the comparisons are physically separate and thus expected to be statistically independent, though empirically there will be nonzero correlations due to the physical similarities between striations.
-%no overlapping data is used.
-This method of counting the number of comparisons assumes
-perfect alignment between the diameter of the wire and the portion of the blade used to cut the wire, which is unrealistic but provides us with a lower bound.
+The first option gives us the maximum number of comparisons ($\ell/r - d/r + 1$), while the second option gives us the minimum number of comparisons.
+In the first case, sequential comparisons share much of the same physical data and are highly related; in the second case, no data is shared between physical comparisons and we can expect that they are statistically independent, though empirically there will be nonzero correlations due to physical similarities between striations.
+For simplicity, let us consider the number of comparisons to lie somewhere between these two estimates.
 
@@
 
-At a maximum, both the blade and the wire are either digitally scanned at resolution $r$ mm, or visually examined using a microscope with a digital resolution that can be expressed as $r$ equivalent to the digital scan.
-In this case, there may be as many as $\ell/r - d/r + 1$ comparisons performed in order to find the optimal alignment between the two samples.
-These comparisons are no longer independent, as each sequential comparison shares much of the same data with the previous and next comparisons.
-% ; such dependency could be expressed via autocorrelation, but we will defer that derivation to a later paper.
-For simplicity, let us consider the number of comparisons to lie somewhere between these two estimates.
-
-Finally, we must consider the number of surfaces which must be compared: the wire may have between one and two sets of striae (depending on tool configuration) and there may be between 2 and 4 different blade cut surfaces to examine.
-This results in a multiplier of between 2 and 8 comparisons to be conducted to compare the striations on the wire to the striations on the blade cuts.
+Finally, we must consider the number of surfaces which must be compared: the wire may have between one and two sets of striae and there may be between 2 and 4 different blade cut surfaces to examine, depending on tool configuration.
+This results in a multiplier of between 2 and 8 to calculate the total number of comparisons required.
 

### 2024-01-19 PNAS/index.Rnw
Slight restructuring

@@
 This is a minimal scenario - the wire will acquire striations from one side of the blade, while the blade itself has two cutting edges, which we will call side A and side B.
-The blade cuts of a sheet of aluminum that is 1.5 cm in width thus produce two striated edges corresponding to side A and side B which can be compared to other cuts to assess similarity.
+A blade cut of a sheet of aluminum will thus produce two striated edges corresponding to side A and side B which can be compared to cut wires to assess similarity.
 
-In addition, we have a 12 gauge aluminum wire (2 mm diameter) which may have been cut  by the wire-cutting tool described above.
+We also have a 12 gauge aluminum wire (2 mm diameter) which may have been cut by the wire-cutting tool described above.
 Class characteristics, which are shared by all tools of similar manufacture, appear to match:
@@
 and the wire is cut such that the blade and the backstop appear to be perpendicular
-(that is, the wire appears to have been cut with a blade similar to the blade configuration of the tool in question).
+(that is, the wire appears to have been cut with a tool of similar configuration).
 
@@
 
-In this example, $\ell = \Sexpr{ell} mm$, $d = \Sexpr{2} mm$, and the minimal number of comparisons between the wire and one blade cut is $\ell/d = \Sexpr{ncomp_min}$.
-As there are two blade cuts (side A and side B), the minimal number of comparisons overall is \Sexpr{ncomp_min*2}, and these comparisons are non-overlapping and independent.
+In this example, $\ell = \Sexpr{ell}$ mm, $d = \Sexpr{2}$ mm, and the minimal number of comparisons between the wire and one blade cut is $\ell/d = \Sexpr{ncomp_min}$.
+As there are two blade cuts (side A and side B), the minimal number of comparisons overall is \Sexpr{ncomp_min*2}, as these comparisons are non-overlapping and independent (on average).
 
-If the blade cuts and the wire edge are scanned at a resolution of $\Sexpr{res*1000} \mu m$, the resulting digital files will require approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons per blade cut; $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall.
-These comparisons are required in order to find the optimal alignment between the wire and the blade cut; they are implicit in the calculation of cross-correlation which is the first and often the only step used to objectively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
-This problem is not unique to algorithms; an examiner would need to physically align the wire and the blade cut by searching along the length of the cut to visually match striations, performing the same process physically that the algorithm performs computationally.
-While these sequential comparisons are highly autocorrelated, and we cannot assume sequential independence when calculating the probability of an error, they serve as an upper bound on the number of comparisons which could be performed.
+Let's assume a resolution of $\Sexpr{res*1000} \mu$m per pixel, the maximum number of comparisons per blade cut is approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons; thus, we need $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall in order to find the optimal alignment between the wire and the blade cut.
+These comparisons are implicit in the calculation of cross-correlation, which is the first and often the only step used to quantitatively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
+Implicit comparisons are not unique to algorithms; an examiner would need to physically align the wire and the blade cut by searching along the length of the cut to visually match striations, performing the same process physically that the algorithm performs computationally.
+While these sequential comparisons are highly auto correlated, and we cannot assume sequential independence when calculating the probability of an error, they serve as an upper bound on the number of comparisons which could be performed.
 

### 2024-01-19 PNAS/index.Rnw
change r to e for error rate

@@
 
-If we begin by considering the multiple comparison problem relative to the $\ell/d$ independent comparisons between a blade cut and a wire, we can compute the overall false positive rate for all comparisons using $r$, the false positive rate for one comparison, and simple probability calculations.
+If we begin by considering the multiple comparison problem relative to the minimal $\ell/d$ independent comparisons between a blade cut and a wire, we can compute the overall false positive rate for all comparisons using $r$, the false positive rate for one comparison, and simple probability calculations.
 Let $$F_n:=\{\text{at least one false pos.
-in }n\text{ disjoint comparisons}\}.$$ Then, $P(F_1) = r$, $P(\overline{F_1}) = 1-r$, $P(F_n) = [1-r]^n$, and $P(\overline{F_n}) = 1 - \left[P(\overline{F_1})\right]^n$.
-% Using this relationship between $F_n$ and the false positive rate $r$, we can determine what $r$ must be in order to limit $P(F_n)$ to an acceptable level by solving for $r$: $r = 1 - \left[1 - P(F_n)\right]^{1/n}$.
+in }n\text{ disjoint comparisons}\}.$$ Then, $P(F_1) = e$, $P(\overline{F_1}) = 1-e$, $P(F_n) = [1-e]^n$, and $P(\overline{F_n}) = 1 - \left[P(\overline{F_1})\right]^n$.
+% Using this relationship between $F_n$ and the false positive rate $r$, we can determine what $e$ must be in order to limit $P(F_n)$ to an acceptable level by solving for $e$: $e = 1 - \left[1 - P(F_n)\right]^{1/n}$.
 \Cref{fig:Accuracy} shows $P(\overline{F_n})$ on the y-axis; if we set 90\% as the acceptable examination-wide false positive rate (FPR), we can make 2 comparisons when the single-comparison FPR is 0.05 and 10 comparisons at 0.01.
@@
 
-It is clear that under these constraints, the accuracy of the forensic evaluation that involves multiple comparisons between a wire and a tool is likely to be low, as the number of candidate alignments that must be examined is too high.
-Even the most innocuous example we could come up with (small blade, only 2 cutting surfaces, and a relatively large wire) involved a minimum of \Sexpr{2*ncomp_min} completely independent comparisons.
+It is clear that under these constraints, the accuracy of a forensic evaluation involving multiple comparisons between a wire and a tool is likely to be low, as the number of candidate alignments that must be examined is too high.
+Even the most innocuous example we could find (small blade, only 2 cutting surfaces, and a relatively large wire) involved a minimum of \Sexpr{2*ncomp_min} completely independent comparisons.
 As a result, findings from  wire comparisons made under current protocols (using error rates recently published in the forensics literature) are probably not reliable enough to be presented at trial.
@@
 
-<<control-overall-fpr, fig.cap="Exploration of overall error rate estimates for different values of $r$, the false positive rate for a single comparison.
+<<control-overall-fpr, fig.cap="Exploration of overall error rate estimates for different values of $e$, the false positive rate for a single comparison.
 Dotted vertical lines show false positive error rate estimates (from left to right, \\citep{best2022,bajic2020,mattijssen2021}).", fig.width = 8, fig.height = 4, include = F>>=

### 2024-01-19 PNAS/index.Rnw
missed one r

@@
 
-If we begin by considering the multiple comparison problem relative to the minimal $\ell/d$ independent comparisons between a blade cut and a wire, we can compute the overall false positive rate for all comparisons using $r$, the false positive rate for one comparison, and simple probability calculations.
+If we begin by considering the multiple comparison problem relative to the minimal $\ell/d$ independent comparisons between a blade cut and a wire, we can compute the overall false positive rate for all comparisons using $e$, the false positive rate for one comparison, and simple probability calculations.
 Let $$F_n:=\{\text{at least one false pos.

### 2024-01-19 PNAS/index.Rnw
Updates to language - passive voice

@@
 In objective disciplines with matching rules, such as DNA, these sources can be distinguished, and when coincidental match rates are very low, human/technical failures make up a large component of the error rate.
-However, in toolmark examination, we do not have an objective matching rule; examiners testify based on subjective, individual rules for how much similarity is enough for an identification.
-As a result, we cannot separate coincidental match rates from human/technical failures.
+However, in toolmark examination, no discipline-wide objective matching rule is used; examiners testify based on subjective, individual rules for how much similarity is enough for an identification.
+As a result, it is not possible to separate coincidental match rates from human/technical failures.
 <<fpr, include = F>>=
@@
 
-%Compounding this problem, o
 Open-set studies of striated comparisons (bullets and firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest that a starting point for estimating the false positive rate for these comparisons is between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
 
-If we begin by considering the multiple comparison problem relative to the minimal $\ell/d$ independent comparisons between a blade cut and a wire, we can compute the overall false positive rate for all comparisons using $e$, the false positive rate for one comparison, and simple probability calculations.
+If we begin by considering the minimal $\ell/d$ independent comparisons between a blade cut and a wire, we can compute the overall false positive rate for all comparisons using $e$, the false positive rate for one comparison, and simple probability calculations.
 Let $$F_n:=\{\text{at least one false pos.
@@
 % Using this relationship between $F_n$ and the false positive rate $r$, we can determine what $e$ must be in order to limit $P(F_n)$ to an acceptable level by solving for $e$: $e = 1 - \left[1 - P(F_n)\right]^{1/n}$.
-\Cref{fig:Accuracy} shows $P(\overline{F_n})$ on the y-axis; if we set 90\% as the acceptable examination-wide false positive rate (FPR), we can make 2 comparisons when the single-comparison FPR is 0.05 and 10 comparisons at 0.01.
-Thus, at an error rate of \Sexpr{sprintf("%.2f", fpr)}, we can conduct approximately 5 comparisons in the course of an examination without exceeding the examination-wide FPR of 10\%.
+\Cref{fig:Accuracy} shows $P(\overline{F_n})$ on the y-axis; if 90\% is an acceptable examination-wide false positive rate (FPR), 2 comparisons can be made when the single-comparison FPR is 0.05; this increases to 10 comparisons at 0.01.
+Thus, at an error rate of \Sexpr{sprintf("%.2f", fpr)}, examiners can conduct approximately 5 comparisons without exceeding the examination-wide FPR of 10\%.
 
@@
 Forensic practitioners often report the findings from their examinations in the form of a categorical conclusion reflecting a single decision.
-This can be misleading when the one decision relies on multiple comparisons which are not individually presented in reports or testimony.
+This can be misleading when this decision relies on multiple comparisons which are not individually presented in reports or testimony.
 In this short contribution, we have shown that the implicit comparisons performed during the forensic analysis of wire cuts have the potential to vastly increase the overall chance that at least one false positive error occurs.

### 2024-01-19 PNAS/index.Rnw
Rearranging the knitr chunks and adding sexpr to things that were previously static.

@@
 
+<<simple-problem-setup, include=F>>=
+ell <- 15
+d <- 2
+res <- 0.645/1000
+ncomp_min <- ell/d
+ncomp_max <- ell/res - d/res + 1
+
+# From 3 studies
+errors <- c(20, 56, 1)
+knm <- c(2842, 774, 220)
+
+fpr <- sum(errors)/sum(knm)
+@
+
 \begin{figure}[t!]
@@
 \includegraphics[width=\columnwidth]{../fig/wire_cutter_wires.png}
-\caption{RJ45 Crimp tool with a 1.5cm razor blade used for cutting. 1mm and 2mm diameter aluminum wires cut with the pliers are shown in a box in the top right corner.}\label{fig:tool-pic}
+\caption{RJ45 Crimp tool with a $\Sexpr{ell/10}$ cm razor blade used for cutting.
+1 mm and 2 mm diameter aluminum wires cut with the pliers are shown in a box in the top right corner.}\label{fig:tool-pic}
 \end{figure}
 
-Let us consider a wire-cutting tool with a 1.5 cm razor blade that meets a cast surface (one such tool is shown in \Cref{fig:tool-pic}); the wire is held against this rectangular cast surface as the blade is pushed into the wire, splitting it in two.
+Let us consider a wire-cutting tool with a $\Sexpr{ell/10}$ cm razor blade that meets a cast surface (one such tool is shown in \Cref{fig:tool-pic}); the wire is held against this rectangular cast surface as the blade is pushed into the wire, splitting it in two.
 This is a minimal scenario - the wire will acquire striations from one side of the blade, while the blade itself has two cutting edges, which we will call side A and side B.
 A blade cut of a sheet of aluminum will thus produce two striated edges corresponding to side A and side B which can be compared to cut wires to assess similarity.
-
-We also have a 12 gauge aluminum wire (2 mm diameter) which may have been cut by the wire-cutting tool described above.
+We also have a 12 gauge aluminum wire ($\Sexpr{d}$ mm diameter) which may have been cut by the wire-cutting tool described above.
 Class characteristics, which are shared by all tools of similar manufacture, appear to match:
@@
 
-<<simple-problem-setup, echo = F, warning = F, message = F>>=
-ell <- 15
-d <- 2
-res <- 0.645/1000
-ncomp_min <- ell/d
-ncomp_max <- ell/res - d/res + 1
-@
-
-In this example, $\ell = \Sexpr{ell}$ mm, $d = \Sexpr{2}$ mm, and the minimal number of comparisons between the wire and one blade cut is $\ell/d = \Sexpr{ncomp_min}$.
+In this example, $\ell = \Sexpr{ell}$ mm, $d = \Sexpr{d}$ mm, and at minimum there are $\ell/d = \Sexpr{ncomp_min}$ comparisons between a wire cut and a blade cut.
 As there are two blade cuts (side A and side B), the minimal number of comparisons overall is \Sexpr{ncomp_min*2}, as these comparisons are non-overlapping and independent (on average).
@@
 \section*{Probability of False Identifications}
-
-There are at least two contributors to a false positive rate: coincidental matches (that is, two pieces of evidence that happen to have similar characteristics but are not from the same source) and human/technical failures \citep[p 50]{pcast}.
-In objective disciplines with matching rules, such as DNA, these sources can be distinguished, and when coincidental match rates are very low, human/technical failures make up a large component of the error rate.
-However, in toolmark examination, no discipline-wide objective matching rule is used; examiners testify based on subjective, individual rules for how much similarity is enough for an identification.
-As a result, it is not possible to separate coincidental match rates from human/technical failures.
-<<fpr, include = F>>=
-errors <- c(20, 56, 1)
-knm <- c(2842, 774, 220)
-
-fpr <- sum(errors)/sum(knm)
-@
-
-Open-set studies of striated comparisons (bullets and firing pin impressions) \citep{bajic2020, mattijssen2021,best2022} suggest that a starting point for estimating the false positive rate for these comparisons is between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these three studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
+There are at least two components of the false positive rate: identifying two pieces of evidence that have similar characteristics but are from different sources (a coincidental match) and procedural failures (e.g.
+lab process errors) \citep[p 50]{pcast}.
+In objective disciplines with standardized evaluation rules (e.g. DNA), these sources can be distinguished.
+However, in toolmark examination, no objective evaluation rules are used; examiners testify based on subjective, individual rules for how much similarity is sufficient for an identification.
+Assuming that lab procedure errors are not a factor in studies, we use reported error rates from striated evidence comparisons (e.g.
+excluding breech face comparisons) in open-set studies as an estimate of the coincidental match rate of a single wire-cut comparison.
+%% Do we need to address the fact that wire cuts don't have external features to align, where bullets and firing pins do???
+The three studies we consider \citep{bajic2020, mattijssen2021,best2022} have false positive rates between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
 
@@
 
-It is clear that under these constraints, the accuracy of a forensic evaluation involving multiple comparisons between a wire and a tool is likely to be low, as the number of candidate alignments that must be examined is too high.
+It is clear that under these constraints, the accuracy of an examination involving multiple comparisons between a wire and a tool is likely to be low, as the number of candidate alignments that must be examined is too high.
 Even the most innocuous example we could find (small blade, only 2 cutting surfaces, and a relatively large wire) involved a minimum of \Sexpr{2*ncomp_min} completely independent comparisons.

### 2024-01-19 PNAS/index.Rnw
A few more edits to the conclusion

@@
 
-In our example we highlight a simple case where a wire  is cut using a two-sided blade, but the findings apply to many other situations where a forensic evaluation involves multiple comparisons against a set of accumulated information.
-Examples include a database search or a situation similar to the example where multiple comparisons of evidence against reference samples are conducted.
+In our example we highlight a simple case where a wire  is cut using a two-sided blade, but the findings apply to many other situations where a forensic evaluation involves multiple comparisons against a set of accumulated information, such as database searches.
 %either in a database or in a series of test samples created during the examination process.
-As a result, it is critically important that forensic practitioners become aware of the impact of carrying out multiple comparisons on their final conclusions
-%science grapple with this multiple comparison problem and
-and that they modify lab procedures and testimony accordingly.
+It is critically important that forensic practitioners become aware of the impact of carrying out multiple comparisons on their final conclusions and that lab procedures and testimony are modified accordingly.
 We suggest three strategies to improve transparency of forensic examinations and to produce more reliable estimates of examination-wide error rates.
-%We have three primary suggestions for improvements targeted at making the multiple comparisons visible to everyone involved in the process of evidence evaluation.
 
@@
 Examiners should also report the total consecutive length or area of the recovered evidence.
-These pieces of information will allow for calculations of minimum and maximum number of comparisons (as we have demonstrated in this paper), and will facilitate calculation of overall error rates.
+These pieces of information will allow for calculations of minimum and maximum number of comparisons (as demonstrated in this paper), facilitating calculation of overall error rates.
 
-Second, researchers in this field should conduct studies relating the length/area of comparison surface to the error rate.
-For instance, here, we have combined studies looking at bullet striations with studies on firing pin shear marks - the striated surfaces are of orders of magnitude different lengths, but they represent our best estimate of the error rate for striated comparisons.
-This was necessary, because we could not find black-box studies of wire cuts.
-These studies should be Type II studies as defined by Koehler \citep{koehlerIntuitiveErrorRate2017a} in that they should be designed to assess the actual error rate when examiners are making difficult comparisons.
+Second, researchers should conduct studies relating the length/area of comparison surface to the error rate.
+For instance, we have pooled studies looking at bullet striations  and firing pin shear marks.
+The striated surfaces are of orders of magnitude different lengths, but they represent our best estimate of the error rate for striated comparisons.
+This was necessary because we could not find black-box error rate studies of wire cuts.
+These studies should be Type II studies as defined by Koehler \citep{koehlerIntuitiveErrorRate2017a}, in that they should be designed to assess the actual error rate when examiners are making difficult comparisons.
 
-Finally, when databases are used at any stage of the forensic evidence evaluation process (from suitability assessment and triage to reports which will be used at trial), the number of database items searched and the number of results returned must be reported.
-Furthermore, the number of results used for futher manual comparison should also be reported.
+Finally, when databases are used at any stage of the forensic evidence evaluation process (from suitability assessment and triage to reports which will be used at trial), the number of database items searched (or comparisons made) and the number of results returned must be reported.
+Furthermore, the number of results used for further manual comparison should also be reported.
 For example, if a firearms examiner searches a local NIBIN database with 1000 entries, requests the 20 closest matches to her evidence, and then carries out a physical examination of five exemplars from the list of 20, all of those values should be clearly reported to enable calculation of the potential FPR.
-This will help make the multiple comparison issue visible for everyone involved in evaluating the forensic evidence, from examiners themselves to jurors and judges.
+This will help make the multiple comparison issue accessible to everyone involved in evaluating the value of forensic evidence, from examiners to lawyers to jurors and judges.
 

### 2024-01-19 PNAS/index.Rnw
Fix probability calcs
Notation update to E_n

@@
 If we begin by considering the minimal $\ell/d$ independent comparisons between a blade cut and a wire, we can compute the overall false positive rate for all comparisons using $e$, the false positive rate for one comparison, and simple probability calculations.
-Let $$F_n:=\{\text{at least one false pos.
-in }n\text{ disjoint comparisons}\}.$$ Then, $P(F_1) = e$, $P(\overline{F_1}) = 1-e$, $P(F_n) = [1-e]^n$, and $P(\overline{F_n}) = 1 - \left[P(\overline{F_1})\right]^n$.
-% Using this relationship between $F_n$ and the false positive rate $r$, we can determine what $e$ must be in order to limit $P(F_n)$ to an acceptable level by solving for $e$: $e = 1 - \left[1 - P(F_n)\right]^{1/n}$.
-\Cref{fig:Accuracy} shows $P(\overline{F_n})$ on the y-axis; if 90\% is an acceptable examination-wide false positive rate (FPR), 2 comparisons can be made when the single-comparison FPR is 0.05; this increases to 10 comparisons at 0.01.
+Let $$E_n:=P\left(\{\geq 1\text{ false pos. in }n\text{ disjoint comparisons}\}\right).$$
+Then, $E_1 = e$, $\overline{E_1} = 1-e$, $\overline{E_n} = \left[\overline{E_1}\right]^n$, and $E_n = 1 - \overline{E_n} = 1 - [1-e]^n$.
+% Using this relationship between $n$ and the false positive rate $e$, we can determine what $e$ must be in order to limit $E_n$ to an acceptable level by solving for $e$: $e = 1 - \left[1 - E_n\right]^{1/n}$.
+\Cref{fig:Accuracy} shows $E_n$ on the y-axis; if 90\% is an acceptable examination-wide false positive rate (FPR), 2 comparisons can be made when the single-comparison FPR is 0.05; this increases to 10 comparisons at 0.01.
 Thus, at an error rate of \Sexpr{sprintf("%.2f", fpr)}, examiners can conduct approximately 5 comparisons without exceeding the examination-wide FPR of 10\%.

### 2024-01-19 PNAS/index.Rnw
dot after e.g. should not be interpreted as the end of a sentence

--- 2024-01-19 PNAS/index.Rnw
@@
 The value {12} represents the number of the paragraph starting the second page.
-If a paragraph runs over onto the second page, include a bracket with the paragraph line number starting the second page, followed by the paragraph number in curly brackets, e.g.
-"\firstpage[4]{11}".
+If a paragraph runs over onto the second page, include a bracket with the paragraph line number starting the second page, followed by the paragraph number in curly brackets, e.g.\ "\firstpage[4]{11}".
 
@@
 \section*{Probability of False Identifications}
-There are at least two components of the false positive rate: identifying two pieces of evidence that have similar characteristics but are from different sources (a coincidental match) and procedural failures (e.g.
-lab process errors) \citep[p 50]{pcast}.
-In objective disciplines with standardized evaluation rules (e.g. DNA), these sources can be distinguished.
+There are at least two components of the false positive rate: identifying two pieces of evidence that have similar characteristics but are from different sources (a coincidental match) and procedural failures (e.g.\ lab process errors) \citep[p 50]{pcast}.
+In objective disciplines with standardized evaluation rules (e.g.\ DNA), these sources can be distinguished.
 However, in toolmark examination, no objective evaluation rules are used; examiners testify based on subjective, individual rules for how much similarity is sufficient for an identification.
-Assuming that lab procedure errors are not a factor in studies, we use reported error rates from striated evidence comparisons (e.g.
-excluding breech face comparisons) in open-set studies as an estimate of the coincidental match rate of a single wire-cut comparison.
+Assuming that lab procedure errors are not a factor in studies, we use reported error rates from striated evidence comparisons (e.g.\ excluding breech face comparisons) in open-set studies as an estimate of the coincidental match rate of a single wire-cut comparison.
 %% Do we need to address the fact that wire cuts don't have external features to align, where bullets and firing pins do???
@@
 %
-% References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g.
-\cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric,phdthesis,masterthesis}.
+% References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g.\ \cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric,phdthesis,masterthesis}.
 All references cited in the main text should be included in the main manuscript file.

### 2024-01-19 PNAS/index.Rnw
Tweaking conclusions, removing can be phrasing

--- 2024-01-19 PNAS/index.Rnw
@@
 
-A forensics examiner tasked with determining whether a wire in evidence was cut by a specific tool will create one or more blade cuts using the putative tool, which can be compared to the cut surface of the wire.
+A forensics examiner tasked with determining whether a wire in evidence was cut by a specific tool will create one or more blade cuts using the putative tool, which are then compared to the cut surface of the wire.
 These cuts are made in a sheet of material similar to the wire's composition, and may be performed at multiple angles, as the angle of the tool to the substrate can affect which striations are recorded on the substrate surface.
@@
 
-The first option gives us the maximum number of comparisons ($\ell/r - d/r + 1$), while the second option gives us the minimum number of comparisons.
+The first option gives us the maximum number of comparisons ($\ell/r - d/r + 1$), while the second option gives us the minimum number of comparisons $\ell/d$.
 In the first case, sequential comparisons share much of the same physical data and are highly related; in the second case, no data is shared between physical comparisons and we can expect that they are statistically independent, though empirically there will be nonzero correlations due to physical similarities between striations.
 For simplicity, let us consider the number of comparisons to lie somewhere between these two estimates.
+Note that when $\ell/d \approx 1$, as in bullet LEA comparisons, the minimal and the maximal number of comparisons are both close to 1.
 
@@
 This is a minimal scenario - the wire will acquire striations from one side of the blade, while the blade itself has two cutting edges, which we will call side A and side B.
-A blade cut of a sheet of aluminum will thus produce two striated edges corresponding to side A and side B which can be compared to cut wires to assess similarity.
+A blade cut of a sheet of aluminum will thus produce two striated edges corresponding to side A and side B which are compared to cut wires to assess similarity.
 We also have a 12 gauge aluminum wire ($\Sexpr{d}$ mm diameter) which may have been cut by the wire-cutting tool described above.
@@
 
-In this example, $\ell = \Sexpr{ell}$ mm, $d = \Sexpr{d}$ mm, and at minimum there are $\ell/d = \Sexpr{ncomp_min}$ comparisons between a wire cut and a blade cut.
+In this example, $\ell = \Sexpr{ell}$ mm, $d = \Sexpr{d}$ mm, and there are at least $\ell/d = \Sexpr{ncomp_min}$ comparisons between a wire cut and a blade cut.
 As there are two blade cuts (side A and side B), the minimal number of comparisons overall is \Sexpr{ncomp_min*2}, as these comparisons are non-overlapping and independent (on average).
 
-Let's assume a resolution of $\Sexpr{res*1000} \mu$m per pixel, the maximum number of comparisons per blade cut is approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons; thus, we need $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall in order to find the optimal alignment between the wire and the blade cut.
+Assuming a resolution of $\Sexpr{res*1000} \mu$m per pixel, the maximum number of comparisons per blade cut is approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons; thus, we need $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall in order to find the optimal alignment between the wire and the blade cut.
 These comparisons are implicit in the calculation of cross-correlation, which is the first and often the only step used to quantitatively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
@@
 While these sequential comparisons are highly auto correlated, and we cannot assume sequential independence when calculating the probability of an error, they serve as an upper bound on the number of comparisons which could be performed.
+
+As the number of comparisons increases, the probability of encountering a coincidental match increases.
+Statisticians call this the \textit{familywise error rate} ("family" = number of comparisons); it is an important quantity to control when conducting a series of tests.
 
@@
 
-If we begin by considering the minimal $\ell/d$ independent comparisons between a blade cut and a wire, we can compute the overall false positive rate for all comparisons using $e$, the false positive rate for one comparison, and simple probability calculations.
-Let $$E_n:=P\left(\{\geq 1\text{ false pos. in }n\text{ disjoint comparisons}\}\right).$$
-Then, $E_1 = e$, $\overline{E_1} = 1-e$, $\overline{E_n} = \left[\overline{E_1}\right]^n$, and $E_n = 1 - \overline{E_n} = 1 - [1-e]^n$.
-% Using this relationship between $n$ and the false positive rate $e$, we can determine what $e$ must be in order to limit $E_n$ to an acceptable level by solving for $e$: $e = 1 - \left[1 - E_n\right]^{1/n}$.
-\Cref{fig:Accuracy} shows $E_n$ on the y-axis; if 90\% is an acceptable examination-wide false positive rate (FPR), 2 comparisons can be made when the single-comparison FPR is 0.05; this increases to 10 comparisons at 0.01.
-Thus, at an error rate of \Sexpr{sprintf("%.2f", fpr)}, examiners can conduct approximately 5 comparisons without exceeding the examination-wide FPR of 10\%.
+For a single-comparison FPR of $e$, the familywise FPR for $n$ comparisons is $1 - [1-e]^n$.
+\Cref{fig:Accuracy} shows the familywise accuracy rate on the y-axis; if 90\% is an acceptable overall accuracy, 2 comparisons can be made when $e = 0.05$; this increases to 10 comparisons when $e =  0.01$.
+Thus, at an error rate of \Sexpr{sprintf("e = %.2f", fpr)}, examiners can conduct approximately 5 comparisons without exceeding the familywise FPR of 10\%.
 
-It is clear that under these constraints, the accuracy of an examination involving multiple comparisons between a wire and a tool is likely to be low, as the number of candidate alignments that must be examined is too high.
-Even the most innocuous example we could find (small blade, only 2 cutting surfaces, and a relatively large wire) involved a minimum of \Sexpr{2*ncomp_min} completely independent comparisons.
-As a result, findings from  wire comparisons made under current protocols (using error rates recently published in the forensics literature) are probably not reliable enough to be presented at trial.
-When we then consider that examiners would make cuts at multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons even in the simple case we have presented here, the probability of a false positive error increases further still.
+It is clear that under these constraints, the accuracy of an examination involving multiple comparisons between a wire and a tool will be low, as the number of candidate alignments that must be examined is high.
+Even the most innocuous example (small blade, only 2 cutting surfaces, and a relatively large wire) involved a minimum of \Sexpr{2*ncomp_min} completely independent comparisons.
+When we then consider that examiners would make cuts at multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons even in the simple case we have presented here, the probability of a false identification becomes even more likely.
+As a result, it is questionable whether wire comparisons made under current protocols are reliable enough to be presented at trial.
 
@@
 Forensic practitioners often report the findings from their examinations in the form of a categorical conclusion reflecting a single decision.
-This can be misleading when this decision relies on multiple comparisons which are not individually presented in reports or testimony.
-In this short contribution, we have shown that the implicit comparisons performed during the forensic analysis of wire cuts have the potential to vastly increase the overall chance that at least one false positive error occurs.
-%While often a forensic conclusion is presented as a single decision, underlying that decision are many more comparisons which are not discussed or reported upon.
-Omission of the underlying details is not necessarily due to malice; rather, it results from protocols which focus on finding the best possible match between two items and then evaluating that match, and also from lack of familiarity with basic statistical ideas.
+This is misleading when the decision relies on multiple comparisons which are not individually presented in reports or testimony.
+In this short contribution, we have shown that the implicit comparisons performed during forensic analysis of wire cuts increase the familywise error rate.
+Omission of the underlying details is not necessarily due to malice; rather, it results from protocols which focus on finding the best possible match between two items and evaluating that match.
 
 In our example we highlight a simple case where a wire is cut using a two-sided blade, but the findings apply to many other situations where a forensic evaluation involves multiple comparisons against a set of accumulated information, such as database searches.
-%either in a database or in a series of test samples created during the examination process.
 It is critically important that forensic practitioners become aware of the impact of carrying out multiple comparisons on their final conclusions and that lab procedures and testimony are modified accordingly.

### 2024-01-19 PNAS/index.Rnw
updated E language and family-wiseness

@@
 As the number of comparisons increases, the probability of encountering a coincidental match increases.
-Statisticians call this the \textit{familywise error rate} ("family" = number of comparisons); it is an important quantity to control when conducting a series of tests.
+Statisticians call this the \textit{family-wise error rate} $E$; it is an important quantity to control when conducting a series ("family") of tests.
 
@@
 
-For a single-comparison FPR of $e$, the familywise FPR for $n$ comparisons is $1 - [1-e]^n$.
-\Cref{fig:Accuracy} shows the familywise accuracy rate on the y-axis; if 90\% is an acceptable overall accuracy, 2 comparisons can be made when $e = 0.05$; this increases to 10 comparisons when $e =  0.01$.
-Thus, at an error rate of \Sexpr{sprintf("e = %.2f", fpr)}, examiners can conduct approximately 5 comparisons without exceeding the familywise FPR of 10\%.
+For a single-comparison FPR of $e$, the family-wise FPR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
+\Cref{fig:Accuracy} shows the family-wise accuracy rate on the y-axis; if 90\% is an acceptable overall accuracy, 2 comparisons can be made when $e = 0.05$; this increases to 10 comparisons when $e =  0.01$.
+Thus, at an error rate of \Sexpr{sprintf("e = %.2f", fpr)}, examiners can conduct approximately 5 comparisons without exceeding the family-wise FPR of 10\%.
 
@@
 This is misleading when the decision relies on multiple comparisons which are not individually presented in reports or testimony.
-In this short contribution, we have shown that the implicit comparisons performed during forensic analysis of wire cuts increase the familywise error rate.
+In this short contribution, we have shown that the implicit comparisons performed during forensic analysis of wire cuts increase the family-wise error rate.
 Omission of the underlying details is not necessarily due to malice; rather, it results from protocols which focus on finding the best possible match between two items and evaluating that match.
@@
 
-First, examiners should report (or defense attorneys should request) the overall length or area of surfaces generated from blade cuts, test fires, and equivalent procedures intended to assess the marks made by a tool submitted for comparison with evidence.
+First, examiners should report (or defense attorneys should request) the overall length or area of surfaces generated during the examination process.
+% from blade cuts, test fires, and equivalent procedures intended to assess the marks made by a tool submitted for comparison with evidence.
 Examiners should also report the total consecutive length or area of the recovered evidence.
-These pieces of information will allow for calculations of minimum and maximum number of comparisons (as demonstrated in this paper), facilitating calculation of overall error rates.
+These pieces of information will take the place of $\ell$ and $d$ and facilitate calculation of examination-wide error rates.
 
@@
 This was necessary because we could not find black-box error rate studies of wire cuts.
-These studies should be Type II studies as defined by Koehler \citep{koehlerIntuitiveErrorRate2017a}, in that they should be designed to assess the actual error rate when examiners are making difficult comparisons.
+New studies should be Type II studies as defined by Koehler \citep{koehlerIntuitiveErrorRate2017a}, designed to assess the actual error rate when examiners are making difficult comparisons.
 
 Finally, when databases are used at any stage of the forensic evidence evaluation process (from suitability assessment and triage to reports which will be used at trial), the number of database items searched (or comparisons made) and the number of results returned must be reported.
-Furthermore, the number of results used for further manual comparison should also be reported.
-For example, if a firearms examiner searches a local NIBIN database with 1000 entries, requests the 20 closest matches to her evidence, and then carries out a physical examination of five exemplars from the list of 20, all of those values should be clearly reported to enable calculation of the potential FPR.
+Additionally, the number of results used for further manual comparison should also be reported.
+For example, if a firearms examiner searches a local NIBIN database with 1000 entries, requests the 20 closest matches to her evidence, and then carries out a physical examination of five exemplars from the list of 20, all of those values should be clearly reported to enable estimation of the examination-wise error rate.
 This will help make the multiple comparison issue accessible to everyone involved in evaluating the value of forensic evidence, from examiners to lawyers to jurors and judges.

### 2024-01-19 PNAS/index.Rnw
Updates to spelling

@@
 When wires are cut, the tool produces striations on the cut surfaced; as in other forms of forensic analysis, these striation marks are used to connect the evidence to the source that created them.
-Here, we argue that the practice of comparing two wire cut surfaces introduces complexities not present in better-investigated forensic examination  of toolmars such as those observed on bullets, as wire comparisons inherently require multiple distinct comparisons, increasing the expected false positive rate.
+Here, we argue that the practice of comparing two wire cut surfaces introduces complexities not present in better-investigated forensic examination  of toolmarks such as those observed on bullets, as wire comparisons inherently require multiple distinct comparisons, increasing the expected false positive rate.
 We call attention to the multiple comparison problem in wire examination and relate it to other situations in forensics that involve multiple comparisons, such as database searches.
@@
 \firstpage[3]{6}
-% Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting.
-In this example, there are a total of 11 paragraphs on the first page, counting the first level heading as a paragraph.
-The value {12} represents the number of the paragraph starting the second page.
-If a paragraph runs over onto the second page, include a bracket with the paragraph line number starting the second page, followed by the paragraph number in curly brackets, e.g.\ "\firstpage[4]{11}".
-
-% If your first paragraph (i.e.
-with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
-If this is the case, add \parshape=0 to the end of the list environment.
-% \dropcap{T}his PNAS journal template is provided to help you write your work in the correct journal format. Instructions for use are provided below.
 
@@
 Multiple comparisons arise persistently when developing statistical methods to address scientific problems \citep{benjaminiControllingFalseDiscovery1995}, and greatly increase the probability of false discoveries.
-%Yet, this potentially significant issue is often ignored.
 Now that vast databases and efficient algorithms are routinely used in forensic evaluations to propose matches to crime scene items, the problem of close non-matches~\citep{pcast}  due to multiple comparisons becomes critically important.
 Scientists have the responsibility to address this often ignored issue to prevent a further erosion of the public's confidence in the justice system.
-%Researchers have been focusing on building up databases to develop algorithms for objective evidence comparisons.
-%\svp{Then, the algorithms facilitate} searches across (vast) databases, reporting the best $N$ matches to a particular piece of evidence.
-%This raises the critically important problem of .
-
-%and explore closest non-matches .
-While these databases (and open data) are critically important in forensics, we also need to be careful about how our methods are affected by the multiple comparison problem.
 
@@
 The probability of finding  close non-matches for an item of evidence is proportional to the size of the database available for searching; close non-matches occur more frequently as databases increase in size and algorithms become more powerful.
-% In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.
-
-%
-% \hh{Database searches are an example where the number of searches is explicitly available as the number of objects meeting the search criteria.
-} % This doesn't seem super-relevant and distracts from the continuity between the algorithm discussion above and the next paragraph
-
-%\hh{XXX director's notes: XXX what is different between full and partial matches is that error rates in the literature are generally reported based on full length comparisons, where macro features are used to anchor comparisons (align along grooves, then zoom in and only 'wiggle' a bit to the left and right to adjust at the micro-scale).
-These kind of adjustments are 'baked' into published error rates.
-}
 
@@
 Assuming that lab procedure errors are not a factor in studies, we use reported error rates from striated evidence comparisons (e.g.\ excluding breech face comparisons) in open-set studies as an estimate of the coincidental match rate of a single wire-cut comparison.
-%% Do we need to address the fact that wire cuts don't have external features to align, where bullets and firing pins do???
 The three studies we consider \citep{bajic2020, mattijssen2021,best2022} have false positive rates between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
@@
 First, examiners should report (or defense attorneys should request) the overall length or area of surfaces generated during the examination process.
-% from blade cuts, test fires, and equivalent procedures intended to assess the marks made by a tool submitted for comparison with evidence.
 Examiners should also report the total consecutive length or area of the recovered evidence.
@@
 
-%We believe that the multiple comparison problems we have discussed here can be overcome with research and changes in evidence assessment procedures, but the first step is making the number of comparisons performed more visible to everyone involved in the process.
-
-% Note: please start your introduction without including the word ``Introduction'' as a section heading (except for math articles in the Physical Sciences section); this heading is implied in the first paragraph.
-
-% \section*{Guide to using this template on Overleaf}
-%
-% Please note that whilst this template provides a preview of the typeset manuscript for submission, to help in this preparation, it will not necessarily be the final publication layout.
-For more detailed information please see the \href{ https://www.pnas.org/page/authors/format}{PNAS Information for Authors}.
-%
-% If you have a question while using this template on Overleaf, please use the help menu (``?'') on the top bar to search for \href{https://www.overleaf.com/help}{help and tutorials}.
-You can also \href{https://www.overleaf.com/contact}{contact the Overleaf support team} at any time with specific questions about your manuscript or feedback on the template.
-%
-% \subsection*{Author Affiliations}
-%
-% Include department, institution, and complete address, with the ZIP/postal code, for each author.
-Use lower case letters to match authors with institutions, as shown in the example.
-PNAS strongly encourages authors to supply an \href{https://orcid.org/}{ORCID identifier} for each author.
-Individual authors must link their ORCID account to their PNAS account at \href{http://www.pnascentral.org/}{www.pnascentral.org}.
-For proper authentication, authors must provide their ORCID at submission and are not permitted to add ORCIDs on proofs.
-%
-% \subsection*{Submitting Manuscripts}
-%
-% All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}.
-If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.
-%
-% \subsection*{Format}
-%
-% Many authors find it useful to organize their manuscripts with the following order of sections: title, author line and affiliations, keywords, abstract, introduction, results, discussion, materials and methods, acknowledgments, and references.
-Other orders and headings are permitted.
-%
-% \subsection*{Manuscript Length}
-%
-% Brief Reports are limited to 3 pages, which is approximately 1,600 words (including the manuscript text, title page, abstract, and figure legends), and 15 references.
-Supporting information (SI) is limited to extended methods, essential supporting datasets, and videos (no additional tables or figures).
-%
-%
-% \subsection*{References}
-%
-% References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g.\ \cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric,phdthesis,masterthesis}.
-All references cited in the main text should be included in the main manuscript file.
-%
-% \subsection*{Data Archival}
-%
-% PNAS must be able to archive the data essential to a published article.
-Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the \href{https://www.pnas.org/author-center/editorial-and-journal-policies#materials-and-data-availability}{Information for Authors}, is acceptable.
-%
-% \subsection*{Language-Editing Services}
-% Prior to submission, authors who believe their manuscripts would benefit from professional editing are encouraged to use a language-editing service (see list at https://www.pnas.org/author-center/language-editing).
-PNAS does not take responsibility for or endorse these services, and their use has no bearing on acceptance of a manuscript for publication.
-%
-% \begin{figure}[t!]
-% \centering
-% \includegraphics[width=.8\linewidth]{frog}
-% \caption{Placeholder image of a frog with a long example legend to show justification setting.}
-% \label{fig:frog}
-% \end{figure}
-%
-%
-% \begin{SCfigure*}[\sidecaptionrelwidth][t]
-% \centering
-% \includegraphics[width=11.4cm,height=11.4cm]{frog}
-% \caption{This legend would be placed at the side of the figure, rather than below it.}\label{fig:side}
-% \end{SCfigure*}
-%
-% \begin{table}[t!]
-% \centering
-% \caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
-% \begin{tabular}{lrrr}
-% Species & CBS & CV & G3 \\
-% \midrule
-% 1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
-% 2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
-% 3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
-% \bottomrule
-% \end{tabular}
-%
-% \addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
-% \end{table}
-%
-%
-% \subsection*{Digital Figures}
-%
-% EPS, high-resolution PDF, and PowerPoint are preferred formats for figures that will be used in the main manuscript.
-Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
-Color images must be in RGB (red, green, blue) mode.
-Include the font files for any text.
-%
-% Images must be provided at final size, preferably 1 column width (8.7cm).
-Figures wider than 1 column should be sized to 11.4cm or 17.8cm wide.
-Numbers, letters, and symbols should be no smaller than 6 points (2mm) and no larger than 12 points (6mm) after reduction and must be consistent.
-%
-% Figures and tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.
-%
-% Figure \ref{fig:frog} shows an example of how to insert a column-wide figure.
-To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment.
-Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
-Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side legends.
-%
-% \subsection*{Tables}
-% Tables should be included in the main manuscript file and should not be uploaded separately.
-%
-% \subsection*{Single column equations}
-%
-% Authors may use 1- or 2-column equations in their article, according to their preference.
-%
-% To allow an equation to span both columns, use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures.
-%
-% Note that the use of the \verb|widetext| environment for equations is not recommended, and should not be used.
-%
-% \begin{figure*}[bt!]
-% \begin{align*}
-% (x+y)^3&=(x+y)(x+y)^2\\
-%        &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
-%        &=x^3+3x^2y+3xy^3+x^3.
-% \end{align*}
-% \end{figure*}
-%
-%
-%
-% \subsection*{Supporting Information (SI)}
-%
-% Authors are limited to the following types of SI: extended methods, datasets, videos, and 3D figures. Extended discussion is not permitted.
-%
-% \subsection*{SI Appendix}
-%
-% Supply extended materials and methods information in a separate SI Appendix PDF file.
-Include SI movie legends.
-SI will be published as provided by the authors; it will not be edited or composed.
-Supplementary figures and tables are not allowed.
-%
-%
-% \subsubsection*{SI Datasets}
-%
-% Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be published in raw format and will not be edited or composed.
-%
-%
-% \subsubsection*{SI Movies}
-%
-% Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and and include a brief legend for each movie in the main manuscript file.
-All movies should be submitted at the desired reproduction size and length.
-Movies should be no more than 10 MB in size.
-%
-%
-%
-%
-% \matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required.
-%
-% \subsection*{Subsection for Method}
-% Example text for subsection.
-% }
-%
-% \showmatmethods{} % Display the Materials and Methods section
-
 \acknow{This work was funded (or partially funded) by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreements 70NANB15H176 and 70NANB20H019 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, Duke University, University of California Irvine, University of Virginia, West Virginia University, University of Pennsylvania, Swarthmore College and University of Nebraska, Lincoln.}
@@
 
-% \bibsplit[3]
-% %Use \bibsplit to split the references from the body of the text. Value "[3]" represents the number of reference in the left column (Note: Please avoid single column figures & tables on this page.)
-
-% Bibliography
-% \bibliography{pnas-sample}
-
 \bibliography{../references}

### 2024-01-19 PNAS/index.Rnw
Combine figures

@@
 Note that when $\ell/d \approx 1$, as in bullet LEA comparisons, the minimal and the maximal number of comparisons are both close to 1.
-
-\begin{figure}[t]
-\includegraphics[width=\columnwidth]{../fig/wire-blade-comparison-label.png}
-\caption{A comparison between a wire and a blade cut requires sliding the wire along the entire blade cut length to determine the best match (or whether there is a match).
-Surfaces shown are rendered 2D topographical scans of a wire and blade cut taken with a confocal light microscope.}\label{fig:wire-blade}
-\end{figure}
-
 Finally, we must consider the number of surfaces which must be compared: the wire may have between one and two sets of striae and there may be between 2 and 4 different blade cut surfaces to examine, depending on tool configuration.
@@
 \centering
+\includegraphics[width=\columnwidth]{../fig/wire-blade-comparison-label.png}
 \includegraphics[width=\columnwidth]{../fig/wire_cutter_wires.png}
-\caption{RJ45 Crimp tool with a $\Sexpr{ell/10}$ cm razor blade used for cutting.
-1 mm and 2 mm diameter aluminum wires cut with the pliers are shown in a box in the top right corner.}\label{fig:tool-pic}
+\caption{(Top) A comparison between a wire and a blade cut requires sliding the wire along the entire blade cut length to determine the best match (or whether there is a match).
+Surfaces shown are rendered 2D topographical scans of a wire and blade cut taken with a confocal light microscope.
+(Bottom) RJ45 Crimp tool with a $\Sexpr{ell/10}$ cm razor blade used for cutting.
+1 mm and 2 mm diameter aluminum wires cut with the pliers are shown in a box in the top right corner.}\label{fig:tool-pic}\label{fig:wire-blade}
 \end{figure}

### 2024-01-30 PNAS/index.Rnw
Fix the title, remove puns :(
Trimming spaces and stuff.
Tweaking things to remove unnecessary space/words.

@@
 
-\title{Mirror, mirror on the wall. Which's the best match of them all?}
+\title{Hidden Multiple Comparisons Increase Forensic Error Rates}
 
@@
 
-\firstpage[3]{6}
+\firstpage[5]{5}
 
@@
 Scientists have the responsibility to address this often ignored issue to prevent a further erosion of the public's confidence in the justice system.
-
 The multiple comparison problem is not new: it has been raised in the past with regard to DNA~\citep{thompson2003} and latent print evaluations~ \citep{koehler2021}.
@@
 
-Compounding this issue, the use of novel algorithms typically results in a large number of comparisons that may not obvious to the user.
+Compounding this issue, the use of algorithms also results in a large number of comparisons that are not obvious to the user.
 For example, the cross-correlation function~\citep{vorburgerApplicationsCrosscorrelationFunctions2011} was one of the first measures proposed for quantifying the similarity between two patterns in response to the 2009 NRC report~\citep{nas2009}, and continues to be used in many pattern searching algorithms to find the best alignment between two images and to quantify their overall similarity.
@@
 This mirrors the forensic examination process: the examiner visually rotates and shifts items under a comparison microscope to align two surfaces.
-In order to avoid false accusations and a further erosion of the public's trust in science, it is important to address the problem of multiple comparisons by controlling their effect on false discovery rates.
+In order to avoid false accusations and further erosions of public trust in science, we must address the problem of multiple comparisons and control their effect on false discovery rates.
 
@@
 Note that when $\ell/d \approx 1$, as in bullet LEA comparisons, the minimal and the maximal number of comparisons are both close to 1.
-Finally, we must consider the number of surfaces which must be compared: the wire may have between one and two sets of striae and there may be between 2 and 4 different blade cut surfaces to examine, depending on tool configuration.
+Finally, we must consider the number of surfaces which must be compared: the wire may have between one and two sets of striae and there may be 2-4 different blade cut surfaces to examine, depending on tool configuration.
 This results in a multiplier of between 2 and 8 to calculate the total number of comparisons required.
@@
 (that is, the wire appears to have been cut with a tool of similar configuration).
-
 In this example, $\ell = \Sexpr{ell}$ mm, $d = \Sexpr{d}$ mm, and there are at least $\ell/d = \Sexpr{ncomp_min}$ comparisons between a wire cut and a blade cut.
@@
 
-Assuming a resolution of $\Sexpr{res*1000} \mu$m per pixel, the maximum number of comparisons per blade cut is approximately $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$ comparisons; thus, we need $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall in order to find the optimal alignment between the wire and the blade cut.
+Assuming a resolution of $\Sexpr{res*1000} \mu$m per pixel, the maximum number of comparisons per blade cut is around $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$; thus, we need $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall in order to find the optimal alignment between the wire and the blade cut.
 These comparisons are implicit in the calculation of cross-correlation, which is the first and often the only step used to quantitatively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
@@
 While these sequential comparisons are highly auto correlated, and we cannot assume sequential independence when calculating the probability of an error, they serve as an upper bound on the number of comparisons which could be performed.
-
 As the number of comparisons increases, the probability of encountering a coincidental match increases.
@@
 \section*{Probability of False Identifications}
-There are at least two components of the false positive rate: identifying two pieces of evidence that have similar characteristics but are from different sources (a coincidental match) and procedural failures (e.g.\ lab process errors) \citep[p 50]{pcast}.
+There are at least two components of the false positive rate (FPR): identifying two pieces of evidence that have similar characteristics but are from different sources (a coincidental match) and procedural failures (e.g.\ lab process errors) \citep[p 50]{pcast}.
 In objective disciplines with standardized evaluation rules (e.g.\ DNA), these sources can be distinguished.
 However, in toolmark examination, no objective evaluation rules are used; examiners testify based on subjective, individual rules for how much similarity is sufficient for an identification.
-Assuming that lab procedure errors are not a factor in studies, we use reported error rates from striated evidence comparisons (e.g.\ excluding breech face comparisons) in open-set studies as an estimate of the coincidental match rate of a single wire-cut comparison.
-The three studies we consider \citep{bajic2020, mattijssen2021,best2022} have false positive rates between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these studies yields a false positive rate of \Sexpr{sprintf("%.2f", fpr)}.
+Assuming that lab procedure errors are not a factor in studies, we use reported error rates from open-set studies of striated evidence comparisons to estimate the coincidental match rate of a single wire-cut comparison.
+The three studies we consider \citep{bajic2020, mattijssen2021,best2022} have FPRs between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these studies yields a FPR of \Sexpr{sprintf("%.2f", fpr)}.
+For a single-comparison FPR of $e$, the family-wise FPR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
+\Cref{fig:Accuracy} shows the family-wise accuracy rate on the y-axis; if 90\% is an acceptable overall accuracy, 2 comparisons can be made when $e = 0.05$; increasing to 10 comparisons when $e =  0.01$.
+Thus, at an error rate of \Sexpr{sprintf("e = %.2f", fpr)}, examiners can conduct comparisons before exceeding the family-wise FPR of 10\%.
 
-For a single-comparison FPR of $e$, the family-wise FPR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
-\Cref{fig:Accuracy} shows the family-wise accuracy rate on the y-axis; if 90\% is an acceptable overall accuracy, 2 comparisons can be made when $e = 0.05$; this increases to 10 comparisons when $e =  0.01$.
-Thus, at an error rate of \Sexpr{sprintf("e = %.2f", fpr)}, examiners can conduct approximately 5 comparisons without exceeding the family-wise FPR of 10\%.
-
-It is clear that under these constraints, the accuracy of an examination involving multiple comparisons between a wire and a tool will be low, as the number of candidate alignments that must be examined is high.
+Under these constraints, the accuracy of an examination involving multiple comparisons between a wire and a tool will be low, as the number of candidate alignments that must be examined is high.
 Even the most innocuous example (small blade, only 2 cutting surfaces, and a relatively large wire) involved a minimum of \Sexpr{2*ncomp_min} completely independent comparisons.
-When we then consider that examiners would make cuts at multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons even in the simple case we have presented here, the probability of a false identification becomes even more likely.
+When we then consider that examiners would make cuts at multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons, the probability of a false identification becomes even more likely.
 As a result, it is questionable whether wire comparisons made under current protocols are reliable enough to be presented at trial.
 
-<<control-overall-fpr, fig.cap="Exploration of overall error rate estimates for different values of $e$, the false positive rate for a single comparison.
+<<control-overall-fpr, fig.cap="Exploration of overall error rate estimates for different values of $e$, the FPR for a single comparison.
 Dotted vertical lines show false positive error rate estimates (from left to right, \\citep{best2022,bajic2020,mattijssen2021}).", fig.width = 8, fig.height = 4, include = F>>=
@@
 
-First, examiners should report (or defense attorneys should request) the overall length or area of surfaces generated during the examination process.
-Examiners should also report the total consecutive length or area of the recovered evidence.
+First, examiners should report (or defense attorneys should request) the overall length or area of surfaces generated during the examination process, along with the total consecutive length or area of the recovered evidence.
 These pieces of information will take the place of $\ell$ and $d$ and facilitate calculation of examination-wide error rates.
@@
 Second, researchers should conduct studies relating the length/area of comparison surface to the error rate.
-For instance, we have pooled studies looking at bullet striations and firing pin shear marks.
-The striated surfaces are of orders of magnitude different lengths, but they represent our best estimate of the error rate for striated comparisons.
-This was necessary because we could not find black-box error rate studies of wire cuts.
+For instance, we have pooled studies looking at bullet striations and firing pin shear marks because we could not find black-box error rate studies of wire cuts.
+The striated surfaces are of orders of magnitude different lengths, but represent the best estimate of the error rate for striated materials.
 New studies should be Type II studies as defined by Koehler \citep{koehlerIntuitiveErrorRate2017a}, designed to assess the actual error rate when examiners are making difficult comparisons.

### 2024-02-01 PNAS/index.Rnw
P(F_1) -> e

@@
 r_vals <- tibble(r = c(.0005, 0.001, 0.005, .01, 0.05),
-                 rlab = sprintf("P(F[1])==%0.05f", r))
+                 rlab = sprintf("e==%0.05f", r))
 res2 <- crossing(r = r_vals$r,

### 2024-02-02 PNAS/index.Rnw
Alicia's changes

@@
 \begin{abstract}
-When wires are cut, the tool produces striations on the cut surfaced; as in other forms of forensic analysis, these striation marks are used to connect the evidence to the source that created them.
+When wires are cut, the tool produces striations on the cut surface; as in other forms of forensic analysis, these striation marks are used to connect the evidence to the source that created them.
 Here, we argue that the practice of comparing two wire cut surfaces introduces complexities not present in better-investigated forensic examination  of toolmarks such as those observed on bullets, as wire comparisons inherently require multiple distinct comparisons, increasing the expected false positive rate.
@@
 Now that vast databases and efficient algorithms are routinely used in forensic evaluations to propose matches to crime scene items, the problem of close non-matches~\citep{pcast} due to multiple comparisons becomes critically important.
-Scientists have the responsibility to address this often ignored issue to prevent a further erosion of the public's confidence in the justice system.
+This often ignored issue increases the false positive rate, and can contribute to the erosion of public trust in the justice system.
 The multiple comparison problem is not new: it has been raised in the past with regard to DNA~\citep{thompson2003} and latent print evaluations~ \citep{koehler2021}.
 One of the root causes~\citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid train bombing case was that the large size of the IAFIS database used to search for similar prints made it possible to locate `unusually' close non-matches.
-The probability of finding  close non-matches for an item of evidence is proportional to the size of the database available for searching; close non-matches occur more frequently as databases increase in size and algorithms become more powerful.
+The probability of finding  close non-matches for an item of evidence is proportional to the size of the database available for searching; close non-matches occur more frequently as databases increase in size.
 
@@
 The first option gives us the maximum number of comparisons ($\ell/r - d/r + 1$), while the second option gives us the minimum number of comparisons $\ell/d$.
-In the first case, sequential comparisons share much of the same physical data and are highly related; in the second case, no data is shared between physical comparisons and we can expect that they are statistically independent, though empirically there will be nonzero correlations due to physical similarities between striations.
+In the first case, sequential comparisons share much of the same physical data and are highly related; in the second case, no data are shared between physical comparisons and we can expect that they are statistically independent, though empirically there will be nonzero correlations due to physical similarities between striations.
 For simplicity, let us consider the number of comparisons to lie somewhere between these two estimates.
-Note that when $\ell/d \approx 1$, as in bullet LEA comparisons, the minimal and the maximal number of comparisons are both close to 1.
-Finally, we must consider the number of surfaces which must be compared: the wire may have between one and two sets of striae and there may be 2-4 different blade cut surfaces to examine, depending on tool configuration.
-This results in a multiplier of between 2 and 8 to calculate the total number of comparisons required.
+Note that when $\ell/d \approx 1$, as in some toolmark comparisons, the upper number of comparisons goes to 1.
+Finally, we must consider the number of surfaces which must be compared: the wire may have one or two sets of striae and there may be  two to four blade cut surfaces to examine, depending on the tool.
+This results in a multiplier of as much as 8.
 
@@
 In this short contribution, we have shown that the implicit comparisons performed during forensic analysis of wire cuts increase the family-wise error rate.
-Omission of the underlying details is not necessarily due to malice; rather, it results from protocols which focus on finding the best possible match between two items and evaluating that match.
+%Omission of the underlying details is not necessarily due to malice; rather, it results from protocols which focus on finding the best possible match between two items and evaluating that match.
 
-In our example we highlight a simple case where a wire is cut using a two-sided blade, but the findings apply to many other situations where a forensic evaluation involves multiple comparisons against a set of accumulated information, such as database searches.
-It is critically important that forensic practitioners become aware of the impact of carrying out multiple comparisons on their final conclusions and that lab procedures and testimony are modified accordingly.
-We suggest three strategies to improve transparency of forensic examinations and to produce more reliable estimates of examination-wide error rates.
+We describe a simple scenario where a wire is cut using a two-sided blade, but findings apply to any situation where a forensic evaluation involves multiple comparisons, including, e.g., database searches.
+Forensic practitioners should understand how the number of comparisons can affect the accuracy of their final conclusion.
+We propose three strategies to enhance transparency and enable more reliable estimates of examination-specific error rates.
+%In our example we highlight a simple case where a wire is cut using a two-sided blade, but the findings apply to many other situations where a forensic evaluation involves multiple comparisons against a set of accumulated information, such as database searches.
+%It is critically important that forensic practitioners become aware of the impact of carrying out multiple comparisons on their final conclusions and that lab procedures and testimony are modified accordingly.
+%We suggest three strategies to improve transparency of forensic examinations and to produce more reliable estimates of examination-wide error rates.
 

### 2024-02-05 PNAS/index.Rnw
new 'figure' ?

@@
 Assuming that lab procedure errors are not a factor in studies, we use reported error rates from open-set studies of striated evidence comparisons to estimate the coincidental match rate of a single wire-cut comparison.
-The three studies we consider \citep{bajic2020, mattijssen2021,best2022} have FPRs between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these studies yields a FPR of \Sexpr{sprintf("%.2f", fpr)}.
+The three studies we consider \citep{bajic2020, mattijssen2021,best2022} have FPRs between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these studies yields an FPR of \Sexpr{sprintf("%.2f", fpr)}.
 For a single-comparison FPR of $e$, the family-wise FPR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
-\Cref{fig:Accuracy} shows the family-wise accuracy rate on the y-axis; if 90\% is an acceptable overall accuracy, 2 comparisons can be made when $e = 0.05$; increasing to 10 comparisons when $e =  0.01$.
-Thus, at an error rate of \Sexpr{sprintf("e = %.2f", fpr)}, examiners can conduct comparisons before exceeding the family-wise FPR of 10\%.
+\hh{\Cref{tab:familywise} shows the impact the number of comparisons has on these published error rates.
+With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.
+only the comparisons of one wire to a single-bladed tool are safely covered without exceeding an upper bound of 10\% for the family wise false identification error.}
+%\Cref{tab:Accuracy} shows the family-wise accuracy rate on the y-axis; if 90\% is an acceptable overall accuracy, 2 comparisons can be made when $e = 0.05$; increasing to 10 comparisons when $e =  0.01$.
+%Thus, at an error rate of \Sexpr{sprintf("e = %.2f", fpr)}, examiners can conduct comparisons before exceeding the family-wise FPR of 10\%.
+\hh{For a modestly sized database search of size 1000, the initial false positive rate should not exceed 1 in 10,000 to guarantee a family-wise total false identification error of at most 10\%. }
 
@@
 
-<<Accuracy, fig.cap="Probability of accurate decisions for $N$ comparisons for different error rates.", fig.width = 8, fig.height = 4>>=
-r_vals <- tibble(r = c(.0005, 0.001, 0.005, .01, 0.05),
-                 rlab = sprintf("e==%0.05f", r))
-res2 <- crossing(r = r_vals$r,
-                 N = c(seq(1, 11, .01),seq(11.1, 109.9, .1), seq(110, 1100, 1))) %>%
-  mutate(overall = (1 - r)^N) %>%
-  left_join(r_vals)
+<<Accuracy, eval=FALSE,  fig.cap="Probability of accurate decisions for $N$ comparisons for different error rates.", fig.width = 8, fig.height = 4>>=
 
-breakpt <- .9
+comp_allowed <- function(e, at_most=0.1) {
+  floor(log(1-at_most)/log(1-e))
+}
 
-error_res <- filter(res2, lead(overall, 1) <= breakpt)
-res2 <- filter(res2, overall >= breakpt)
+family_wise <- function(e, N) {
+  1 - (1 - e)^N
+}
 
-res2sum <- res2 %>% group_by(r) %>% filter(overall == min(overall))
+r_vals <- tibble(r = c(0.004545, 0.00704, 0.02, 0.0724, 0.001, 0.0001, 0.00001),
+                 Study = c("Best (2022)","Bajic (2020)", "Pooled Error",  "Mattijssen (2021)", "", "", "")) %>% arrange(desc(r))
+#                 rlab = sprintf("e==%0.05f", r)) %>% arrange(desc(r))
 
-ggplot(data = res2, aes(x = N, y = overall, color = factor(r), label = rlab)) +
-  geom_textline(parse = T, gap = F, straight = F,
-                padding = unit(1, "mm"), vjust = 0, linewidth = 1) +
-  xlab("Number of comparisons") +
-  scale_x_log10(
-    breaks = c(1, 1000, 10000, res2sum$N),
-    labels = c("1", "1000", "10000", round(res2sum$N)),
-    minor_breaks = c((1:9)[-2],10*(3:9), 100*3:9, 1000*1:9)
-  ) +
-  geom_hline(yintercept = breakpt, linewidth = 1) +
-#  geom_label(
-#    data = res2sum,
-#    aes(x = N, y = breakpt, label = paste0("N == ", round(N))),
-#    parse = T, vjust = 1, hjust = 1) +
-  geom_segment(data = res2sum,
-     aes(x = N, xend=N, y = -Inf, yend=0.9, color=factor(r)),
-     linewidth=0.5) +
-  geom_line(data = error_res, linetype = 2, linewidth = 0.75, alpha =0.5) +
-  ylab("1 - Probability of false positive\n(N comparisons)") +
-  ggtitle("# Comparisons and Overall Accuracy") +
-  scale_color_discrete("FP Rate for single comparison", guide = 'none') +
-  theme_bw() +
-  coord_cartesian(ylim = c(0.775, 1.0)) +
-  theme(axis.text.y = element_text(hjust = 0.5, vjust = 0.5, angle = 90))
+r_vals <- r_vals %>% mutate(
+#  E1 = r,
+  E10 = family_wise(r, 10),
+  E100 = family_wise(r, 100),
+  E1000 = family_wise(r, 1000),
+  allowed = comp_allowed(r)
+)
+
+data <- r_vals %>%
+  mutate(
+    FPR = c("7.24%", "2.00%", "0.70%", "0.45%", "1 in 1,000", "1 in 10,000", "1 in 100,000")
+  ) %>%
+  select(Study, FPR, E10, E100, E1000, allowed)
+
+library(gt)
+
+gt_tab <- data %>%
+  mutate(across(c(E10, E100, E1000), function(x) 100*x)) %>%
+  gt() %>%
+  tab_spanner(label = "False Positive Error (in %) based on N comparisons", columns = matches("E10")) %>%
+  fmt_number(columns = c(E10, E100, E1000), decimals = 1) %>%
+  fmt_number(columns = c(allowed), sep_mark = ",", decimals=0) %>%
+  tab_header(
+    title = md("False identification rates increase with the number of comparisons made.")
+  ) %>%
+ cols_label(
+    E10 = "N = 10",
+    E100 = "N = 100",
+    E1000 = "N = 1,000",
+    allowed = "Number of comparisons\nfor error below 10%",
+    FPR = "False Positive Rate"
+  )
+
+# include modified latex based on latex produced below:
+as_latex(gt_tab)
+
+# r_vals <- r_vals %>% mutate(
+#   study = sprintf("%s:e=%s", study, r)
+# )
+
+# res2 <- crossing(r = r_vals$r,
+#                  N = c(seq(1, 11, .01),seq(11.1, 109.9, .1), seq(110, 1100, 1))) %>%
+#   mutate(overall = (1 - r)^N) %>%
+#   left_join(r_vals)
+#
+
+# breakpt <- .9
+#
+# error_res <- filter(res2, lead(overall, 1) <= breakpt)
+# res2 <- filter(res2, overall >= breakpt)
+#
+# res2sum <- res2 %>% group_by(r) %>% filter(overall == min(overall))
+#
+#
+# ggplot(data = res2, aes(x = N, y = 1-overall, color = study, label = rlab)) +
+# #  geom_textline(parse = T, gap = F, straight = F,
+# #                padding = unit(1, "mm"), vjust = 0, linewidth = 1) +
+#   geom_line(linewidth = 1) +
+#   geom_line(linewidth = 1, data = error_res, alpha = 0.5) +
+#   xlab("Number of comparisons n") +
+#   scale_x_log10(
+#     breaks = c(1, 10000, res2sum$N),
+#     labels = c("1", "10000", round(res2sum$N)[-4], "1.4"),
+#     minor_breaks = c((1:9)[-2],10*(3:9), 100*3:9, 1000*1:9)
+#   ) +
+#   geom_hline(yintercept = 0.1, linewidth = 1) +
+# #  geom_label(
+# #    data = res2sum,
+# #    aes(x = N, y = breakpt, label = paste0("N == ", round(N))),
+# #    parse = T, vjust = 1, hjust = 1) +
+#   geom_point(data = res2sum,
+#      aes(x = N, y = 0.1, color=study),
+#      size=10) +
+# #  geom_segment(data = res2sum,
+# #     aes(x = N, xend=N, y = -Inf, yend=0.1, color=study),
+# #     linewidth=0.5) +
+# #  geom_line(data = error_res, linetype = 2, linewidth = 0.75, alpha =0.5) +
+#   ylab("Family-wise false positive rate E") +
+#   ggtitle("# Comparisons and Overall Accuracy") +
+#   scale_color_discrete("FP Rate for \nsingle comparison") +
+#   theme_bw() +
+#   coord_cartesian(ylim = c(0, .15)) +
+#   theme(axis.text.y = element_text(hjust = 0.5, vjust = 0.5, angle = 90))
 
 @
+
+\begin{table}
+\scalebox{.95}{
+\begin{tabular}{lcrrrr}
+\toprule
+ &  & \multicolumn{3}{p{25mm}}{False Positives (\%) in N comparisons} &  \\
+\cmidrule(lr){3-5}
+Study & FPR $e$ & $E_{10}$ & $E_{100}$ & $E_{1,000}$ & $E_N < 0.1$ \\
+\midrule\addlinespace[2.5pt]
+Mattijssen (2021) & 7.24\% & $52.8$ & $99.9$ & $100.0$ & $1$ \\
+Pooled Error & 2.00\% & $18.3$ & $86.7$ & $100.0$ & $5$ \\
+Bajic (2020) & 0.70\% & $6.8$ & $50.7$ & $99.9$ & $14$ \\
+Best (2022) & 0.45\% & $4.5$ & $36.6$ & $98.9$ & $23$ \\
+ & 1 in 1,000 & $1.0$ & $9.5$ & $63.2$ & $105$ \\
+ & 1 in 10,000 & $0.1$ & $1.0$ & $9.5$ & $1,053$ \\
+ & 1 in 100,000 & $10^{-4}$ & $0.1$ & $1.0$ & $10,535$ \\
+\bottomrule
+\end{tabular}}
+\caption{\label{tab:family-wise}Table showing the relationship between false positive rates and the chance of false identifications in $N$ comparisons for a set of different FPRs and different number of comparisons.
+The last column gives the number of comparisons allowed while ensuring a false identification percentage of at most 10\%.}
+\end{table}
 

### 2024-02-05 PNAS/index.Rnw
and we are back to 3.17

@@
 For a single-comparison FPR of $e$, the family-wise FPR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
-\hh{\Cref{tab:familywise} shows the impact the number of comparisons has on these published error rates.
+\Cref{tab:familywise} \hh{shows the impact the number of comparisons has on these published error rates.
 With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.
@@
 Under these constraints, the accuracy of an examination involving multiple comparisons between a wire and a tool will be low, as the number of candidate alignments that must be examined is high.
-Even the most innocuous example (small blade, only 2 cutting surfaces, and a relatively large wire) involved a minimum of \Sexpr{2*ncomp_min} completely independent comparisons.
-When we then consider that examiners would make cuts at multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons, the probability of a false identification becomes even more likely.
+Even the most innocuous example (small blade, only 2 cutting surfaces, and a relatively large wire) involves a minimum of \Sexpr{2*ncomp_min} %completely independent
+comparisons.
+When we then consider that examiners would make cuts under multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons, the probability of a false identification becomes even more likely.
 As a result, it is questionable whether wire comparisons made under current protocols are reliable enough to be presented at trial.
+
+\hh{Obviously, we need studies for wire evidence. It is also clear, based on the error rates, that larger studies are needed. }
+\hh{Going forward, we need to move away from  the binary assessment of 'match' or 'non-match' and quantify (a) the similarity of striations expressed between two pieces of evidence, and (b) the frequency of the observed pattern.
+Unusual striation patterns should be assigned more weight in the process.
+}
 

### 2024-02-06 PNAS/index.Rnw
Updates - now back to 3 pages

--- 2024-02-05 PNAS/index.Rnw
@@
 
-\firstpage[5]{5}
+\firstpage[7]{5}
 
@@
 One of the root causes~\citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid train bombing case was that the large size of the IAFIS database used to search for similar prints made it possible to locate `unusually' close non-matches.
-The probability of finding  close non-matches for an item of evidence is proportional to the size of the database available for searching; close non-matches occur more frequently as databases increase in size.
+% The probability of finding  close non-matches for an item of evidence is proportional to the size of the database available for searching; close non-matches occur more frequently as databases increase in size.
+\svp{As database size increases, so does the probability of finding a close non-match.}
 
@@
 The blade cuts will then be compared to the wire under a comparison microscope, though eventually, automatic comparison algorithms may also be validated for lab use.
-Each side of each blade cut will be compared to each side of the wire; depending on the tool design, there can be between two and four cutting surfaces in contact with the substrate.
+Each side of each blade cut will be compared to each side of the wire;
+\svp{different tool designs have between 2 and 4 cutting surfaces in contact with the substrate.}
+% depending on the tool design, there can be between two and four cutting surfaces in contact with the substrate.
 
@@
 For a single-comparison FPR of $e$, the family-wise FPR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
-\hh{\Cref{tab:familywise} shows the impact the number of comparisons has on these published error rates.
-With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.
-only the comparisons of one wire to a single-bladed tool are safely covered without exceeding an upper bound of 10\% for the family wise false identification error.}
+\Cref{tab:family-wise} \hh{shows the impact the number of comparisons has on these published error rates.
+With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.} \svp{even the simple example in this paper exceeds} \hh{an upper bound of 10\% for the family wise false identification error.}
 %\Cref{tab:Accuracy} shows the family-wise accuracy rate on the y-axis; if 90\% is an acceptable overall accuracy, 2 comparisons can be made when $e = 0.05$; increasing to 10 comparisons when $e =  0.01$.
 %Thus, at an error rate of \Sexpr{sprintf("e = %.2f", fpr)}, examiners can conduct comparisons before exceeding the family-wise FPR of 10\%.
-\hh{For a modestly sized database search of size 1000, the initial false positive rate should not exceed 1 in 10,000 to guarantee a family-wise total false identification error of at most 10\%. }
+\svp{To conduct a search of a modestly sized database with 1000 entries, the initial FPR cannot}
+% \hh{For a modestly sized database search of size 1000, the initial false positive rate should not
+\hh{exceed 1 in 10,000 to guarantee a family-wise total false identification error of at most 10\%. }
 
 Under these constraints, the accuracy of an examination involving multiple comparisons between a wire and a tool will be low, as the number of candidate alignments that must be examined is high.
-Even the most innocuous example (small blade, only 2 cutting surfaces, and a relatively large wire) involved a minimum of \Sexpr{2*ncomp_min} completely independent comparisons.
-When we then consider that examiners would make cuts at multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons, the probability of a false identification becomes even more likely.
+Even the most innocuous example (small blade, only 2 cutting surfaces, and a relatively large wire) involves a minimum of \Sexpr{2*ncomp_min} %completely independent
+comparisons.
+Examiners would make cuts under multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons \svp{and making a} false identification even more \svp{probable}.
 As a result, it is questionable whether wire comparisons made under current protocols are reliable enough to be presented at trial.
+
+\svp{Clearly, studies for wire evidence, and larger studies for striated evidence in general, are necessary.}
+% \hh{Obviously, we need studies for wire evidence. It is also clear, based on the error rates, that larger studies are needed. }
+\svp{Moving away from binary assessments toward quantification of striation similarity and observed pattern frequency will also reduce the severity of this issue and allow examiners to assign unusual striation patterns more weight in the process.}
+% \hh{Going forward, we need to move away from  the binary assessment of 'match' or 'non-match' and quantify (a) the similarity of striations expressed between two pieces of evidence, and (b) the frequency of the observed pattern.
+Unusual striation patterns should be assigned more weight in the process.
+}
 

### 2024-02-06 PNAS/index.Rnw
some color out

@@
 For a single-comparison FPR of $e$, the family-wise FPR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
-\Cref{tab:family-wise} \hh{shows the impact the number of comparisons has on these published error rates.
-With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.} \svp{even the simple example in this paper exceeds} \hh{an upper bound of 10\% for the family wise false identification error.}
+\Cref{tab:family-wise} shows the impact the number of comparisons has on these published error rates.
+With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.\ even the simple example in this paper exceeds an upper bound of 10\% for the family wise false identification error.
 %\Cref{tab:Accuracy} shows the family-wise accuracy rate on the y-axis; if 90\% is an acceptable overall accuracy, 2 comparisons can be made when $e = 0.05$; increasing to 10 comparisons when $e =  0.01$.

### 2024-02-06 PNAS/index.Rnw
color out

@@
 % The probability of finding  close non-matches for an item of evidence is proportional to the size of the database available for searching; close non-matches occur more frequently as databases increase in size.
-\svp{As database size increases, so does the probability of finding a close non-match.}
+As database size increases, so does the probability of finding a close non-match.
 
@@
 Each side of each blade cut will be compared to each side of the wire;
-\svp{different tool designs have between 2 and 4 cutting surfaces in contact with the substrate.}
+different tool designs have between 2 and 4 cutting surfaces in contact with the substrate.
 % depending on the tool design, there can be between two and four cutting surfaces in contact with the substrate.
@@
 %Thus, at an error rate of \Sexpr{sprintf("e = %.2f", fpr)}, examiners can conduct comparisons before exceeding the family-wise FPR of 10\%.
-\svp{To conduct a search of a modestly sized database with 1000 entries, the initial FPR cannot}
-% \hh{For a modestly sized database search of size 1000, the initial false positive rate should not
-\hh{exceed 1 in 10,000 to guarantee a family-wise total false identification error of at most 10\%. }
+To conduct a search of a modestly sized database with 1000 entries, the initial FPR cannot
+exceed 1 in 10,000 to guarantee a family-wise total false identification error of at most 10\%.
 
@@
 comparisons.
-Examiners would make cuts under multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons \svp{and making a} false identification even more \svp{probable}.
+Examiners would make cuts under multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons and making a false identification even more probable.
 As a result, it is questionable whether wire comparisons made under current protocols are reliable enough to be presented at trial.
 
-\svp{Clearly, studies for wire evidence, and larger studies for striated evidence in general, are necessary.}
-% \hh{Obviously, we need studies for wire evidence. It is also clear, based on the error rates, that larger studies are needed. }
-\svp{Moving away from binary assessments toward quantification of striation similarity and observed pattern frequency will also reduce the severity of this issue and allow examiners to assign unusual striation patterns more weight in the process.}
-% \hh{Going forward, we need to move away from  the binary assessment of 'match' or 'non-match' and quantify (a) the similarity of striations expressed between two pieces of evidence, and (b) the frequency of the observed pattern.
-Unusual striation patterns should be assigned more weight in the process.
-}
+Clearly, studies for wire evidence, and larger studies for striated evidence in general, are necessary.
+Moving away from binary assessments toward quantification of striation similarity and observed pattern frequency will also reduce the severity of this issue and allow examiners to assign unusual striation patterns more weight in the process.
 
@@
 
-\acknow{This work was funded (or partially funded) by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreements 70NANB15H176 and 70NANB20H019 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, Duke University, University of California Irvine, University of Virginia, West Virginia University, University of Pennsylvania, Swarthmore College and University of Nebraska, Lincoln.}
+\acknow{This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreements 70NANB15H176 and 70NANB20H019 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, Duke University, University of California Irvine, University of Virginia, West Virginia University, University of Pennsylvania, Swarthmore College and University of Nebraska, Lincoln.}
 

### 2024-03-26 PNAS/index.Rnw
horizontal lines in the table are now in the column
A few word tweaks

@@
 This mirrors the forensic examination process: the examiner visually rotates and shifts items under a comparison microscope to align two surfaces.
-In order to avoid false accusations and further erosions of public trust in science, we must address the problem of multiple comparisons and control their effect on false discovery rates.
+In order to avoid false accusations and the corresponding impact on public perception of forensics, we must address the problem of multiple comparisons and control their effect on false discovery rates.
 
@@
 
-In order to calculate the number of comparisons carried out in the course of one examination, we define $\ell$ to be the length of the blade cut, and $d$ to be the diameter of the wire.
+In order to calculate the number of comparisons carried out in the course of one examination, we define $b$ to be the length of the blade cut, and $d$ to be the diameter of the wire.
 We assume that the wire is covered with striations suitable for comparison across its full diameter $d$.
@@
 
-The first option gives us the maximum number of comparisons ($\ell/r - d/r + 1$), while the second option gives us the minimum number of comparisons $\ell/d$.
+The first option gives us the maximum number of comparisons ($b/r - d/r + 1$), while the second option gives us the minimum number of comparisons $b/d$.
 In the first case, sequential comparisons share much of the same physical data and are highly related; in the second case, no data are shared between physical comparisons and we can expect that they are statistically independent, though empirically there will be nonzero correlations due to physical similarities between striations.
 For simplicity, let us consider the number of comparisons to lie somewhere between these two estimates.
-Note that when $\ell/d \approx 1$, as in some toolmark comparisons, the upper number of comparisons goes to 1.
+Note that when $b/d \approx 1$, as in some toolmark comparisons, the upper number of comparisons goes to 1.
 Finally, we must consider the number of surfaces which must be compared: the wire may have one or two sets of striae and there may be  two to four blade cut surfaces to examine, depending on the tool.
@@
 <<simple-problem-setup, include=F>>=
-ell <- 15
+b <- 15
 d <- 2
 res <- 0.645/1000
-ncomp_min <- ell/d
-ncomp_max <- ell/res - d/res + 1
+ncomp_min <- b/d
+ncomp_max <- b/res - d/res + 1
 
@@
 Surfaces shown are rendered 2D topographical scans of a wire and blade cut taken with a confocal light microscope.
-(Bottom) RJ45 Crimp tool with a $\Sexpr{ell/10}$ cm razor blade used for cutting.
+(Bottom) RJ45 Crimp tool with a $\Sexpr{b/10}$ cm razor blade used for cutting.
 1 mm and 2 mm diameter aluminum wires cut with the pliers are shown in a box in the top right corner.}\label{fig:tool-pic}\label{fig:wire-blade}
@@
 
-Let us consider a wire-cutting tool with a $\Sexpr{ell/10}$ cm razor blade that meets a cast surface (one such tool is shown in \Cref{fig:tool-pic}); the wire is held against this rectangular cast surface as the blade is pushed into the wire, splitting it in two.
+Let us consider a wire-cutting tool with a $\Sexpr{b/10}$ cm razor blade that meets a cast surface (one such tool is shown in \Cref{fig:tool-pic}); the wire is held against this rectangular cast surface as the blade is pushed into the wire, splitting it in two.
 This is a minimal scenario - the wire will acquire striations from one side of the blade, while the blade itself has two cutting edges, which we will call side A and side B.
@@
 (that is, the wire appears to have been cut with a tool of similar configuration).
-In this example, $\ell = \Sexpr{ell}$ mm, $d = \Sexpr{d}$ mm, and there are at least $\ell/d = \Sexpr{ncomp_min}$ comparisons between a wire cut and a blade cut.
+In this example, $b = \Sexpr{b}$ mm, $d = \Sexpr{d}$ mm, and there are at least $b/d = \Sexpr{ncomp_min}$ comparisons between a wire cut and a blade cut.
 As there are two blade cuts (side A and side B), the minimal number of comparisons overall is \Sexpr{ncomp_min*2}, as these comparisons are non-overlapping and independent (on average).
@@
 In objective disciplines with standardized evaluation rules (e.g.\ DNA), these sources can be distinguished.
-However, in toolmark examination, no objective evaluation rules are used; examiners testify based on subjective, individual rules for how much similarity is sufficient for an identification.
-Assuming that lab procedure errors are not a factor in studies, we use reported error rates from open-set studies of striated evidence comparisons to estimate the coincidental match rate of a single wire-cut comparison.
-The three studies we consider \citep{bajic2020, mattijssen2021,best2022} have FPRs between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these studies yields an FPR of \Sexpr{sprintf("%.2f", fpr)}.
+However, in toolmark examination, no objective evaluation rules are used; examiners testify based on subjective rules for how much similarity is sufficient for an identification.
+Assuming that lab procedure errors are not a factor in studies, we use reported error rates from three open-set studies of striated evidence\citep{bajic2020, mattijssen2021,best2022} to obtain a ballpark estimate of the coincidental match rate of a single wire-cut comparison.
+These studies have FPRs between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these studies weighted by sample size yields an FPR of \Sexpr{sprintf("%.2f", fpr)}.
 For a single-comparison FPR of $e$, the family-wise FPR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
@@
 \begin{table}
+\hspace{-8pt}
 \scalebox{.95}{
@@
 First, examiners should report (or defense attorneys should request) the overall length or area of surfaces generated during the examination process, along with the total consecutive length or area of the recovered evidence.
-These pieces of information will take the place of $\ell$ and $d$ and facilitate calculation of examination-wide error rates.
+These pieces of information will take the place of $b$ and $d$ and facilitate calculation of examination-wide error rates.
 

### 2024-04-01 PNAS/reviews/response.qmd
A few minor tweaks + reviewer response letter

---
title: "Reviewer Response"
author: "Susan Vanderplas, Heike Hofmann, Alicia Carriquiry"
format: pdf
---

# General Comments

We thank both reviewers for their insightful comments. While we respond point-by-point below, the reviews of this paper are both careful, considerate, and helpful, and will serve to improve this paper.

# Reviewer 1

- **Points 1 and 2**: Existing alternative solutions to multiple comparisons problems and existing psychological literature that speaks to this kind of comparison task.

> We appreciate the reviewer's helpful suggestions about additional situations which may be similar and/or which may help to resolve the multiple comparison problem. One extremely unique factor of wire cuts that is not true of duct tape fits, bullets, or cartridge cases is that there is not any external structure which can assist with the multiple comparison problem. In duct tape, the tape has edges (and internal fibers) that provide structure. In bullets, there are lands and grooves which provide some structure and a way to sequence the lands according to the physical bullet orientation; this reduces a $n^2$ comparison problem to an $n$ comparison problem, where $n$ is the number of lands (though of course, database searches make even this a challenge). In cartridge cases, the extractor marks are of limited width and comparable size to the mechanism itself. In contrast, with wire cuts, there is a blade cut and a wire, and the wire could have been cut at any point along the blade; in most cases there is not any structure beyond the striation marks themselves. Thus, wire cuts serve as a uniquely problematic example within the wider issue of multiple comparisons in forensics. The Prusinowski paper is a good paper, but the holistic judgement process demonstrated in e.g. Fig 7 of the manuscript is predicated on edge alignment as a first step; no physical analogue to this exists for wire cuts. While the Prusinowski paper's method is one way to avoid cross-correlation, it inherently relies on an assumption which is not true in most striation marks: that the length of the striated region is the same for the evidence as for the comparison sample.

> Similarly, the psychological concepts relating to holistic judgements are difficult to apply to wire cuts and even bullets, because examination of the striations visually precludes examination of the entire entity, either due to the length (e.g. the blade cut is many times longer than the wire diameter) or due to curvature (in the case of bullets). While it is possible that a talented cognitive psychologist could provide an improved examination procedure that would systematize this sort of comparison, we aren't aware of any such procedures implemented in labs at the moment that are aimed at addressing this problem; in addition, as we are limited to 3 pages, it seems quite impossible to do the topic any justice. We would be most interested to collaborate with Reviewer 1 on an exloration of the zoomed-in vs. holistic cognition of striated mark comparisons, if they are a cognitive psychologist, as two of the three authors of this paper are statisticians who occasionally dabble in psychology-adjacent disciplines such as perception of statistical graphics.

### 2024-04-01 PNAS/index.Rnw
A few minor tweaks + reviewer response letter

@@
 Now that vast databases and efficient algorithms are routinely used in forensic evaluations to propose matches to crime scene items, the problem of close non-matches~\citep{pcast} due to multiple comparisons becomes critically important.
-This often ignored issue increases the false positive rate, and can contribute to the erosion of public trust in the justice system.
+This often ignored issue increases the false discovery rate, and can contribute to the erosion of public trust in the justice system through conviction of innocent individuals.
 The multiple comparison problem is not new: it has been raised in the past with regard to DNA~\citep{thompson2003} and latent print evaluations~ \citep{koehler2021}.
 One of the root causes~\citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid train bombing case was that the large size of the IAFIS database used to search for similar prints made it possible to locate `unusually' close non-matches.
-% The probability of finding  close non-matches for an item of evidence is proportional to the size of the database available for searching; close non-matches occur more frequently as databases increase in size.
 As database size increases, so does the probability of finding a close non-match.
@@
 This mirrors the forensic examination process: the examiner visually rotates and shifts items under a comparison microscope to align two surfaces.
-In order to avoid false accusations and the corresponding impact on public perception of forensics, we must address the problem of multiple comparisons and control their effect on false discovery rates.
+In order to avoid false accusations and the corresponding impact on public perception of forensics, we must address the problem of multiple comparisons in database and alignment searches and control their effect on false discovery rates.
 

### 2024-04-01 PNAS/index.Rnw
Fixing FPR -> FDR and overall -> familywise

@@
 When wires are cut, the tool produces striations on the cut surface; as in other forms of forensic analysis, these striation marks are used to connect the evidence to the source that created them.
-Here, we argue that the practice of comparing two wire cut surfaces introduces complexities not present in better-investigated forensic examination  of toolmarks such as those observed on bullets, as wire comparisons inherently require multiple distinct comparisons, increasing the expected false positive rate.
+Here, we argue that the practice of comparing two wire cut surfaces introduces complexities not present in better-investigated forensic examination  of toolmarks such as those observed on bullets, as wire comparisons inherently require multiple distinct comparisons, increasing the expected false discovery rate.
 We call attention to the multiple comparison problem in wire examination and relate it to other situations in forensics that involve multiple comparisons, such as database searches.
@@
 \section*{Probability of False Identifications}
-There are at least two components of the false positive rate (FPR): identifying two pieces of evidence that have similar characteristics but are from different sources (a coincidental match) and procedural failures (e.g.\ lab process errors) \citep[p 50]{pcast}.
+There are at least two components of the false discovery rate (FDR): identifying two pieces of evidence that have similar characteristics but are from different sources (a coincidental match) and procedural failures (e.g.\ lab process errors) \citep[p 50]{pcast}.
 In objective disciplines with standardized evaluation rules (e.g.\ DNA), these sources can be distinguished.
@@
 Assuming that lab procedure errors are not a factor in studies, we use reported error rates from three open-set studies of striated evidence\citep{bajic2020, mattijssen2021,best2022} to obtain a ballpark estimate of the coincidental match rate of a single wire-cut comparison.
-These studies have FPRs between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these studies weighted by sample size yields an FPR of \Sexpr{sprintf("%.2f", fpr)}.
-For a single-comparison FPR of $e$, the family-wise FPR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
+These studies have FDRs between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these studies weighted by sample size yields an FDR of \Sexpr{sprintf("%.2f", fpr)}.
+For a single-comparison FDR of $e$, the family-wise FDR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
 \Cref{tab:family-wise} shows the impact the number of comparisons has on these published error rates.
@@
 %\Cref{tab:Accuracy} shows the family-wise accuracy rate on the y-axis; if 90\% is an acceptable overall accuracy, 2 comparisons can be made when $e = 0.05$; increasing to 10 comparisons when $e =  0.01$.
-%Thus, at an error rate of \Sexpr{sprintf("e = %.2f", fpr)}, examiners can conduct comparisons before exceeding the family-wise FPR of 10\%.
-To conduct a search of a modestly sized database with 1000 entries, the initial FPR cannot
+%Thus, at an error rate of \Sexpr{sprintf("e = %.2f", fpr)}, examiners can conduct comparisons before exceeding the family-wise FDR of 10\%.
+To conduct a search of a modestly sized database with 1000 entries, the initial FDR cannot
 exceed 1 in 10,000 to guarantee a family-wise total false identification error of at most 10\%.
@@
 
-<<control-overall-fpr, fig.cap="Exploration of overall error rate estimates for different values of $e$, the FPR for a single comparison.
-Dotted vertical lines show false positive error rate estimates (from left to right, \\citep{best2022,bajic2020,mattijssen2021}).", fig.width = 8, fig.height = 4, include = F>>=
+<<control-overall-fpr, fig.cap="Exploration of familywise error rate estimates for different values of $e$, the FDR for a single comparison.
+Dotted vertical lines show false discovery error rate estimates (from left to right, \\citep{best2022,bajic2020,mattijssen2021}).", fig.width = 8, fig.height = 4, include = F>>=
 library(ggplot2)
@@
   coord_cartesian(xlim = c(0, 0.07), ylim = c(0, 1)) +
-  ylab("Probability of a false positive\n(N comparisons)") +
-  xlab("Probability of a false positive\n(1 comparison)") +
-  ggtitle("Wire Comparison: Controlling Overall False Positive Rate") +
+  ylab("Probability of a false discovery\n(N comparisons)") +
+  xlab("Probability of a false discovery\n(1 comparison)") +
+  ggtitle("Wire Comparison: Controlling Familywise False Discovery Rate") +
   geom_vline(xintercept = 0.02, linetype = "dashed", color = "grey40") +
@@
   gt() %>%
-  tab_spanner(label = "False Positive Error (in %) based on N comparisons", columns = matches("E10")) %>%
+  tab_spanner(label = "False Discovery Error (in %) based on N comparisons", columns = matches("E10")) %>%
   fmt_number(columns = c(E10, E100, E1000), decimals = 1) %>%
@@
   tab_header(
-    title = md("False identification rates increase with the number of comparisons made.")
+    title = md("False discovery rates increase with the number of comparisons made.")
   ) %>%
@@
     allowed = "Number of comparisons\nfor error below 10%",
-    FPR = "False Positive Rate"
+    FPR = "False Discovery Rate"
   )
@@
 # #  geom_line(data = error_res, linetype = 2, linewidth = 0.75, alpha =0.5) +
-#   ylab("Family-wise false positive rate E") +
+#   ylab("Family-wise false discovery rate E") +
 #   ggtitle("# Comparisons and Overall Accuracy") +
-#   scale_color_discrete("FP Rate for \nsingle comparison") +
+#   scale_color_discrete("FD Rate for \nsingle comparison") +
 #   theme_bw() +
@@
 \toprule
- &  & \multicolumn{3}{p{25mm}}{False Positives (\%) in N comparisons} &  \\
+ &  & \multicolumn{3}{p{25mm}}{False Discoveries (\%) in N comparisons} &  \\
 \cmidrule(lr){3-5}
-Study & FPR $e$ & $E_{10}$ & $E_{100}$ & $E_{1,000}$ & $E_N < 0.1$ \\
+Study & FDR $e$ & $E_{10}$ & $E_{100}$ & $E_{1,000}$ & $E_N < 0.1$ \\
 \midrule\addlinespace[2.5pt]
@@
 \end{tabular}}
-\caption{\label{tab:family-wise}Table showing the relationship between false positive rates and the chance of false identifications in $N$ comparisons for a set of different FPRs and different number of comparisons.
+\caption{\label{tab:family-wise}Table showing the relationship between false discovery rates and the chance of false identifications in $N$ comparisons for a set of different FDRs and different number of comparisons.
 The last column gives the number of comparisons allowed while ensuring a false identification percentage of at most 10\%.}

### 2024-04-01 PNAS/index.Rnw
Fixing overall

@@
 In this example, $b = \Sexpr{b}$ mm, $d = \Sexpr{d}$ mm, and there are at least $b/d = \Sexpr{ncomp_min}$ comparisons between a wire cut and a blade cut.
-As there are two blade cuts (side A and side B), the minimal number of comparisons overall is \Sexpr{ncomp_min*2}, as these comparisons are non-overlapping and independent (on average).
+As there are two blade cuts (side A and side B), the minimal number of comparisons is \Sexpr{ncomp_min*2}, as these comparisons are non-overlapping and independent (on average).
 
-Assuming a resolution of $\Sexpr{res*1000} \mu$m per pixel, the maximum number of comparisons per blade cut is around $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$; thus, we need $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall in order to find the optimal alignment between the wire and the blade cut.
+Assuming a resolution of $\Sexpr{res*1000} \mu$m per pixel, the maximum number of comparisons per blade cut is around $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$; thus, we need $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons in order to find the optimal alignment between the wire and the blade cut.
 These comparisons are implicit in the calculation of cross-correlation, which is the first and often the only step used to quantitatively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
@@
 With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.\ even the simple example in this paper exceeds an upper bound of 10\% for the family wise false identification error.
-%\Cref{tab:Accuracy} shows the family-wise accuracy rate on the y-axis; if 90\% is an acceptable overall accuracy, 2 comparisons can be made when $e = 0.05$; increasing to 10 comparisons when $e =  0.01$.
+%\Cref{tab:Accuracy} shows the family-wise accuracy rate on the y-axis; if 90\% is an acceptable familywise accuracy, 2 comparisons can be made when $e = 0.05$; increasing to 10 comparisons when $e =  0.01$.
 %Thus, at an error rate of \Sexpr{sprintf("e = %.2f", fpr)}, examiners can conduct comparisons before exceeding the family-wise FDR of 10\%.

### 2024-04-02 PNAS/reviews/response.qmd
Updates

@@
 We thank both reviewers for their insightful comments. 
-While we respond point-by-point below, the reviews of this paper are both careful, considerate, and helpful, and will serve to improve this paper.
+While we respond point-by-point below, the reviews of this paper are careful, considerate, and helpful, and will serve to improve this paper.
+We regret that due to the space constraints PNAS imposes on brief reports, we could not address all of the suggestions in the paper, but in many cases we would have liked to include the extra nuance you (quite correctly) identified in your reviews.
 
@@
 
-- **Points 1 and 2**: Existing alternative solutions to multiple comparisons problems and existing psychological literature that speaks to this kind of comparison task.
-
-> We appreciate the reviewer's helpful suggestions about additional situations which may be similar and/or which may help to resolve the multiple comparison problem.
+- **Point 1**: Existing alternative solutions to multiple comparisons problems.    
+We appreciate the reviewer's helpful suggestions about additional situations which may be similar and/or which may help to resolve the multiple comparison problem.
 One extremely unique factor of wire cuts that is not true of duct tape fits, bullets, or cartridge cases is that there is not any external structure which can assist with the multiple comparison problem.
@@
 Thus, wire cuts serve as a uniquely problematic example within the wider issue of multiple comparisons in forensics.
-The Prusinowski paper is a good paper, but the holistic judgement process demonstrated in e.g.
+The Prusinowski paper is excellent, but the holistic judgement process demonstrated in e.g.
 Fig 7 of the manuscript is predicated on edge alignment as a first step; no physical analogue to this exists for wire cuts.
 While the Prusinowski paper's method is one way to avoid cross-correlation, it inherently relies on an assumption which is not true in most striation marks: that the length of the striated region is the same for the evidence as for the comparison sample.
+We also note that previous papers by Prusinowski (e.g.
+2020) use cross-correlation functions to describe the alignment between two samples.
 
-> Similarly, the psychological concepts relating to holistic judgements are difficult to apply to wire cuts and even bullets, because examination of the striations visually precludes examination of the entire entity, either due to the length (e.g.
+- **Point 2** - Psychological concepts relating to visual comparison and decision making.    
+The psychological concepts relating to holistic judgments are difficult to apply to wire cuts and even bullets, because examination of the striations visually precludes examination of the entire entity, either due to the length (e.g.
 the blade cut is many times longer than the wire diameter) or due to curvature (in the case of bullets).
-While it is possible that a talented cognitive psychologist could provide an improved examination procedure that would systematize this sort of comparison, we aren't aware of any such procedures implemented in labs at the moment that are aimed at addressing this problem; in addition, as we are limited to 3 pages, it seems quite impossible to do the topic any justice.
-We would be most interested to collaborate with Reviewer 1 on an exloration of the zoomed-in vs.
+While it is possible that a talented cognitive psychologist could provide an improved examination procedure that would systematize this sort of comparison and enable holistic assessments, we aren't aware of any such procedures implemented in labs at the moment that are aimed at addressing this problem; in addition, as we are limited to 3 pages, it seems quite impossible to do the topic any justice.
+We would be most interested to collaborate with Reviewer 1 on an exploration of the zoomed-in vs.
 holistic cognition of striated mark comparisons, if they are a cognitive psychologist, as two of the three authors of this paper are statisticians who occasionally dabble in psychology-adjacent disciplines such as perception of statistical graphics.
+
+- **Point 3** - Clarity, precision, and minor issues. 
+    - We have addressed the first bullet point by changing the language to emphasize that the issue is both multiple comparisons and large databases (which lead to multiple comparisons). 
+    - While we agree that the partial fingerprint was a contributing factor, given that we only have 3 pages, we have chosen to focus primarily on the database size factor, which is most relevant to this paper.
+It is certainly true that there are some analogies between partial fingerprints and e.g.
+stranded wires which may each be considered "partial" due to the nature of how patterns may be transferred from the tool to the wire fragment, but this distinction isn't necessary to make the main point of the paper.
+    - We have added a description of what cross-correlation does, but the reference to the Vorburger paper will have to suffice for the definition because there is a lot of notation necessary to provide a definition of the theoretical quantity, and the computational implementations used in most algorithm implementations are even more complex to define and discuss.
+    - Brief reports are limited to 15 references, so we did not have the option to additionally include citations to the following algorithms/papers, which is far from an exhaustive list of papers using the cross-correlation function in forensics to align items or evaluate fit:
+
+        - Chu, W., Song, J., Vorburger, T., Yen, J., Ballou, S., & Bachrach, B.
+(2010).
+Pilot Study of Automated Bullet Signature Identification Based on Topography Measurements and Correlations*†.
+Journal of Forensic Sciences, 55(2), 341–347.
+https://doi.org/10/bpkwtx
+        - Hare, E., Hofmann, H., Carriquiry, A., & others.
+(2017).
+Automatic matching of bullet land impressions.
+The Annals of Applied Statistics, 11(4), 2332–2356.
+https://doi.org/10.1214/17-AOAS1080
+        - Krishnan, G., & Hofmann, H.
+(2019).
+Adapting the Chumbley Score to Match Striae on Land Engraved Areas (LEAs) of Bullets,.
+Journal of Forensic Sciences, 64(3), 728–740.
+https://doi.org/10.1111/1556-4029.13950
+        - Ma, L., Song, J., Whitenton, E., Zheng, A., Vorburger, T., & Zhou, J.
+(2004).
+NIST Bullet Signature Measurement System for RM (Reference Material) 8240 Standard Bullets.
+Journal of Forensic Sciences, 49(4), 1–11.
+https://doi.org/10/cdsbv8
+        - Chen, Z., Chu, W., Soons, J.
+A., Thompson, R.
+M., Song, J., & Zhao, X.
+(2019).
+Fired bullet signature correlation using the Congruent Matching Profile Segments (CMPS) method.
+Forensic Science International, 305, 109964.
+https://doi.org/10/gn649n
+        - Ott, D., Thompson, R., & Song, J.
+(2017).
+Applying 3D measurements and computer matching algorithms to two firearm examination proficiency tests.
+Forensic Science International, 271, 98–106.
+https://doi.org/10/gn649s
+        - Prusinowski, M., Brooks, E., & Trejos, T.
+(2020).
+Development and validation of a systematic approach for the quantitative assessment of the quality of duct tape physical fits.
+Forensic Science International, 307, 110103.
+https://doi.org/10.1016/j.forsciint.2019.110103
+        - Song, H., & Song, J.
+(2022).
+Virtual image standard (VIS) for performance evaluation of the congruent matching cells (CMC) algorithms in firearm evidence identifications.
+Journal of Forensic Sciences, 67(4), 1417–1430.
+https://doi.org/10.1111/1556-4029.15026
+        - Song, J., Chu, W., Tong, M., & Soons, J.
+(2014).
+3D topography measurements on correlation cells—A new approach to forensic ballistics identifications.
+Measurement Science and Technology, 25(6), 064005.
+https://doi.org/10/gn65br
+        - Song, J., & Song, H. (2023). Reporting likelihood ratio for casework in firearm evidence identification. Journal of Forensic Sciences, 68(2), 399–406. https://doi.org/10.1111/1556-4029.15186
+        - Song, J.-F., Vorburger, T.
+V., Ma, L., Libert, J.
+M., & Ballou, S.
+M.
+(2005).
+A Metric for the Comparison of Surface Topographies of Standard Reference Material (SRM) Bullets and Casings.
+NIST.
+https://www.nist.gov/publications/metric-comparison-surface-topographies-standard-reference-material-srm-bullets-and
+        - Tong, M., Song, J., Chu, W., & Thompson, R.
+M.
+(2014).
+Fired Cartridge Case Identification Using Optical Images and the Congruent Matching Cells (CMC) Method.
+Journal of Research of the National Institute of Standards and Technology, 119, 575.
+https://doi.org/10/gn65bm
+        - Venkatasubramanian, G., Hegde, V., Padi, S., Iyer, H., & Herman, M.
+(2021).
+Comparing footwear impressions that are close non-matches using correlation-based approaches.
+Journal of Forensic Sciences, 66(3), 890–909.
+https://doi.org/10/gpjnz3
+        - Wen, Z., Curran, J.
+M., & Wevers, G.
+(2023).
+Shoeprint image retrieval and crime scene shoeprint image linking by using convolutional neural network and normalized cross correlation.
+Science & Justice, 63(4), 439–450.
+https://doi.org/10.1016/j.scijus.2023.04.014
+        - Zemmels, J., VanderPlas, S., & Hofmann, H.
+(2023).
+A Study in Reproducibility: The Congruent Matching Cells Algorithm and cmcR Package.
+The R Journal, 14(4), 79–102.
+https://doi.org/10.32614/RJ-2023-014
+        - Zhang, N.
+F.
+(2021).
+Statistical models for firearm and tool mark image comparisons based on the congruent matching cells (CMC) method.
+Forensic Science International, 326, 110912.
+https://doi.org/10.1016/j.forsciint.2021.110912
+        - Zhang, H., Zhu, J., Hong, R., Wang, H., Sun, F., & Malik, A.
+(2021).
+Convergence-improved congruent matching cells (CMC) method for firing pin impression comparison.
+Journal of Forensic Sciences, 66(2), 571–582.
+https://doi.org/10/gpjn2f
+
+    - We have changed the terminology to use $d$ for the wire cross-section length, and $b$ for the blade cut length. Hopefully this is somewhat more intuitive. 
+    - We would love to use a more verbose explanation with less sophisticated vocabulary, but we are working within the confines of the 3 page limit imposed by Brief Reports - this requires linguistic brevity which sometimes necessitates a more terse explanation.
+    - The AFTE guidelines (emphasis added) are provided below.
+As interpretation is subjective and based on the individual's training and experience according to the AFTE guidelines, the rules used to evaluate the evidence are (it stands to reason) also subjective and based on individual training and experience, e.g.
+with the best agreement demonstrated between toolmarks produced by different tools, and consistent with the agreement demonstrated by toolmarks produced by the same tool.
+
+    > The three principles of the AFTE Theory of Identification as it Relates to Toolmarks are:
+    > 1.
+The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the unique surface contours of two toolmarks are in sufficient agreement.
+    > 2.
+This sufficient agreement is related to the significant duplication of random toolmarks as evidenced by the correspondence of a pattern or combination of patterns of surface contours.
+Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows.
+Specifically, the relative height or depth, width, curvature and spatial relationship of the individual peaks, ridges and furrows within one set of surface contours are defined and compared to the corresponding features in the second set of surface contours.
+Agreement is significant when it **exceeds the best agreement demonstrated between toolmarks known to have been produced by different tools** and is consistent with agreement demonstrated by toolmarks known to have been produced by the same tool.
+The statement that sufficient agreement exists between two toolmarks means that the agreement is of a quantity and quality that the likelihood another tool could have made the mark is so remote as to be considered a practical impossibility.
+    > 3. Currently the interpretation of individualization/identification is **subjective in nature**, founded on scientific principles and **based on the examiners training and experience.**
+
+    - We have added additional information about data pooling - the calculations were indeed weighted by sample size.
+
+# Reviewer 2
+
+- We have attempted to streamline the number of terms used for error rates, using false discovery rate and familywise false discovery rate consistently throughout the paper.
+In the conclusion, we reference error rates more generally, but have clarified that we mean both false discovery and false elimination errors.
+
+- There are two separate questions in the second comment:
+    - Implicit multiple comparisons in different toolmark disciplines.    
+    While brevity forced us to distinguish implicitly between multiple comparisons in bullets or cartridge cases and multiple comparisons in wire cuts - the latter are a much more extreme case than the former, as there is some external structure present in both cartridge cases and bullets which limits the scale of comparisons necessary.
+For instance, in bullets, there are the land-groove transition areas, and we know that each land must occur in sequence; these physical constraints reduce the potential number of comparisons over the surface of the bullet.
+In wire cuts, however, there is an extremely limited external structure that would assist with identifying which part of the blade was used to cut the wire - the only indicators are the striation marks themselves and the potential configuration of the blade - whether two cutting surfaces or 4 contributed to the wire cut, and how those surfaces might be aligned.
+This magnifies the problem (which is still very present in bullets) to an extreme which is useful for presentation in a brief report such as this one - the problem itself is relatively simple to explain and sufficiently extreme to illustrate the multiple comparison problem and motivate its importance when examining the scientific basis of forensic pattern comparisons.
+    - Suggesting that the establishment of firearms examination bona fides is a done deal.    
+    While we do not agree that forensic firearm examination has been scientifically shown to be reliable to the levels necessary in the 2009 NAS report and 2015 PCAST report, within the category of striated toolmark evidence, firearm examination is certainly the most well-developed and well-studied discipline.
+We are well aware of the statistical issues with firearm studies, having published several papers ourselves highlighting these problems; with that said, we must also acknowledge that some more recent studies have fewer design flaws than historically well-cited studies.
+Even with the problems shared by these more recent studies (participant self-selection, lack of reporting of participant dropout, lack of casework level realism, participant blindness to the study, breadth of firearm/ammunition selection), the study design is sufficient to provide some information about error rates.
+This information may not be generalizable to a population or useful for statistical inference, but the studies we cite are sufficient to provide descriptive statistics; these statistics can be aggregated to provide a useful hypothetical error rate that represents a "ballpark" estimate of striated error rates.
+We have worked through the math with a whole host of different error rates and the conclusions are similarly dramatic even if the error rate changes by a factor of 10.
+Thus, our decision to ground the hypothetical error rate in estimates from firearm studies is motivated primarily by an attempt to be charitable and suggest that the multiple comparison problem is still a problem even if we accept the low error rates in better-designed (but still suboptimal) published studies.
+- Putative tool - we have changed the language, using 'recovered tool' to indicate a tool recovered from the suspect.
+
+- Figure 1 clutter and "two sides"    
+We've cleaned up the figure as much as possible.
+We agree that it is cluttered, but we needed to increase the information density of the picture in order to meet the space constraints of PNAS brief reports.
+We have simplified the tool labels, using $b$ for the length of the blade cut and $d$ for the diameter of the wire.
+Hopefully these are easier to read.
+The tool shown in Figure 1 has a single blade with two sides (which are shown as inset figures between the legs of the tool).
+In other tool configurations, it is possible to have two blades that meet at the tip, and thus to have 4 cutting surfaces, 2 of which may be represented on each half of the cut wire.
+As we mention ("number of surfaces"), this setup may make additional comparisons necessary, further compounding the situation we illustrate in the paper.
+Examiners may also take multiple blade cuts at different angles, which only makes the problem more serious.
+We have written this paper to account for the simplest possible wire cut solution with a tool that has minimal cutting surfaces, and our results are still catastrophic when we try to enforce even moderate controls on false discovery errors.
+
+- "on average"    
+Strictly speaking, because striated surfaces tend to have periodic oscillations due to the striations, non-overlapping comparisons may have some dependence.
+Striations are typically random, and so the expected value of this dependence is zero, even if for a single instance of a striated blade, there may be non-zero correlation between non-overlapping comparisons.
+We've examined and established this result empirically, but wanted to be technically precise while maintaining readability for an audience who are not all statistically sophisticated enough to identify the nuances of dependence structures in this type of data.
+
+- "family-wise error rate"    
+We've added a citation to Tukey's 1953 JASA paper on multiple comparisons.
+
+- "two components."    
+The problem is that with a subjective set of evaluation rules that are examiner-dependent, the two sources cannot be separated.
+We address this a few sentences later when we say "Assuming that lab procedure errors are not a factor in studies, we use reported error rates from three open-set studies of striated evidence"...that is, we are explicitly assuming that the procedure errors are not a factor in the studies we are using.
+
+- Relevance of the results of firearms studies.    
+We agree that the firearms studies have their weak points, which we've briefly summarized in a previous comment response.
+We decided after some discussion that using an error rate grounded in some empirical study was more concrete (and thus easier to relate to) than using an example error rate not connected to reality  (which, statistically, would have worked just as well without the baggage).
+There are certainly pros and cons to each approach, but we wanted to make this paper as concrete as possible with the hopes of motivating action on the issue of multiple comparisons more generally and wire cut forensic analysis specifically.
+
+- We've removed the reference to Type II studies, as it is a distraction from the main point of this paper.

### 2024-04-02 PNAS/index.Rnw
Updates

@@
 
-\firstpage[7]{5}
+\firstpage[4]{5}
 
@@
 Compounding this issue, the use of algorithms also results in a large number of comparisons that are not obvious to the user.
-For example, the cross-correlation function~\citep{vorburgerApplicationsCrosscorrelationFunctions2011} was one of the first measures proposed for quantifying the similarity between two patterns in response to the 2009 NRC report~\citep{nas2009}, and continues to be used in many pattern searching algorithms to find the best alignment between two images and to quantify their overall similarity.
+For example, the cross-correlation function~\citep{vorburgerApplicationsCrosscorrelationFunctions2011}, which computes the correlation for each alignment of two sequences, was one of the first measures proposed to quantify the similarity between two patterns in response to the 2009 NRC report~\citep{nas2009}, and continues to be used in many pattern searching algorithms to find the best alignment between two images and to quantify their overall similarity.
 Finding the best alignment often consists in sliding one surface across the whole length (for one-dimensional patterns, such as striations) or area (for two dimensional sources, such as impression marks) of the other item while keeping track of the value of a similarity measure.
@@
 
-A forensics examiner tasked with determining whether a wire in evidence was cut by a specific tool will create one or more blade cuts using the putative tool, which are then compared to the cut surface of the wire.
-These cuts are made in a sheet of material similar to the wire's composition, and may be performed at multiple angles, as the angle of the tool to the substrate can affect which striations are recorded on the substrate surface.
+A forensics examiner tasked with determining whether a wire in evidence was cut by a recovered tool will create one or more blade cuts, which are then compared to the cut surface of the wire recovered from the scene.
+These cuts are made in a sheet of material matching the wire composition, and may be performed at multiple angles, as the angle of the tool to the substrate can affect which striations are recorded on the substrate surface.
 The blade cuts will then be compared to the wire under a comparison microscope, though eventually, automatic comparison algorithms may also be validated for lab use.
@@
 different tool designs have between 2 and 4 cutting surfaces in contact with the substrate.
-% depending on the tool design, there can be between two and four cutting surfaces in contact with the substrate.
 
@@
 As the number of comparisons increases, the probability of encountering a coincidental match increases.
-Statisticians call this the \textit{family-wise error rate} $E$; it is an important quantity to control when conducting a series ("family") of tests.
+Statisticians call this the \textit{family-wise error rate} $E$; it is an important quantity to control when conducting a series ("family") of tests\citep{tukey1953multiple}.
 
-\section*{Probability of False Identifications}
+\section*{Probability of False Discoveries}
 There are at least two components of the false discovery rate (FDR): identifying two pieces of evidence that have similar characteristics but are from different sources (a coincidental match) and procedural failures (e.g.\ lab process errors) \citep[p 50]{pcast}.
@@
 \Cref{tab:family-wise} shows the impact the number of comparisons has on these published error rates.
-With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.\ even the simple example in this paper exceeds an upper bound of 10\% for the family wise false identification error.
-%\Cref{tab:Accuracy} shows the family-wise accuracy rate on the y-axis; if 90\% is an acceptable familywise accuracy, 2 comparisons can be made when $e = 0.05$; increasing to 10 comparisons when $e =  0.01$.
-%Thus, at an error rate of \Sexpr{sprintf("e = %.2f", fpr)}, examiners can conduct comparisons before exceeding the family-wise FDR of 10\%.
+With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.\ even the simple example in this paper exceeds an upper bound of 10\% for the family wise false discovery error.
 To conduct a search of a modestly sized database with 1000 entries, the initial FDR cannot
-exceed 1 in 10,000 to guarantee a family-wise total false identification error of at most 10\%.
+exceed 1 in 10,000 to guarantee a family-wise total false discovery error of at most 10\%.
 
 Under these constraints, the accuracy of an examination involving multiple comparisons between a wire and a tool will be low, as the number of candidate alignments that must be examined is high.
-Even the most innocuous example (small blade, only 2 cutting surfaces, and a relatively large wire) involves a minimum of \Sexpr{2*ncomp_min} %completely independent
-comparisons.
-Examiners would make cuts under multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons and making a false identification even more probable.
+Even the most innocuous example (small blade, only 2 cutting surfaces, and a relatively large wire) involves a minimum of \Sexpr{2*ncomp_min}  comparisons.
+Examiners would make cuts under multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons and making a false discovery even more probable.
 As a result, it is questionable whether wire comparisons made under current protocols are reliable enough to be presented at trial.
@@
 \end{tabular}}
-\caption{\label{tab:family-wise}Table showing the relationship between false discovery rates and the chance of false identifications in $N$ comparisons for a set of different FDRs and different number of comparisons.
-The last column gives the number of comparisons allowed while ensuring a false identification percentage of at most 10\%.}
+\caption{\label{tab:family-wise}Table showing the relationship between false discovery rates and the chance of a false discovery in $N$ comparisons for a set of different FDRs and different number of comparisons.
+The last column gives the number of comparisons allowed while ensuring a familywise false discovery percentage of at most 10\%.}
 \end{table}
@@
 The striated surfaces are of orders of magnitude different lengths, but represent the best estimate of the error rate for striated materials.
-New studies should be Type II studies as defined by Koehler \citep{koehlerIntuitiveErrorRate2017a}, designed to assess the actual error rate when examiners are making difficult comparisons.
+New studies should be designed to assess error rates (false discovery and false elimination) when examiners are making difficult comparisons.
 
@@
 Additionally, the number of results used for further manual comparison should also be reported.
-For example, if a firearms examiner searches a local NIBIN database with 1000 entries, requests the 20 closest matches to her evidence, and then carries out a physical examination of five exemplars from the list of 20, all of those values should be clearly reported to enable estimation of the examination-wise error rate.
-This will help make the multiple comparison issue accessible to everyone involved in evaluating the value of forensic evidence, from examiners to lawyers to jurors and judges.
+For example, if a firearms examiner searches a local NIBIN database with 1000 entries, requests the 20 closest matches to her evidence, and then carries out a physical examination of five exemplars from the list of 20, all of those values should be clearly reported to enable estimation of the familywise error rate.
+This will help make the multiple comparison issue accessible to everyone involved in evaluating the value of forensic evidence: examiners, lawyers, jurors, and judges.
 

### 2024-04-02 PNAS/index-old.tex
Update with old/changed tex files

--- 2024-04-01 PNAS/index.Rnw
@@
-\documentclass[9pt,twocolumn,twoside]{pnas-new}
+\documentclass[9pt,twocolumn,twoside]{pnas-new}\usepackage[]{graphicx}\usepackage[dvipsnames]{xcolor}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 % Use the lineno option to display guide line numbers if required.
@@
 When wires are cut, the tool produces striations on the cut surface; as in other forms of forensic analysis, these striation marks are used to connect the evidence to the source that created them.
-Here, we argue that the practice of comparing two wire cut surfaces introduces complexities not present in better-investigated forensic examination  of toolmarks such as those observed on bullets, as wire comparisons inherently require multiple distinct comparisons, increasing the expected false discovery rate.
+Here, we argue that the practice of comparing two wire cut surfaces introduces complexities not present in better-investigated forensic examination  of toolmarks such as those observed on bullets, as wire comparisons inherently require multiple distinct comparisons, increasing the expected false positive rate.
 We call attention to the multiple comparison problem in wire examination and relate it to other situations in forensics that involve multiple comparisons, such as database searches.
@@
 \doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
-<<setup, include = F>>=
-library(knitr)
-opts_chunk$set(error=F, warning=F, message = F, dpi=300, echo = F)
-options(digits = 4)
-@
 
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 Now that vast databases and efficient algorithms are routinely used in forensic evaluations to propose matches to crime scene items, the problem of close non-matches~\citep{pcast} due to multiple comparisons becomes critically important.
-This often ignored issue increases the false discovery rate, and can contribute to the erosion of public trust in the justice system through conviction of innocent individuals.
+This often ignored issue increases the false positive rate, and can contribute to the erosion of public trust in the justice system.
 The multiple comparison problem is not new: it has been raised in the past with regard to DNA~\citep{thompson2003} and latent print evaluations~ \citep{koehler2021}.
 One of the root causes~\citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid train bombing case was that the large size of the IAFIS database used to search for similar prints made it possible to locate `unusually' close non-matches.
-As database size increases, so does the probability of finding a close non-match.
+% The probability of finding  close non-matches for an item of evidence is proportional to the size of the database available for searching; close non-matches occur more frequently as databases increase in size.
+\svp{As database size increases, so does the probability of finding a close non-match.}
 
@@
 This mirrors the forensic examination process: the examiner visually rotates and shifts items under a comparison microscope to align two surfaces.
-In order to avoid false accusations and the corresponding impact on public perception of forensics, we must address the problem of multiple comparisons in database and alignment searches and control their effect on false discovery rates.
+In order to avoid false accusations and further erosions of public trust in science, we must address the problem of multiple comparisons and control their effect on false discovery rates.
 
@@
 Each side of each blade cut will be compared to each side of the wire;
-different tool designs have between 2 and 4 cutting surfaces in contact with the substrate.
+\svp{different tool designs have between 2 and 4 cutting surfaces in contact with the substrate.}
 % depending on the tool design, there can be between two and four cutting surfaces in contact with the substrate.
@@
 
-In order to calculate the number of comparisons carried out in the course of one examination, we define $b$ to be the length of the blade cut, and $d$ to be the diameter of the wire.
+In order to calculate the number of comparisons carried out in the course of one examination, we define $\ell$ to be the length of the blade cut, and $d$ to be the diameter of the wire.
 We assume that the wire is covered with striations suitable for comparison across its full diameter $d$.
@@
 
-The first option gives us the maximum number of comparisons ($b/r - d/r + 1$), while the second option gives us the minimum number of comparisons $b/d$.
+The first option gives us the maximum number of comparisons ($\ell/r - d/r + 1$), while the second option gives us the minimum number of comparisons $\ell/d$.
 In the first case, sequential comparisons share much of the same physical data and are highly related; in the second case, no data are shared between physical comparisons and we can expect that they are statistically independent, though empirically there will be nonzero correlations due to physical similarities between striations.
 For simplicity, let us consider the number of comparisons to lie somewhere between these two estimates.
-Note that when $b/d \approx 1$, as in some toolmark comparisons, the upper number of comparisons goes to 1.
+Note that when $\ell/d \approx 1$, as in some toolmark comparisons, the upper number of comparisons goes to 1.
 Finally, we must consider the number of surfaces which must be compared: the wire may have one or two sets of striae and there may be  two to four blade cut surfaces to examine, depending on the tool.
@@
 \subsection*{A concrete example}
-
-<<simple-problem-setup, include=F>>=
-b <- 15
-d <- 2
-res <- 0.645/1000
-ncomp_min <- b/d
-ncomp_max <- b/res - d/res + 1
-
-# From 3 studies
-errors <- c(20, 56, 1)
-knm <- c(2842, 774, 220)
-
-fpr <- sum(errors)/sum(knm)
-@
 
@@
 Surfaces shown are rendered 2D topographical scans of a wire and blade cut taken with a confocal light microscope.
-(Bottom) RJ45 Crimp tool with a $\Sexpr{b/10}$ cm razor blade used for cutting.
+(Bottom) RJ45 Crimp tool with a $1.5$ cm razor blade used for cutting.
 1 mm and 2 mm diameter aluminum wires cut with the pliers are shown in a box in the top right corner.}\label{fig:tool-pic}\label{fig:wire-blade}
@@
 
-Let us consider a wire-cutting tool with a $\Sexpr{b/10}$ cm razor blade that meets a cast surface (one such tool is shown in \Cref{fig:tool-pic}); the wire is held against this rectangular cast surface as the blade is pushed into the wire, splitting it in two.
+Let us consider a wire-cutting tool with a $1.5$ cm razor blade that meets a cast surface (one such tool is shown in \Cref{fig:tool-pic}); the wire is held against this rectangular cast surface as the blade is pushed into the wire, splitting it in two.
 This is a minimal scenario - the wire will acquire striations from one side of the blade, while the blade itself has two cutting edges, which we will call side A and side B.
 A blade cut of a sheet of aluminum will thus produce two striated edges corresponding to side A and side B which are compared to cut wires to assess similarity.
-We also have a 12 gauge aluminum wire ($\Sexpr{d}$ mm diameter) which may have been cut by the wire-cutting tool described above.
+We also have a 12 gauge aluminum wire ($2$ mm diameter) which may have been cut by the wire-cutting tool described above.
 Class characteristics, which are shared by all tools of similar manufacture, appear to match:
@@
 (that is, the wire appears to have been cut with a tool of similar configuration).
-In this example, $b = \Sexpr{b}$ mm, $d = \Sexpr{d}$ mm, and there are at least $b/d = \Sexpr{ncomp_min}$ comparisons between a wire cut and a blade cut.
-As there are two blade cuts (side A and side B), the minimal number of comparisons overall is \Sexpr{ncomp_min*2}, as these comparisons are non-overlapping and independent (on average).
+In this example, $\ell = 15$ mm, $d = 2$ mm, and there are at least $\ell/d = 7.5$ comparisons between a wire cut and a blade cut.
+As there are two blade cuts (side A and side B), the minimal number of comparisons overall is 15, as these comparisons are non-overlapping and independent (on average).
 
-Assuming a resolution of $\Sexpr{res*1000} \mu$m per pixel, the maximum number of comparisons per blade cut is around $\Sexpr{format(round(ncomp_max/1000)*1000, big.mark=',')}$; thus, we need $\Sexpr{format(round(ncomp_max/1000)*2000, big.mark=',')}$ comparisons overall in order to find the optimal alignment between the wire and the blade cut.
+Assuming a resolution of $0.645 \mu$m per pixel, the maximum number of comparisons per blade cut is around $20,000$; thus, we need $40,000$ comparisons overall in order to find the optimal alignment between the wire and the blade cut.
 These comparisons are implicit in the calculation of cross-correlation, which is the first and often the only step used to quantitatively assess the similarity between striated evidence such as bullets, aperture shear, and firing pin impressions.
@@
 \section*{Probability of False Identifications}
-There are at least two components of the false discovery rate (FDR): identifying two pieces of evidence that have similar characteristics but are from different sources (a coincidental match) and procedural failures (e.g.\ lab process errors) \citep[p 50]{pcast}.
+There are at least two components of the false positive rate (FPR): identifying two pieces of evidence that have similar characteristics but are from different sources (a coincidental match) and procedural failures (e.g.\ lab process errors) \citep[p 50]{pcast}.
 In objective disciplines with standardized evaluation rules (e.g.\ DNA), these sources can be distinguished.
-However, in toolmark examination, no objective evaluation rules are used; examiners testify based on subjective rules for how much similarity is sufficient for an identification.
-Assuming that lab procedure errors are not a factor in studies, we use reported error rates from three open-set studies of striated evidence\citep{bajic2020, mattijssen2021,best2022} to obtain a ballpark estimate of the coincidental match rate of a single wire-cut comparison.
-These studies have FDRs between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these studies weighted by sample size yields an FDR of \Sexpr{sprintf("%.2f", fpr)}.
-For a single-comparison FDR of $e$, the family-wise FDR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
-\Cref{tab:family-wise} shows the impact the number of comparisons has on these published error rates.
-With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.\ even the simple example in this paper exceeds an upper bound of 10\% for the family wise false identification error.
+However, in toolmark examination, no objective evaluation rules are used; examiners testify based on subjective, individual rules for how much similarity is sufficient for an identification.
+Assuming that lab procedure errors are not a factor in studies, we use reported error rates from open-set studies of striated evidence comparisons to estimate the coincidental match rate of a single wire-cut comparison.
+The three studies we consider \citep{bajic2020, mattijssen2021,best2022} have FPRs between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these studies yields an FPR of 0.02.
+For a single-comparison FPR of $e$, the family-wise FPR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
+\Cref{tab:family-wise} \hh{shows the impact the number of comparisons has on these published error rates.
+With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.} \svp{even the simple example in this paper exceeds} \hh{an upper bound of 10\% for the family wise false identification error.}
 %\Cref{tab:Accuracy} shows the family-wise accuracy rate on the y-axis; if 90\% is an acceptable overall accuracy, 2 comparisons can be made when $e = 0.05$; increasing to 10 comparisons when $e =  0.01$.
-%Thus, at an error rate of \Sexpr{sprintf("e = %.2f", fpr)}, examiners can conduct comparisons before exceeding the family-wise FDR of 10\%.
-To conduct a search of a modestly sized database with 1000 entries, the initial FDR cannot
-exceed 1 in 10,000 to guarantee a family-wise total false identification error of at most 10\%.
+%Thus, at an error rate of sprintf("e = %.2f", fpr), examiners can conduct comparisons before exceeding the family-wise FPR of 10\%.
+\svp{To conduct a search of a modestly sized database with 1000 entries, the initial FPR cannot}
+% \hh{For a modestly sized database search of size 1000, the initial false positive rate should not
+\hh{exceed 1 in 10,000 to guarantee a family-wise total false identification error of at most 10\%. }
 
 Under these constraints, the accuracy of an examination involving multiple comparisons between a wire and a tool will be low, as the number of candidate alignments that must be examined is high.
-Even the most innocuous example (small blade, only 2 cutting surfaces, and a relatively large wire) involves a minimum of \Sexpr{2*ncomp_min} %completely independent
+Even the most innocuous example (small blade, only 2 cutting surfaces, and a relatively large wire) involves a minimum of 15 %completely independent
 comparisons.
-Examiners would make cuts under multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons and making a false identification even more probable.
+Examiners would make cuts under multiple angles \citep{baikerToolmarkVariabilityQuality2015}, increasing the number of comparisons \svp{and making a} false identification even more \svp{probable}.
 As a result, it is questionable whether wire comparisons made under current protocols are reliable enough to be presented at trial.
 
-Clearly, studies for wire evidence, and larger studies for striated evidence in general, are necessary.
-Moving away from binary assessments toward quantification of striation similarity and observed pattern frequency will also reduce the severity of this issue and allow examiners to assign unusual striation patterns more weight in the process.
-
-<<control-overall-fpr, fig.cap="Exploration of familywise error rate estimates for different values of $e$, the FDR for a single comparison.
-Dotted vertical lines show false discovery error rate estimates (from left to right, \\citep{best2022,bajic2020,mattijssen2021}).", fig.width = 8, fig.height = 4, include = F>>=
-library(ggplot2)
-library(tidyr)
-library(dplyr)
-# library(latex2exp)
-library(geomtextpath)
-single_rmp = exp(seq(-1000, 0, .1))
-N = c(10, 100, 1000, 10000)
-
-res = crossing(single_rmp, N) %>%
-  mutate(overall = 1 - (1 - single_rmp)^N)
-
-studies <- tibble(single_rmp = c( 0.004545,0.00704, 0.0724), study = c("Best (2022)","Bajic (2020)",  "Mattijssen (2021)"), y = c(-0.025, .15, 0.6), hjust = c(0, 0, 1))
-ggplot(res, aes(x = single_rmp, y = overall, group = factor(N), color = factor(N))) +
-  coord_cartesian(xlim = c(0, 0.07), ylim = c(0, 1)) +
-  ylab("Probability of a false discovery\n(N comparisons)") +
-  xlab("Probability of a false discovery\n(1 comparison)") +
-  ggtitle("Wire Comparison: Controlling Familywise False Discovery Rate") +
-  geom_vline(xintercept = 0.02, linetype = "dashed", color = "grey40") +
-  geom_vline(xintercept = studies$single_rmp, linetype = "51", color = "grey50") +
-  geom_label(data = studies[1,], aes(x = single_rmp, y = y, label = study), inherit.aes = F, hjust = studies$hjust[1], vjust = 0) +
-  geom_label(data = studies[2,], aes(x = single_rmp, y = y, label = study), inherit.aes = F, hjust = studies$hjust[2], vjust = 0) +
-  geom_label(data = studies[3,], aes(x = single_rmp, y = y, label = study), inherit.aes = F, hjust = studies$hjust[3], vjust = 0) +
-  annotate(geom = "label", x = 0.02, y = 0.1, label = "Pooled error rate \nused in this paper", color = "grey40", hjust=0) +
-  scale_color_discrete("# comparisons  ") +
-  geom_line(linewidth=1) +
-  theme_bw() +
-  theme(legend.position = c(1, 0),
-        legend.justification = c(1, 0),
-        legend.background = element_rect(color = "grey40", fill=alpha("white", 0.1))) +
-  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = seq(0, 1, .05)) +
-  scale_x_continuous(breaks = c(0, 0.02, 0.04, 0.06, studies$single_rmp),
-                     labels=c(0, 0.02, 0.04, 0.06, "0.004545", "\n0.00704","0.0724"))
-#  geom_hline(yintercept = 0.1, colour="black")
-@
-
-<<Accuracy, eval=FALSE,  fig.cap="Probability of accurate decisions for $N$ comparisons for different error rates.", fig.width = 8, fig.height = 4>>=
-
-comp_allowed <- function(e, at_most=0.1) {
-  floor(log(1-at_most)/log(1-e))
+\svp{Clearly, studies for wire evidence, and larger studies for striated evidence in general, are necessary.}
+% \hh{Obviously, we need studies for wire evidence. It is also clear, based on the error rates, that larger studies are needed. }
+\svp{Moving away from binary assessments toward quantification of striation similarity and observed pattern frequency will also reduce the severity of this issue and allow examiners to assign unusual striation patterns more weight in the process.}
+% \hh{Going forward, we need to move away from  the binary assessment of 'match' or 'non-match' and quantify (a) the similarity of striations expressed between two pieces of evidence, and (b) the frequency of the observed pattern.
+Unusual striation patterns should be assigned more weight in the process.
 }
 
-family_wise <- function(e, N) {
-  1 - (1 - e)^N
-}
-
-r_vals <- tibble(r = c(0.004545, 0.00704, 0.02, 0.0724, 0.001, 0.0001, 0.00001),
-                 Study = c("Best (2022)","Bajic (2020)", "Pooled Error",  "Mattijssen (2021)", "", "", "")) %>% arrange(desc(r))
-#                 rlab = sprintf("e==%0.05f", r)) %>% arrange(desc(r))
-
-r_vals <- r_vals %>% mutate(
-#  E1 = r,
-  E10 = family_wise(r, 10),
-  E100 = family_wise(r, 100),
-  E1000 = family_wise(r, 1000),
-  allowed = comp_allowed(r)
-)
-
-data <- r_vals %>%
-  mutate(
-    FPR = c("7.24%", "2.00%", "0.70%", "0.45%", "1 in 1,000", "1 in 10,000", "1 in 100,000")
-  ) %>%
-  select(Study, FPR, E10, E100, E1000, allowed)
-
-library(gt)
-
-gt_tab <- data %>%
-  mutate(across(c(E10, E100, E1000), function(x) 100*x)) %>%
-  gt() %>%
-  tab_spanner(label = "False Discovery Error (in %) based on N comparisons", columns = matches("E10")) %>%
-  fmt_number(columns = c(E10, E100, E1000), decimals = 1) %>%
-  fmt_number(columns = c(allowed), sep_mark = ",", decimals=0) %>%
-  tab_header(
-    title = md("False discovery rates increase with the number of comparisons made.")
-  ) %>%
- cols_label(
-    E10 = "N = 10",
-    E100 = "N = 100",
-    E1000 = "N = 1,000",
-    allowed = "Number of comparisons\nfor error below 10%",
-    FPR = "False Discovery Rate"
-  )
-
-# include modified latex based on latex produced below:
-as_latex(gt_tab)
-
-# r_vals <- r_vals %>% mutate(
-#   study = sprintf("%s:e=%s", study, r)
-# )
-
-# res2 <- crossing(r = r_vals$r,
-#                  N = c(seq(1, 11, .01),seq(11.1, 109.9, .1), seq(110, 1100, 1))) %>%
-#   mutate(overall = (1 - r)^N) %>%
-#   left_join(r_vals)
-#
-
-# breakpt <- .9
-#
-# error_res <- filter(res2, lead(overall, 1) <= breakpt)
-# res2 <- filter(res2, overall >= breakpt)
-#
-# res2sum <- res2 %>% group_by(r) %>% filter(overall == min(overall))
-#
-#
-# ggplot(data = res2, aes(x = N, y = 1-overall, color = study, label = rlab)) +
-# #  geom_textline(parse = T, gap = F, straight = F,
-# #                padding = unit(1, "mm"), vjust = 0, linewidth = 1) +
-#   geom_line(linewidth = 1) +
-#   geom_line(linewidth = 1, data = error_res, alpha = 0.5) +
-#   xlab("Number of comparisons n") +
-#   scale_x_log10(
-#     breaks = c(1, 10000, res2sum$N),
-#     labels = c("1", "10000", round(res2sum$N)[-4], "1.4"),
-#     minor_breaks = c((1:9)[-2],10*(3:9), 100*3:9, 1000*1:9)
-#   ) +
-#   geom_hline(yintercept = 0.1, linewidth = 1) +
-# #  geom_label(
-# #    data = res2sum,
-# #    aes(x = N, y = breakpt, label = paste0("N == ", round(N))),
-# #    parse = T, vjust = 1, hjust = 1) +
-#   geom_point(data = res2sum,
-#      aes(x = N, y = 0.1, color=study),
-#      size=10) +
-# #  geom_segment(data = res2sum,
-# #     aes(x = N, xend=N, y = -Inf, yend=0.1, color=study),
-# #     linewidth=0.5) +
-# #  geom_line(data = error_res, linetype = 2, linewidth = 0.75, alpha =0.5) +
-#   ylab("Family-wise false discovery rate E") +
-#   ggtitle("# Comparisons and Overall Accuracy") +
-#   scale_color_discrete("FD Rate for \nsingle comparison") +
-#   theme_bw() +
-#   coord_cartesian(ylim = c(0, .15)) +
-#   theme(axis.text.y = element_text(hjust = 0.5, vjust = 0.5, angle = 90))
-
-@
-
 \begin{table}
-\hspace{-8pt}
 \scalebox{.95}{
@@
 \toprule
- &  & \multicolumn{3}{p{25mm}}{False Discoveries (\%) in N comparisons} &  \\
+ &  & \multicolumn{3}{p{25mm}}{False Positives (\%) in N comparisons} &  \\
 \cmidrule(lr){3-5}
-Study & FDR $e$ & $E_{10}$ & $E_{100}$ & $E_{1,000}$ & $E_N < 0.1$ \\
+Study & FPR $e$ & $E_{10}$ & $E_{100}$ & $E_{1,000}$ & $E_N < 0.1$ \\
 \midrule\addlinespace[2.5pt]
@@
 \end{tabular}}
-\caption{\label{tab:family-wise}Table showing the relationship between false discovery rates and the chance of false identifications in $N$ comparisons for a set of different FDRs and different number of comparisons.
+\caption{\label{tab:family-wise}Table showing the relationship between false positive rates and the chance of false identifications in $N$ comparisons for a set of different FPRs and different number of comparisons.
 The last column gives the number of comparisons allowed while ensuring a false identification percentage of at most 10\%.}
@@
 First, examiners should report (or defense attorneys should request) the overall length or area of surfaces generated during the examination process, along with the total consecutive length or area of the recovered evidence.
-These pieces of information will take the place of $b$ and $d$ and facilitate calculation of examination-wide error rates.
+These pieces of information will take the place of $\ell$ and $d$ and facilitate calculation of examination-wide error rates.
 
@@
 
-\acknow{This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreements 70NANB15H176 and 70NANB20H019 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, Duke University, University of California Irvine, University of Virginia, West Virginia University, University of Pennsylvania, Swarthmore College and University of Nebraska, Lincoln.}
+\acknow{This work was funded (or partially funded) by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreements 70NANB15H176 and 70NANB20H019 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, Duke University, University of California Irvine, University of Virginia, West Virginia University, University of Pennsylvania, Swarthmore College and University of Nebraska, Lincoln.}
 

### 2024-04-02 PNAS/index-old.tex
Removing coloration from index-old and remaking the diff file

@@
 Each side of each blade cut will be compared to each side of the wire;
-\svp{different tool designs have between 2 and 4 cutting surfaces in contact with the substrate.}
-% depending on the tool design, there can be between two and four cutting surfaces in contact with the substrate.
+different tool designs have between 2 and 4 cutting surfaces in contact with the substrate.
 
@@
 For a single-comparison FPR of $e$, the family-wise FPR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
-\Cref{tab:family-wise} \hh{shows the impact the number of comparisons has on these published error rates.
-With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.} \svp{even the simple example in this paper exceeds} \hh{an upper bound of 10\% for the family wise false identification error.}
-%\Cref{tab:Accuracy} shows the family-wise accuracy rate on the y-axis; if 90\% is an acceptable overall accuracy, 2 comparisons can be made when $e = 0.05$; increasing to 10 comparisons when $e =  0.01$.
-%Thus, at an error rate of sprintf("e = %.2f", fpr), examiners can conduct comparisons before exceeding the family-wise FPR of 10\%.
-\svp{To conduct a search of a modestly sized database with 1000 entries, the initial FPR cannot}
-% \hh{For a modestly sized database search of size 1000, the initial false positive rate should not
-\hh{exceed 1 in 10,000 to guarantee a family-wise total false identification error of at most 10\%. }
+\Cref{tab:family-wise} shows the impact the number of comparisons has on these published error rates.
+With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.
+even the simple example in this paper exceeds an upper bound of 10\% for the family wise false identification error.
+To conduct a search of a modestly sized database with 1000 entries, the initial FPR cannot
+exceed 1 in 10,000 to guarantee a family-wise total false identification error of at most 10\%.
 

### 2024-05-15 PNAS/index.Rnw
Trying to ensure the PNAS version works the first time... lol
Movinng all the pnas-required files into the damn folder so that everything works... hopefully?
Final updates, hopefully.

@@
 
+\graphicspath{{../fig/}}
 \title{Hidden Multiple Comparisons Increase Forensic Error Rates}
@@
 
-\section*{Calculating the Number of Comparisons}
+\section*{Methods: Calculating the Number of Comparisons}
 
@@
 \centering
-\includegraphics[width=\columnwidth]{../fig/wire-blade-comparison-label.png}
-\includegraphics[width=\columnwidth]{../fig/wire_cutter_wires.png}
+\includegraphics[width=\columnwidth]{full-wire-cutter-fig.pdf}
 \caption{(Top) A comparison between a wire and a blade cut requires sliding the wire along the entire blade cut length to determine the best match (or whether there is a match).
@@
 \Cref{tab:family-wise} shows the impact the number of comparisons has on these published error rates.
-With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.\ even the simple example in this paper exceeds an upper bound of 10\% for the family wise false discovery error.
+With an error rate of 0.007, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.\ even the simple example in this paper exceeds an upper bound of 10\% for the family wise false discovery error.
 To conduct a search of a modestly sized database with 1000 entries, the initial FDR cannot
